---
title: "Model SIM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model SIM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(magrittr)
```

Please find below guidelines how godley package can be used to create model SIM — the simplest model with government money described by *Wynne Godley* and *Marc Lavoie* in Chapter 3 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*

```{r setup}
library(godley)
```

### Base scenario
First you need to create empty SFC model:
```{r}
# Create empty model
model_sim <- create_model(name = "SFC SIM")
```

Next you have to add variables:
```{r}
# Add variables
model_sim <- add_variable(model_sim, "C_d", desc = "Consumption demand by households")
model_sim <- add_variable(model_sim, "C_s", desc = "Consumption supply")
model_sim <- add_variable(model_sim, "G_s", desc = "Government supply")
model_sim <- add_variable(model_sim, "H_h", desc = "Cash money held by households")
model_sim <- add_variable(model_sim, "H_s", desc = "Cash money supplied by the government")
model_sim <- add_variable(model_sim, "N_d", desc = "Demand for labor")
model_sim <- add_variable(model_sim, "N_s", desc = "Supply of labor")
model_sim <- add_variable(model_sim, "T_d", desc = "Taxes, demand")
model_sim <- add_variable(model_sim, "T_s", desc = "Taxes, supply")
model_sim <- add_variable(model_sim, "Y", desc = "Income = GDP")
model_sim <- add_variable(model_sim, "Yd", desc = "Disposable income of households")
model_sim <- add_variable(model_sim, "alpha1", init = 0.6, desc = "Propensity to consume out of income")
model_sim <- add_variable(model_sim, "alpha2", init = 0.4, desc = "Propensity to consume out of wealth")
model_sim <- add_variable(model_sim, "theta", init = 0.2, desc = "Tax rate")
model_sim <- add_variable(model_sim, "G_d", init = 20, desc = "Government demand")
model_sim <- add_variable(model_sim, "W", init = 1, desc = "Wage rate")
```

You also have to add equations:
```{r}
# Add equations
model_sim <- add_equation(model_sim, "C_s = C_d", desc = "Consumption")
model_sim <- add_equation(model_sim, "G_s = G_d")
model_sim <- add_equation(model_sim, "T_s = T_d")
model_sim <- add_equation(model_sim, "N_s = N_d")
model_sim <- add_equation(model_sim, "Yd = W * N_s - T_s")
model_sim <- add_equation(model_sim, "T_d = theta * W * N_s")
model_sim <- add_equation(model_sim, "C_d = alpha1 * Yd + alpha2 * H_h[-1]")
model_sim <- add_equation(model_sim, "H_s = G_d - T_d + H_s[-1]")
model_sim <- add_equation(model_sim, "H_h = Yd - C_d + H_h[-1]")
model_sim <- add_equation(model_sim, "Y = C_s + G_s")
model_sim <- add_equation(model_sim, "N_d = Y/W")
model_sim <- add_equation(model_sim, "H_s = H_h", desc = "Money equilibrium", hidden = TRUE)
```

Now you can simulate model (in this example for the next 100 periods using Newton method):
```{r}
# Simulate model
model_sim <- simulate_scenario(model_sim, scenario = "baseline", max_iter = 350, periods = 100, 
                               hidden_tol = 0.1, tol = 1e-08, method = "Newton")
```

Based on calculated simulation you can display results for desired set of variables:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_sim, scenario = c("baseline"), from = 1, to = 50, 
                expressions = c("Y", "C_d", "G_s"))
```

### Shock scenario
With godley package you can simulate how a shock will affect the base scenario. In this example you will implement a permanent increase in government expenditures. \
\
First you need to create empty shock object:
```{r}
# Create empty shock
shock_sim <- create_shock()
```

Next you need to add appropriate equation:
```{r}
# Add shock equation with increased government expenditures
shock_sim <- add_shock(shock_sim, equation = "G_d = 25", 
                       desc = "Increase in government expenditures", start = 5, end = 50)
```

Later you need to create the shock scenario:
```{r}
# Create new scenario with this shock
model_sim <- add_scenario(model_sim, name = "expansion", origin = "baseline", 
                          origin_period = 1, shock = shock_sim)
```

Now you can simulate new shock scenario:
```{r}
# Simulate shock
model_sim <- simulate_scenario(model_sim, scenario = "expansion", max_iter = 350, periods = 100, 
                               hidden_tol = 0.1, tol = 1e-08, method = "Newton")
```

You can also compare results from shock and base scenarios:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_sim, scenario = c("baseline", "expansion"), from = 1, to = 50, 
                expressions = c("Y"))
```

```{r, out.width="100%"}
plot_simulation(model = model_sim, scenario = c("baseline", "expansion"), from = 1, to = 50, 
                expressions = c("C_d"))
```

```{r, out.width="100%"}
plot_simulation(model = model_sim, scenario = c("baseline", "expansion"), from = 1, to = 50, 
                expressions = c("G_s"))
```

### References
More information about model SIM can be found in Chapter 3 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*
