---
title: "Model SIMEX"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model SIMEX}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Please find below guidelines how godley package can be used to create model SIMEX — model with government money and expectations described by *Wynne Godley* and *Marc Lavoie* in Chapter 3 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*

```{r setup}
library(godley)
```

### Base scenario
First you need to create empty SFC model:
```{r}
# Create empty model
model_simex <- create_model(name = "SFC SIMEX")
```

Next you have to add variables:
```{r}
# Add variables
model_simex <- add_variable(model_simex, "C_d", desc = "Consumption demand by households")
model_simex <- add_variable(model_simex, "C_s", desc = "Consumption supply")
model_simex <- add_variable(model_simex, "G_s", desc = "Government supply")
model_simex <- add_variable(model_simex, "T_d", desc = "Taxes, demand")
model_simex <- add_variable(model_simex, "T_s", desc = "Taxes, supply")
model_simex <- add_variable(model_simex, "N_d", desc = "Demand for labor")
model_simex <- add_variable(model_simex, "N_s", desc = "Supply of labor")
model_simex <- add_variable(model_simex, "H_h", desc = "Cash money held by households")
model_simex <- add_variable(model_simex, "H_s", desc = "Cash money supplied by the government")
model_simex <- add_variable(model_simex, "H_d", desc = "Cash money demanded by the government")
model_simex <- add_variable(model_simex, "Y", desc = "Income = GDP")
model_simex <- add_variable(model_simex, "Yd", desc = "Disposable income of households")
model_simex <- add_variable(model_simex, "Yd_e", desc = "Expected disposable income of households")
model_simex <- add_variable(model_simex, "alpha1", init = 0.6, 
                            desc = "Propensity to consume out of income")
model_simex <- add_variable(model_simex, "alpha2", init = 0.4, 
                            desc = "Propensity to consume out of wealth")
model_simex <- add_variable(model_simex, "theta", init = 0.2, desc = "Tax rate")
model_simex <- add_variable(model_simex, "G_d", init = 20, desc = "Government demand")
model_simex <- add_variable(model_simex, "W", init = 1, desc = "Wage rate")

```

You also have to add equations:
```{r}
# Add equations
model_simex <- add_equation(model_simex, "C_s = C_d", desc = "Consumption")
model_simex <- add_equation(model_simex, "G_s = G_d")
model_simex <- add_equation(model_simex, "T_s = T_d")
model_simex <- add_equation(model_simex, "N_s = N_d")
model_simex <- add_equation(model_simex, "Yd = W * N_s - T_s")
model_simex <- add_equation(model_simex, "T_d = theta * W * N_s")
model_simex <- add_equation(model_simex, "C_d = alpha1 * Yd_e + alpha2 * H_h[-1]")
model_simex <- add_equation(model_simex, "H_s = G_d - T_d + H_s[-1]")
model_simex <- add_equation(model_simex, "H_h = Yd - C_d + H_h[-1]")
model_simex <- add_equation(model_simex, "Y = C_s + G_s")
model_simex <- add_equation(model_simex, "N_d = Y/W")
model_simex <- add_equation(model_simex, "H_d = Yd_e - C_d + H_h[-1]")
model_simex <- add_equation(model_simex, "Yd[-1] = Yd_e")
model_simex <- add_equation(model_simex, "H_s = H_h", desc = "Money equilibrium", hidden = TRUE)
```

Now you can simulate model (in this example for the next 100 periods using Newton method):
```{r}
# Simulate model
model_simex <- simulate_scenario(model_simex, scenario = "baseline", max_iter = 350, periods = 100, 
                                 hidden_tol = 0.1, tol = 1e-08, method = "Newton")
```

Based on calculated simulation you can display results for desired set of variables:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_simex, scenario = c("baseline"), from = 1, to = 50, 
                expressions = c("Y", "C_d", "C_s / alpha1"))
```

### Shock scenario
You can simulate how a shock will affect the base scenario. In this example you will implement an increased government expenditures. \
\
First you need to create empty shock:
```{r}
# Create empty shock
shock_simex <- create_shock()
```

Next you need to add appropriate equation:
```{r}
# Add shock equation with increased government expenditures
shock_simex <- add_shock(shock_simex, equation = "G_d = 25", 
                         desc = "Increase in government expenditures", start = 5, end = 50)
```

Then you need to create new shock scenario:
```{r}
# Create new scenario with this shock
model_simex <- add_scenario(model_simex, name = "expansion", origin = "baseline", 
                            origin_period = 100, shock = shock_simex)
```

Now you can simulate new shock scenario:
```{r}
# Simulate shock
model_simex <- simulate_scenario(model_simex, scenario = "expansion", max_iter = 350, periods = 100, 
                                 hidden_tol = 0.1, tol = 1e-08, method = "Newton")
```

You can also display results from the shock scenario:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_simex, scenario = c("expansion"), from = 1, to = 50, 
                expressions = c("Y", "C_d", "C_s / alpha1"))
```

### References
More information about model SIMEX can be found in Chapter 3 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*