---
title: "Model PC"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model PC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Please find below guidelines how godley package can be used to create model PC — model with government money with portfolio choice described by *Wynne Godley* and *Marc Lavoie* in Chapter 4 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*

```{r setup}
library(godley)
```

### Base scenario
First you need to create empty SFC model:
```{r}
# Create empty model
model_pc <- create_model(name = "SFC PC")
```

Next you have to add variables:
```{r}
# Add variables
model_pc <- add_variable(model_pc, "B_cb", desc = "") 
model_pc <- add_variable(model_pc, "H_s", desc = "") 
model_pc <- add_variable(model_pc, "B_s", desc = "") 
model_pc <- add_variable(model_pc, "B_h", desc = "") 
model_pc <- add_variable(model_pc, "H_h1", desc = "") 
model_pc <- add_variable(model_pc, "H_h", desc = "") 
model_pc <- add_variable(model_pc, "C", desc = "") 
model_pc <- add_variable(model_pc, "V", desc = "") 
model_pc <- add_variable(model_pc, "T_x", desc = "") 
model_pc <- add_variable(model_pc, "Y", desc = "Income = GDP") 
model_pc <- add_variable(model_pc, "Yd", desc = "Disposable income of households") 
model_pc <- add_variable(model_pc, "alpha1", init = 0.6, desc = "Propensity to consume out of income") 
model_pc <- add_variable(model_pc, "alpha2", init = 0.4, desc = "Propensity to consume out of wealth") 
model_pc <- add_variable(model_pc, "theta", init = 0.2, desc = "Tax rate") 
model_pc <- add_variable(model_pc, "r", init = 0.025, desc = "") 
model_pc <- add_variable(model_pc, "G", init = 20, desc = "Government demand") 
model_pc <- add_variable(model_pc, "lambda0", init = 0.635, desc = "") 
model_pc <- add_variable(model_pc, "lambda1", init = 0.05, desc = "") 
model_pc <- add_variable(model_pc, "lambda2", init = 0.01, desc = "")
```

you also have to add equations:
```{r}
# Add equations
model_pc <- add_equation(model_pc, "Y = C + G", desc = "") 
model_pc <- add_equation(model_pc, "Yd = Y - T_x + r[-1] * B_h[-1]") 
model_pc <- add_equation(model_pc, "T_x = theta * (Y + r[-1] * B_h[-1])") 
model_pc <- add_equation(model_pc, "V = V[-1] + (Yd - C)") 
model_pc <- add_equation(model_pc, "C = alpha1 * Yd + alpha2 * V[-1]") 
model_pc <- add_equation(model_pc, "H_h = V - B_h") 
model_pc <- add_equation(model_pc, "H_h1 = V * ((1 - lambda0) - lambda1 * r + lambda2 * ( Yd/V ))") 
model_pc <- add_equation(model_pc, "B_h = V * (lambda0 + lambda1 * r - lambda2 * ( Yd/V ))") 
model_pc <- add_equation(model_pc, "B_s = B_s[-1] + (G + r[-1] * B_s[-1]) - (T_x + r[-1] * B_cb[-1])") 
model_pc <- add_equation(model_pc, "H_s = H_s[-1] + B_cb - B_cb[-1]") 
model_pc <- add_equation(model_pc, "B_cb = B_s - B_h") 
model_pc <- add_equation(model_pc, "H_h = H_s", hidden = T)
```

Now you can simulate model (in this example for the next 100 periods using Newton method):
```{r}
# Simulate model
model_pc <- simulate_scenario(model_pc, scenario = "baseline", max_iter = 350, periods = 100, 
                              hidden_tol = 0.1, tol = 1e-08, method = "Newton")
```

Based on calculated simulation you can display results for desired set of variables:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_pc, scenario = c("baseline"), from = 1, to = 100, 
                expressions = c("B_h / V"))
```

```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_pc, scenario = c("baseline"), from = 1, to = 100, 
                expressions = c("H_h / V"))
```

### Shock scenario
you can simulate how a shock will affect the base scenario. In this example you will implement an increased rate of interest on bills. \
\
First you need to create empty shock:
```{r}
# Create empty shock
shock_pc <- create_shock()
```

Next you need to add appropriate equation:
```{r}
# Add shock equation with increased rate of interest on bills
shock_pc <- add_shock(shock_pc, equation = "r = 0.035", 
                      desc = "Increase in the rate of interest on bills", start = 5, end = 50)
```

Then you need to create new shock scenario:
```{r}
# Create new scenario with this shock
model_pc <- add_scenario(model_pc, name = "expansion", origin = "baseline", 
                         origin_period = 100, shock = shock_pc)
```

Now you can simulate new shock scenario:
```{r}
# Simulate shock
model_pc <- simulate_scenario(model_pc, scenario = "expansion", max_iter = 350, periods = 100, 
                              hidden_tol = 0.1, tol = 1e-08, method = "Newton")
```

You can also display results from shock scenario:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_pc, scenario = c("expansion"), from = 1, to = 50, 
                expressions = c("B_h / V"))
```

```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_pc, scenario = c("expansion"), from = 1, to = 50, 
                expressions = c("H_h / V"))
```

### References
More information about model PC can be found in Chapter 4 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*
