---
title: "Model BMW"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model BMW}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Please find below guidelines how godley package can be used to create model BMW — the simplest model with private bank money described by *Wynne Godley* and *Marc Lavoie* in Chapter 7 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*

```{r setup}
library(godley)
```

### Base scenario
First you need to create empty SFC model:
```{r}
# Create empty model
model_bmw <- create_model(name = "SFC BMW")
```

Next you have to add variables:
```{r}
# Add variables
model_bmw <- model_bmw |>
  add_variable("rl", init = 0.025) |>
  add_variable("alpha0", init = 20) |>
  add_variable("alpha1", init = 0.75) |>
  add_variable("alpha2", init = 0.10) |>
  add_variable("delta", init = 0.10) |>
  add_variable("gamma", init = 0.15) |>
  add_variable("kappa", init = 1) |>
  add_variable("pr", init = 1) |>
  add_variable("Nd", init = .001) |>
  add_variable("Ns", init = .001) |>
  add_variable("Y", init = .001) |>
  add_variable("Cs") |>
  add_variable("Cd") |>
  add_variable("Is") |>
  add_variable("Id") |>
  add_variable("Ls") |>
  add_variable("Ld") |>
  add_variable("WBd") |>
  add_variable("AF") |>
  add_variable("K") |>
  add_variable("YD") |>
  add_variable("WBs") |>
  add_variable("rm") |>
  add_variable("Mh") |>
  add_variable("Ms") |>
  add_variable("W") |>
  add_variable("DA") |>
  add_variable("KT")
```

You also have to add equations:
```{r}
# Add equations
model_bmw <- model_bmw |>
  add_equation("Cs = Cd") |>
  add_equation("Is = Id") |>
  add_equation("Ns = Nd") |>
  add_equation("Ls = Ls[-1] + Ld - Ld[-1]") |>
  add_equation("Y = Cs + Is") |>
  add_equation("WBd = Y - rl[-1] * Ld[-1] - AF") |>
  add_equation("AF = delta * K[-1]") |>
  add_equation("Ld = Ld[-1] + Id - AF") |>
  add_equation("YD = WBs + rm[-1] * Mh[-1]") |>
  add_equation("Mh = Mh[-1] + YD - Cd") |>
  add_equation("Ms = Ms[-1] + Ls - Ls[-1]") |>
  add_equation("rm = rl") |>
  add_equation("WBs = W * Ns") |>
  add_equation("Nd = Y / pr") |>
  add_equation("W = WBd / Nd") |>
  add_equation("Cd = alpha0 + alpha1 * YD + alpha2 * Mh[-1]") |>
  add_equation("K = K[-1] + Id - DA") |>
  add_equation("DA = delta * K[-1]") |>
  add_equation("KT = kappa * Y[-1]") |>
  add_equation("Id = gamma * (KT - K[-1]) + DA") |>
  add_equation("Ms = Mh", hidden = TRUE)
```

Now you can simulate model (in this example for the next 100 periods using Gauss method):
```{r}
# Simulate model
model_bmw <- simulate_scenario(model_bmw,
  scenario = "baseline", max_iter = 350, periods = 100,
  hidden_tol = 0.1, tol = 1e-08, method = "Gauss"
)
```

Based on calculated simulation you can display results for desired set of variables:
```{r, out.width="100%"}
# Plot results
plot_simulation(
  model = model_bmw, scenario = c("baseline"), from = 1, to = 50,
  expressions = c("Cd", "YD")
)
```

Note: In above example new pipe operator was applied. It does not work in older R versions (before R 4.1). 

### Shock scenarios
With godley you can simulate how a shock will affect the base scenario. 

#### Shock 1
In this example you will implement increased autonomous consumption expenditures. \
\
First you need to create empty shock:
```{r}
# Create empty shock
shock_bmw <- create_shock()
```

Next you need to add appropriate equation:
```{r}
# Add shock equation with increase in autonomous consumption expenditures
shock_bmw <- add_shock(shock_bmw,
  equation = "alpha0 = 30",
  desc = "Increase in autonomous consumption expenditures", start = 5, end = 50
)
```

Then you need to create new shock scenario:
```{r}
# Create new scenario with this shock
model_bmw <- add_scenario(model_bmw,
  name = "expansion1", origin = "baseline",
  origin_period = 100, shock = shock_bmw
)
```

Now you can simulate new shock scenario:
```{r}
# Simulate shock
model_bmw <- simulate_scenario(model_bmw,
  scenario = "expansion1", max_iter = 350, periods = 100,
  hidden_tol = 0.1, tol = 1e-08, method = "Newton"
)
```

You can also plot results from the shock scenarios:
```{r, out.width="100%"}
# Plot results
plot_simulation(
  model = model_bmw, scenario = c("expansion1"), from = 1, to = 50,
  expressions = c("Cd", "YD")
)
```

#### Shock 2
In this example you will implement increased propensity to save. \
\
First you need to create empty shock:
```{r}
# Create empty shock
shock_bmw <- create_shock()
```

Next you need to add appropriate equation:
```{r}
# Add shock equation with increased propensity to save
shock_bmw <- add_shock(shock_bmw,
  equation = "alpha1 = 0.7",
  desc = "Increase in the propensity to save", start = 5, end = 50
)
```

Later you need to create new shock scenario:
```{r}
# Create new scenario with this shock
model_bmw <- add_scenario(model_bmw,
  name = "expansion2", origin = "baseline",
  origin_period = 100, shock = shock_bmw
)
```

Now you can simulate new shock scenario:
```{r}
# Simulate shock
model_bmw <- simulate_scenario(model_bmw,
  scenario = "expansion2", max_iter = 350, periods = 100,
  hidden_tol = 0.1, tol = 1e-08, method = "Newton"
)
```

You can also plot results from the shock scenarios:
```{r, out.width="100%"}
# Plot results
plot_simulation(
  model = model_bmw, scenario = c("expansion2"), from = 1, to = 50,
  expressions = c("Cd", "YD")
)
```

### References
More information about model BMW can be found in Chapter 7 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*
