---
title: "Model BMW"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model BMW}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Please find below guidelines how godley package can be used to create model BMW — the simplest model with private bank money described by *Wynne Godley* and *Marc Lavoie* in Chapter 7 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*

```{r setup}
library(godley)
```

### Base scenario
First you need to create empty SFC model:
```{r}
# Create empty model
model_bmw <- create_model(name = "SFC BMW")
```

Next you have to add variables:
```{r}
# Add variables
model_bmw <- add_variable(model_bmw, "rl", init = 0.025) 
model_bmw <- add_variable(model_bmw, "alpha0", init = 20) 
model_bmw <- add_variable(model_bmw, "alpha1", init = 0.75) 
model_bmw <- add_variable(model_bmw, "alpha2", init = 0.10) 
model_bmw <- add_variable(model_bmw, "delta", init = 0.10) 
model_bmw <- add_variable(model_bmw, "gamma", init = 0.15) 
model_bmw <- add_variable(model_bmw, "kappa", init = 1) 
model_bmw <- add_variable(model_bmw, "pr", init = 1) 
model_bmw <- add_variable(model_bmw, "Nd", init = .001) 
model_bmw <- add_variable(model_bmw, "Ns", init = .001) 
model_bmw <- add_variable(model_bmw, "Y", init = .001) 
model_bmw <- add_variable(model_bmw, "Cs") 
model_bmw <- add_variable(model_bmw, "Cd") 
model_bmw <- add_variable(model_bmw, "Is") 
model_bmw <- add_variable(model_bmw, "Id") 
model_bmw <- add_variable(model_bmw, "Ls") 
model_bmw <- add_variable(model_bmw, "Ld") 
model_bmw <- add_variable(model_bmw, "WBd") 
model_bmw <- add_variable(model_bmw, "AF") 
model_bmw <- add_variable(model_bmw, "K") 
model_bmw <- add_variable(model_bmw, "YD") 
model_bmw <- add_variable(model_bmw, "WBs") 
model_bmw <- add_variable(model_bmw, "rm") 
model_bmw <- add_variable(model_bmw, "Mh") 
model_bmw <- add_variable(model_bmw, "Ms") 
model_bmw <- add_variable(model_bmw, "W") 
model_bmw <- add_variable(model_bmw, "DA") 
model_bmw <- add_variable(model_bmw, "KT")
```

You also have to add equations:
```{r}
# Add equations
model_bmw <- add_equation(model_bmw, "Cs = Cd") 
model_bmw <- add_equation(model_bmw, "Is = Id") 
model_bmw <- add_equation(model_bmw, "Ns = Nd") 
model_bmw <- add_equation(model_bmw, "Ls = Ls[-1] + Ld - Ld[-1]") 
model_bmw <- add_equation(model_bmw, "Y = Cs + Is") 
model_bmw <- add_equation(model_bmw, "WBd = Y - rl[-1] * Ld[-1] - AF") 
model_bmw <- add_equation(model_bmw, "AF = delta * K[-1]") 
model_bmw <- add_equation(model_bmw, "Ld = Ld[-1] + Id - AF") 
model_bmw <- add_equation(model_bmw, "YD = WBs + rm[-1] * Mh[-1]") 
model_bmw <- add_equation(model_bmw, "Mh = Mh[-1] + YD - Cd") 
model_bmw <- add_equation(model_bmw, "Ms = Ms[-1] + Ls - Ls[-1]") 
model_bmw <- add_equation(model_bmw, "rm = rl") 
model_bmw <- add_equation(model_bmw, "WBs = W * Ns") 
model_bmw <- add_equation(model_bmw, "Nd = Y / pr") 
model_bmw <- add_equation(model_bmw, "W = WBd / Nd") 
model_bmw <- add_equation(model_bmw, "Cd = alpha0 + alpha1 * YD + alpha2 * Mh[-1]") 
model_bmw <- add_equation(model_bmw, "K = K[-1] + Id - DA") 
model_bmw <- add_equation(model_bmw, "DA = delta * K[-1]") 
model_bmw <- add_equation(model_bmw, "KT = kappa * Y[-1]") 
model_bmw <- add_equation(model_bmw, "Id = gamma * (KT - K[-1]) + DA") 
model_bmw <- add_equation(model_bmw, "Ms = Mh", hidden = T)
```

Now you can simulate model (in this example for the next 100 periods using Gauss method):
```{r}
# Simulate model
model_bmw <- simulate_scenario(model_bmw, scenario = "baseline", max_iter = 350, periods = 100, 
                               hidden_tol = 0.1, tol = 1e-08, method = "Gauss")
```

Based on calculated simulation you can display results for desired set of variables:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_bmw, scenario = c("baseline"), from = 1, to = 50, 
                expressions = c("Cd", "YD"))
```

### Shock scenarios
With godley you can simulate how a shock will affect the base scenario. 

#### Shock 1
In this example you will implement increased autonomous consumption expenditures. \
\
First you need to create empty shock:
```{r}
# Create empty shock
shock_bmw <- create_shock()
```

Next you need to add appropriate equation:
```{r}
# Add shock equation with increase in autonomous consumption expenditures
shock_bmw <- add_shock(shock_bmw, equation = "alpha0 = 30", 
                       desc = "Increase in autonomous consumption expenditures", start = 5, end = 50)
```

Then you need to create new shock scenario:
```{r}
# Create new scenario with this shock
model_bmw <- add_scenario(model_bmw, name = "expansion1", origin = "baseline", 
                          origin_period = 100, shock = shock_bmw)
```

Now you can simulate new shock scenario:
```{r}
# Simulate shock
model_bmw <- simulate_scenario(model_bmw, scenario = "expansion1", max_iter = 350, periods = 100, 
                               hidden_tol = 0.1, tol = 1e-08, method = "Newton")
```

You can also plot results from the shock scenarios:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_bmw, scenario = c("expansion1"), from = 1, to = 50, 
                expressions = c("Cd", "YD"))
```

#### Shock 2
In this example you will implement increased propensity to save. \
\
First you need to create empty shock:
```{r}
# Create empty shock
shock_bmw <- create_shock()
```

Next you need to add appropriate equation:
```{r}
# Add shock equation with increased propensity to save
shock_bmw <- add_shock(shock_bmw, equation = "alpha1 = 0.7", 
                       desc = "Increase in the propensity to save", start = 5, end = 50)
```

Later you need to create new shock scenario:
```{r}
# Create new scenario with this shock
model_bmw <- add_scenario(model_bmw, name = "expansion2", origin = "baseline", 
                          origin_period = 100, shock = shock_bmw)
```

Now you can simulate new shock scenario:
```{r}
# Simulate shock
model_bmw <- simulate_scenario(model_bmw, scenario = "expansion2", max_iter = 350, periods = 100, 
                               hidden_tol = 0.1, tol = 1e-08, method = "Newton")
```

You can also plot results from the shock scenarios:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_bmw, scenario = c("expansion2"), from = 1, to = 50, 
                expressions = c("Cd", "YD"))
```

### References
More information about model BMW can be found in Chapter 7 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*
