---
title: "Model BMWK"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model BMWK}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Please find below guidelines how godley package can be used to create the model BMWK — Kaldorian model with private bank money described by *Wynne Godley* and *Marc Lavoie* in Chapter 7 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*

```{r setup}
library(godley)
```

### Base scenario
First you need to create empty SFC model:
```{r}
# Create empty model
model_bmwk <- create_model(name = "SFC BMWK")
```

Next you have to add variables:
```{r}
# Add variables
model_bmwk <- add_variable(model_bmwk, "rl", init = 0.025)
model_bmwk <- add_variable(model_bmwk, "alpha0", init = 20)
model_bmwk <- add_variable(model_bmwk, "alpha2", init = 0.10)
model_bmwk <- add_variable(model_bmwk, "delta", init = 0.10)
model_bmwk <- add_variable(model_bmwk, "gamma", init = 0.15)
model_bmwk <- add_variable(model_bmwk, "kappa", init = 1)
model_bmwk <- add_variable(model_bmwk, "pr", init = 1)
model_bmwk <- add_variable(model_bmwk, "Nd", init = .001)
model_bmwk <- add_variable(model_bmwk, "Ns", init = .001)
model_bmwk <- add_variable(model_bmwk, "Y", init = .001)
model_bmwk <- add_variable(model_bmwk, "alpha1w", init = .8)
model_bmwk <- add_variable(model_bmwk, "alpha1r", init = .15)
model_bmwk <- add_variable(model_bmwk, "Cs")
model_bmwk <- add_variable(model_bmwk, "Cd")
model_bmwk <- add_variable(model_bmwk, "Is")
model_bmwk <- add_variable(model_bmwk, "Id")
model_bmwk <- add_variable(model_bmwk, "Ls")
model_bmwk <- add_variable(model_bmwk, "Ld")
model_bmwk <- add_variable(model_bmwk, "WBd")
model_bmwk <- add_variable(model_bmwk, "AF")
model_bmwk <- add_variable(model_bmwk, "K")
model_bmwk <- add_variable(model_bmwk, "YD")
model_bmwk <- add_variable(model_bmwk, "WBs")
model_bmwk <- add_variable(model_bmwk, "rm")
model_bmwk <- add_variable(model_bmwk, "Mh")
model_bmwk <- add_variable(model_bmwk, "Ms")
model_bmwk <- add_variable(model_bmwk, "W")
model_bmwk <- add_variable(model_bmwk, "DA")
model_bmwk <- add_variable(model_bmwk, "KT")
```

You also have to add equations:
```{r}
# Add equations
model_bmwk <- add_equation(model_bmwk, "Cs = Cd")
model_bmwk <- add_equation(model_bmwk, "Is = Id")
model_bmwk <- add_equation(model_bmwk, "Ns = Nd")
model_bmwk <- add_equation(model_bmwk, "Ls = Ls[-1] + Ld - Ld[-1]")
model_bmwk <- add_equation(model_bmwk, "Y = Cs + Is")
model_bmwk <- add_equation(model_bmwk, "WBd = Y - rl[-1] * Ld[-1] - AF")
model_bmwk <- add_equation(model_bmwk, "AF = delta * K[-1]")
model_bmwk <- add_equation(model_bmwk, "Ld = Ld[-1] + Id - AF")
model_bmwk <- add_equation(model_bmwk, "YD = WBs + rm[-1] * Mh[-1]")
model_bmwk <- add_equation(model_bmwk, "Mh = Mh[-1] + YD - Cd")
model_bmwk <- add_equation(model_bmwk, "Ms = Ms[-1] + Ls - Ls[-1]")
model_bmwk <- add_equation(model_bmwk, "rm = rl")
model_bmwk <- add_equation(model_bmwk, "WBs = W * Ns")
model_bmwk <- add_equation(model_bmwk, "Nd = Y / pr")
model_bmwk <- add_equation(model_bmwk, "W = WBd / Nd")
model_bmwk <- add_equation(model_bmwk, "Cd = alpha0 + alpha1w * WBs + alpha1r * rm[-1] * Mh[-1] + alpha2 * Mh")
model_bmwk <- add_equation(model_bmwk, "K = K[-1] + Id - DA")
model_bmwk <- add_equation(model_bmwk, "DA = delta * K[-1]")
model_bmwk <- add_equation(model_bmwk, "KT = kappa * Y[-1]")
model_bmwk <- add_equation(model_bmwk, "Id = gamma * (KT - K[-1]) + DA")
model_bmwk <- add_equation(model_bmwk, "Ms = Mh", hidden = T)
```

Now you can simulate model (in this example for the next 100 periods using Gauss method):
```{r}
# Simulate model
model_bmwk <- simulate_scenario(model_bmwk, scenario = "baseline", max_iter = 350, periods = 100, 
                               hidden_tol = 0.1, tol = 1e-08, method = "Gauss")
```

Based on calculated simulation you can display results for desired set of variables:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_bmwk, scenario = c("baseline"), from = 1, to = 50, 
                expressions = c("Y"))
```

### Shock scenario
You can simulate now how a shock will affect the base scenario. In this example you will implement increase in the interest rate. \
\
First you need to create empty shock:
```{r}
# Create empty shock
shock_bmwk <- create_shock()
```

Next you need to add appropriate equation:
```{r}
# Add shock equation
shock_bmwk <- add_shock(shock_bmwk, equation = "rl = 0.035", 
                       desc = "Increase in the interest rate", start = 5, end = 50)
```

Then you need to create new shock scenario:
```{r}
# Create new scenario with this shock
model_bmwk <- add_scenario(model_bmwk, name = "expansion", origin = "baseline", 
                          origin_period = 100, shock = shock_bmwk)
```

Now you can simulate new shock scenario:
```{r}
# Simulate shock
model_bmwk <- simulate_scenario(model_bmwk, scenario = "expansion", max_iter = 350, periods = 100, 
                               hidden_tol = 0.1, tol = 1e-08, method = "Gauss")
```

You can also plot simulation results from the shock scenario:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_bmwk, scenario = c("expansion"), from = 1, to = 50, 
                expressions = c("Y"))
```

### References
More information about model BMWK can be found in Chapter 7 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*
