---
title: "Model DISINF"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model DISINF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Please find below guidelines how godley package can be used to create model DISINF — model with private bank money, inventories and inflation described by *Wynne Godley* and *Marc Lavoie* in Chapter 9 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*

```{r setup}
library(godley)
```

### Base scenario
First you need to create empty SFC model:
```{r}
# Create empty model
model_disinf <- create_model(name = "SFC DISINF")
```

Next you have to add variables:
```{r}
# Add variables
model_disinf <- add_variable(model_disinf, "rrc", init = 0.025)
model_disinf <- add_variable(model_disinf, "pr", init = 1)
model_disinf <- add_variable(model_disinf, "add", init = 0.02)
model_disinf <- add_variable(model_disinf, "alpha0", init = 15)
model_disinf <- add_variable(model_disinf, "alpha1", init = 0.8)
model_disinf <- add_variable(model_disinf, "alpha2", init = 0.1)
model_disinf <- add_variable(model_disinf, "beta", init = 0.9)
model_disinf <- add_variable(model_disinf, "epsilon", init = 0.8)
model_disinf <- add_variable(model_disinf, "gamma", init = 0.25)
model_disinf <- add_variable(model_disinf, "phi", init = 0.24)
model_disinf <- add_variable(model_disinf, "sigma_T", init = 0.2)
model_disinf <- add_variable(model_disinf, "Omega0", init = -1.4)
model_disinf <- add_variable(model_disinf, "Omega1", init = 1)
model_disinf <- add_variable(model_disinf, "Omega2", init = 1.2)
model_disinf <- add_variable(model_disinf, "Omega3", init = 0.3)
model_disinf <- add_variable(model_disinf, "p", init = 1)
model_disinf <- add_variable(model_disinf, "W", init = 1)
model_disinf <- add_variable(model_disinf, "UC", init = 1)
model_disinf <- add_variable(model_disinf, "s_E", init = .00001)
model_disinf <- add_variable(model_disinf, "inv_T")
model_disinf <- add_variable(model_disinf, "inv_E")
model_disinf <- add_variable(model_disinf, "inv")
model_disinf <- add_variable(model_disinf, "s")
model_disinf <- add_variable(model_disinf, "c")
model_disinf <- add_variable(model_disinf, "N")
model_disinf <- add_variable(model_disinf, "WB")
model_disinf <- add_variable(model_disinf, "INV")
model_disinf <- add_variable(model_disinf, "S")
model_disinf <- add_variable(model_disinf, "EF")
model_disinf <- add_variable(model_disinf, "Ld")
model_disinf <- add_variable(model_disinf, "Ls")
model_disinf <- add_variable(model_disinf, "Ms")
model_disinf <- add_variable(model_disinf, "rm")
model_disinf <- add_variable(model_disinf, "EFb")
model_disinf <- add_variable(model_disinf, "Mh")
model_disinf <- add_variable(model_disinf, "YD")
model_disinf <- add_variable(model_disinf, "C")
model_disinf <- add_variable(model_disinf, "omega_T")
model_disinf <- add_variable(model_disinf, "Nfe")
model_disinf <- add_variable(model_disinf, "yfe")
model_disinf <- add_variable(model_disinf, "mh")
model_disinf <- add_variable(model_disinf, "y")
model_disinf <- add_variable(model_disinf, "rl")
model_disinf <- add_variable(model_disinf, "pic")
model_disinf <- add_variable(model_disinf, "ydhs")
model_disinf <- add_variable(model_disinf, "yd")
model_disinf <- add_variable(model_disinf, "ydhs_E")
```

You also have to add equations:
```{r}
# Add equations
model_disinf <- add_equation(model_disinf, "y = s_E + inv_E - inv[-1]")
model_disinf <- add_equation(model_disinf, "inv_T = sigma_T * s_E")
model_disinf <- add_equation(model_disinf, "inv_E = inv[-1] + gamma * (inv_T - inv[-1])")
model_disinf <- add_equation(model_disinf, "inv = inv[-1] + (y - s)")
model_disinf <- add_equation(model_disinf, "s_E = beta * s[-1] + (1 - beta) * s_E[-1]")
model_disinf <- add_equation(model_disinf, "s = c")
model_disinf <- add_equation(model_disinf, "N = y / pr")
model_disinf <- add_equation(model_disinf, "WB = N * W")
model_disinf <- add_equation(model_disinf, "UC = WB / y")
model_disinf <- add_equation(model_disinf, "INV = inv * UC")
model_disinf <- add_equation(model_disinf, "S = p * s")
model_disinf <- add_equation(model_disinf, "p = (1 + phi) * (1 + rrc * sigma_T) * UC")
model_disinf <- add_equation(model_disinf, "EF = S - WB + (INV - INV[-1]) - rl * INV[-1]")
model_disinf <- add_equation(model_disinf, "Ld = INV")
model_disinf <- add_equation(model_disinf, "Ls = Ld")
model_disinf <- add_equation(model_disinf, "Ms = Ls")
model_disinf <- add_equation(model_disinf, "rm = rl - add")
model_disinf <- add_equation(model_disinf, "EFb = rl[-1] * Ls[-1] - rm[-1] * Mh[-1]")
model_disinf <- add_equation(model_disinf, "pic = (UC / UC[-1]) - 1")
model_disinf <- add_equation(model_disinf, "rl = (1 + rrc) * (1 + pic) - 1")
model_disinf <- add_equation(model_disinf, "YD = WB + EF + EFb + rm * Mh[-1]")
model_disinf <- add_equation(model_disinf, "Mh = Mh[-1] + YD - C")
model_disinf <- add_equation(model_disinf, "ydhs = c + (mh - mh[-1])")
model_disinf <- add_equation(model_disinf, "yd = YD / p")
model_disinf <- add_equation(model_disinf, "C = c * p")
model_disinf <- add_equation(model_disinf, "mh = Mh / p")
model_disinf <- add_equation(model_disinf, "c = alpha0 + alpha1 * ydhs_E + alpha2 * mh[-1]")
model_disinf <- add_equation(model_disinf, "ydhs_E = epsilon * ydhs[-1] + (1 - epsilon) * ydhs_E[-1]")
model_disinf <- add_equation(model_disinf, "omega_T = Omega0 + Omega1 * pr + Omega2 * (N / Nfe)")
model_disinf <- add_equation(model_disinf, "W = W[-1] * (1 + Omega3 * (omega_T[-1] - (W[-1]/p[-1])))")
model_disinf <- add_equation(model_disinf, "yfe = (1 + sigma_T) * s - inv[-1]")
model_disinf <- add_equation(model_disinf, "Nfe = s / pr")
```

Now you can simulate model (in this example for the next 100 periods using Gauss method):
```{r}
# Simulate model
model_disinf<- simulate_scenario(model_disinf, scenario = "baseline", max_iter = 350, periods = 100, 
                                 hidden_tol = 0.1, tol = 1e-08, method = "Gauss")
```

Based on calculated simulation you can display results for desired set of variables:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_disinf, scenario = c("baseline"), from = 1, to = 50, 
                expressions = c("p", "UC", "UCp = UC/p"))
```

```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_disinf, scenario = c("baseline"), from = 1, to = 50, 
                expressions = c("inflation = (p - dplyr::lag(p)) / dplyr::lag(p)"))
```

```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_disinf, scenario = c("baseline"), from = 1, to = 50, 
                expressions = c("ydhs_ss = alpha0 / (1 - alpha1 - alpha2*sigma_T * (UC/p))", 
                                "ydhs", "c", "s"))
```

### Shock scenario
With godley you can simulate how a shock will affect the base scenario. 

#### Shock 1
In this example you will implement an increase in the costing margins. \
\
First you need to create empty shock:
```{r}
# Create empty shock
shock_disinf <- create_shock()
```

Next you need to add appropriate equation:
```{r}
# Add shock equation
shock_disinf <- add_shock(shock_disinf, equation = "phi = 0.3", 
                          desc = "Increase in the costing margins", start = 5, end = 50)
```

Then you need to create new shock scenario:
```{r}
# Create new scenario with this shock
model_disinf <- add_scenario(model_disinf, name = "expansion1", origin = "baseline", 
                         origin_period = 100, shock = shock_disinf)
```

Now you can simulate new shock scenario:
```{r}
# Simulate shock
model_disinf <- simulate_scenario(model_disinf, scenario = "expansion1", max_iter = 350, periods = 100, 
                              hidden_tol = 0.1, tol = 1e-08, method = "Gauss")
```

You can also display simulation results on the plots:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_disinf, scenario = c("expansion1"), from = 1, to = 50, 
                expressions = c("p", "UC", "UCp = UC/p"))
```

```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_disinf, scenario = c("expansion1"), from = 1, to = 50, 
                expressions = c("inflation = (p - dplyr::lag(p)) / dplyr::lag(p)"))
```

```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_disinf, scenario = c("expansion1"), from = 1, to = 50, 
                expressions = c("ydhs_ss = alpha0 / (1 - alpha1 - alpha2*sigma_T * (UC/p))", 
                                "ydhs", "c", "s"))
```

#### Shock 2
In this example you will implement an increase in the target real wage. \
\
First you need to create empty shock:
```{r}
# Create empty shock
shock_disinf <- create_shock()
```

Next you need to add appropriate equation:
```{r}
# Add shock equation
shock_disinf <- add_shock(shock_disinf, equation = "Omega0   = -1", 
                          desc = "Increase in the target real wage", start = 5, end = 100)
```

Then you need to create new shock scenario:
```{r}
# Create new scenario with this shock
model_disinf <- add_scenario(model_disinf, name = "expansion2", origin = "baseline", 
                             origin_period = 1, shock = shock_disinf)
```

Now you can simulate new shock scenario:
```{r}
# Simulate shock
model_disinf <- simulate_scenario(model_disinf, scenario = "expansion2", max_iter = 350, periods = 100, 
                                  hidden_tol = 0.1, tol = 1e-08, method = "Gauss")
```

You can also display simulation results on the plots:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_disinf, scenario = c("expansion2"), from = 1, to = 50, 
                expressions = c("p", "UC", "UCp = UC/p"))
```

```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_disinf, scenario = c("expansion2"), from = 1, to = 50, 
                expressions = c("inflation = (p - dplyr::lag(p)) / dplyr::lag(p)"))
```

```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_disinf, scenario = c("expansion2"), from = 1, to = 50, 
                expressions = c("ydhs_ss = alpha0 / (1 - alpha1 - alpha2*sigma_T * (UC/p))", 
                                "ydhs", "c", "s"))
```

### References
More information about model DISINF can be found in Chapter 9 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*