---
title: "Model PCEX"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model PCEX}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Please find below guidelines how godley package can be used to create models PCEX — models with government money and portfolio choice with expectations described by *Wynne Godley* and *Marc Lavoie* in Chapter 4 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*

```{r setup}
library(godley)
```

### Model PCEX - random shocks
In this example expectations are subject to a random process, described in equation with a new variable "Ra" normally distributed with mean equal to 0: ```Yd_e = Yd * (1 + Ra)``` \
To create this model first you need to create empty SFC model:
```{r}
# Create empty model
model_pcex <- create_model(name = "SFC PCEX")
```

Next you have to add variables:
```{r}
# Add variables
model_pcex <- model_pcex |>
  add_variable("B_cb", desc = "") |>
  add_variable("H_s", desc = "") |>
  add_variable("B_s", desc = "") |>
  add_variable("B_h", desc = "") |>
  add_variable("H_d1", desc = "") |>
  add_variable("H_d", desc = "") |>
  add_variable("H_h", desc = "") |>
  add_variable("C", desc = "") |>
  add_variable("V", desc = "") |>
  add_variable("T_x", desc = "") |>
  add_variable("Y", desc = "Income = GDP") |>
  add_variable("Yd", desc = "Disposable income of households") |>
  add_variable("Yd_e") |>
  add_variable("B_d") |>
  add_variable("V_e") |>
  add_variable("Ra") |>
  add_variable("alpha1", init = 0.6, desc = "Propensity to consume out of income") |>
  add_variable("alpha2", init = 0.4, desc = "Propensity to consume out of wealth") |>
  add_variable("theta", init = 0.2, desc = "Tax rate") |>
  add_variable("r", init = 0.025, desc = "") |>
  add_variable("G", init = 20, desc = "Government demand") |>
  add_variable("lambda0", init = 0.635, desc = "") |>
  add_variable("lambda1", init = 0.05, desc = "") |>
  add_variable("lambda2", init = 0.01, desc = "")
```

you also have to add equations:
```{r}
# Add equations
model_pcex <- model_pcex |>
  add_equation("Y = C + G", desc = "") |>
  add_equation("Yd = Y - T_x + r[-1] * B_h[-1]") |>
  add_equation("T_x = theta * (Y + r[-1] * B_h[-1])") |>
  add_equation("V = V[-1] + (Yd - C)") |>
  add_equation("C = alpha1 * Yd_e + alpha2 * V[-1]") |>
  add_equation("B_d =  V_e * lambda0 + V_e * lambda1 * r - lambda2 * Yd_e") |>
  add_equation("H_d1 = V_e * (1 - lambda0) - V_e * lambda1 * r + lambda2 * Yd_e") |>
  add_equation("H_d = V_e - B_d") |>
  add_equation("V_e = V[-1] + (Yd_e - C)") |>
  add_equation("H_h = V - B_h") |>
  add_equation("B_h = B_d") |>
  add_equation("B_s = B_s[-1] + (G + r[-1] * B_s[-1]) - (T_x + r[-1] * B_cb[-1])") |>
  add_equation("H_s = H_s[-1] + B_cb - B_cb[-1]") |>
  add_equation("B_cb = B_s - B_h") |>
  add_equation("Yd_e = Yd * (1 + Ra)") |>
  add_equation("Ra = rnorm(1, 0, 0.05)") |>
  add_equation("H_h = H_s", hidden = T)
```

Now you can simulate model (in this example for the next 100 periods using Newton method):
```{r}
# Simulate model
model_pcex <- simulate_scenario(model_pcex, scenario = "baseline", max_iter = 350, periods = 100, 
                                hidden_tol = 0.1, tol = 1e-08, method = "Newton")
```

Based on calculated simulation you can display results for desired set of variables:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_pcex, scenario = c("baseline"), from = 1, to = 50,
                expressions = c("H_h", 'H_d'))
```

Note: In above example new pipe operator was applied. It does not work in older R versions (before R 4.1). 

### Model PCEX1 - adaptive expectations
In this example we assume that expectations about current disposable income are based on actual disposable income from the previous period. \
To apply this model first you need to create empty SFC model:
```{r}
# Create empty model
model_pcex1 <- create_model(name = "SFC PCEX1")
```

Next you have to add variables:
```{r}
# Add variables
model_pcex1 <- model_pcex1 |>
  add_variable("B_cb", desc = "") |>
  add_variable("H_s", desc = "") |>
  add_variable("B_s", desc = "") |>
  add_variable("B_h", desc = "") |>
  add_variable("H_d1", desc = "") |>
  add_variable("H_d", desc = "") |>
  add_variable("H_h", desc = "") |>
  add_variable("C", desc = "") |>
  add_variable("V", desc = "") |>
  add_variable("T_x", desc = "") |>
  add_variable("Y", desc = "Income = GDP") |>
  add_variable("Yd", desc = "Disposable income of households") |>
  add_variable("Yd_e") |>
  add_variable("B_d") |>
  add_variable("V_e") |>
  add_variable("alpha1", init = 0.6, desc = "Propensity to consume out of income") |>
  add_variable("alpha2", init = 0.4, desc = "Propensity to consume out of wealth") |>
  add_variable("theta", init = 0.2, desc = "Tax rate") |>
  add_variable("r", init = 0.025, desc = "") |>
  add_variable("G", init = 20, desc = "Government demand") |>
  add_variable("lambda0", init = 0.635, desc = "") |>
  add_variable("lambda1", init = 0.05, desc = "") |>
  add_variable("lambda2", init = 0.01, desc = "")
```

you also have to change equation about expected disposable income from previous example to ```Yd_e = Yd[-1]```:
```{r}
# Add equations
model_pcex1 <- model_pcex1 |>
  add_equation("Y = C + G", desc = "") |>
  add_equation("Yd = Y - T_x + r[-1] * B_h[-1]") |>
  add_equation("T_x = theta * (Y + r[-1] * B_h[-1])") |>
  add_equation("V = V[-1] + (Yd - C)") |>
  add_equation("C = alpha1 * Yd_e + alpha2 * V[-1]") |>
  add_equation("B_d =  V_e * lambda0 + V_e * lambda1 * r - lambda2 * Yd_e") |>
  add_equation("H_d1 = V_e * (1 - lambda0) - V_e * lambda1 * r + lambda2 * Yd_e") |>
  add_equation("H_d = V_e - B_d") |>
  add_equation("V_e = V[-1] + (Yd_e - C)") |>
  add_equation("H_h = V - B_h") |>
  add_equation("B_h = B_d") |>
  add_equation("B_s = B_s[-1] + (G + r[-1] * B_s[-1]) - (T_x + r[-1] * B_cb[-1])") |>
  add_equation("H_s = H_s[-1] + B_cb - B_cb[-1]") |>
  add_equation("B_cb = B_s - B_h") |>
  add_equation("Yd_e = Yd[-1]") |>
  add_equation("H_h = H_s", hidden = T)
```

Now you can simulate model (in this example for the next 100 periods using Newton method):
```{r}
# Simulate model
model_pcex1 <- simulate_scenario(model_pcex1, scenario = "baseline", max_iter = 350, 
                                 periods = 100, hidden_tol = 0.1, tol = 1e-08, method = "Newton")
```

Based on calculated simulation you can display results for desired set of variables:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_pcex1, scenario = c("baseline"), from = 1, to = 50,
                expressions = c("Y"))
```

Note: In above example new pipe operator was applied. It does not work in older R versions (before R 4.1).

#### Shock - increase in the propensity to consume out of disposable income

Now you can simulate what will happen with economy if the propensity to consume out of disposable income increase to 0.7.\
First you need to create empty shock object and add new equation for ```alpha1```:
```{r}
# Create empty shock
shock_pcex1 <- create_shock()

# Add shock equation with increased government expenditures
shock_pcex1 <- add_shock(shock_pcex1,
                         equation = "alpha1 = 0.7",
                         desc = "Increase in the propensity to consume out of disposable income", start = 5, end = 50
)

# Create new scenario with this shock
model_pcex1 <- add_scenario(model_pcex1,
                            name = "expansion", origin = "baseline",
                            origin_period = 100, shock = shock_pcex1
)
```

Now you can simulate shock scenario:
```{r}
# Simulate shock
model_pcex1 <- simulate_scenario(model_pcex1,
                                 scenario = "expansion", max_iter = 350, periods = 100,
                                 hidden_tol = 0.1, tol = 1e-08, method = "Newton"
)
```

Based on results you can plot national income, evolution of consumption, expected disposable income and lagged wealth:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_pcex1, scenario = c("expansion"), from = 1, to = 50,
                expressions = c("Y"))
```

```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_pcex1, scenario = c("expansion"), from = 1, to = 50,
                expressions = c("Yd_e", "C", "V"))
```

### Model PCEX2 - where the propensity to consume reacts negatively to higher interest rates
In this example we assume that expectations about current disposable income are based on actual disposable income from the previous period like in previous example but the propensity to consume reacts negatively to higher interest rates. To apply this model we need to add two new variables ```iota``` and ```alpha10``` to variables used in PCEX1 and add one additional equation ```alpha1 = alpha10 - iota * r[-1]```.\
To apply this model first you need to create empty SFC model:
```{r}
# Create empty model
model_pcex2 <- create_model(name = "SFC PCEX2")
```

Next you have to add variables:
```{r}
# Add variables
model_pcex2 <- model_pcex2 |>
  add_variable("B_cb", desc = "") |>
  add_variable("H_s", desc = "") |>
  add_variable("B_s", desc = "") |>
  add_variable("B_h", desc = "") |>
  add_variable("H_d1", desc = "") |>
  add_variable("H_d", desc = "") |>
  add_variable("H_h", desc = "") |>
  add_variable("C", desc = "") |>
  add_variable("V", desc = "") |>
  add_variable("T_x", desc = "") |>
  add_variable("Y", desc = "Income = GDP") |>
  add_variable("Yd", desc = "Disposable income of households") |>
  add_variable("Yd_e") |>
  add_variable("B_d") |>
  add_variable("V_e") |>
  add_variable("iota", init = 4) |>
  add_variable("alpha10", init = 0.7) |>
  add_variable("alpha1", init = 0.6, desc = "Propensity to consume out of income") |>
  add_variable("alpha2", init = 0.4, desc = "Propensity to consume out of wealth") |>
  add_variable("theta", init = 0.2, desc = "Tax rate") |>
  add_variable("r", init = 0.025, desc = "") |>
  add_variable("G", init = 20, desc = "Government demand") |>
  add_variable("lambda0", init = 0.635, desc = "") |>
  add_variable("lambda1", init = 0.05, desc = "") |>
  add_variable("lambda2", init = 0.01, desc = "")
```

You also have to add equations: 
```{r}
# Add equations
model_pcex2 <- model_pcex2 |>
  add_equation("Y = C + G", desc = "") |>
  add_equation("Yd = Y - T_x + r[-1] * B_h[-1]") |>
  add_equation("T_x = theta * (Y + r[-1] * B_h[-1])") |>
  add_equation("V = V[-1] + (Yd - C)") |>
  add_equation("C = alpha1 * Yd_e + alpha2 * V[-1]") |>
  add_equation("B_d =  V_e * lambda0 + V_e * lambda1 * r - lambda2 * Yd_e") |>
  add_equation("H_d1 = V_e * (1 - lambda0) - V_e * lambda1 * r + lambda2 * Yd_e") |>
  add_equation("H_d = V_e - B_d") |>
  add_equation("V_e = V[-1] + (Yd_e - C)") |>
  add_equation("H_h = V - B_h") |>
  add_equation("B_h = B_d") |>
  add_equation("B_s = B_s[-1] + (G + r[-1] * B_s[-1]) - (T_x + r[-1] * B_cb[-1])") |>
  add_equation("H_s = H_s[-1] + B_cb - B_cb[-1]") |>
  add_equation("B_cb = B_s - B_h") |>
  add_equation("Yd_e = Yd[-1]") |>
  add_equation("alpha1 = alpha10 - iota * r[-1]") |>
  add_equation("H_h = H_s", hidden = T)
```

Now you can simulate model (in this example for the next 100 periods using Newton method):
```{r}
# Simulate model
model_pcex2 <- simulate_scenario(model_pcex2, scenario = "baseline", max_iter = 350, 
                                 periods = 100, hidden_tol = 0.1, tol = 1e-08, method = "Newton")
```

Based on calculated simulation you can display results for desired set of variables:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_pcex2, scenario = c("baseline"), from = 1, to = 50,
                expressions = c("Y"))
```

Note: In above example new pipe operator was applied. It does not work in older R versions (before R 4.1). 

#### Shock - increase in the rate of interest on bills

Now you can simulate what will happen with the economy if the rate of interest on bills increase by 100 points.\
First you need to create empty shock object and add new equation for ```r```:
```{r}
# Create empty shock
shock_pcex2 <- create_shock()

# Add shock equation with increased government expenditures
shock_pcex2 <- add_shock(shock_pcex2,
                         equation = "r = 0.035",
                         desc = "Increase in the rate of interest on bills", start = 5, end = 50
)

# Create new scenario with this shock
model_pcex2 <- add_scenario(model_pcex2,
                            name = "expansion", origin = "baseline",
                            origin_period = 100, shock = shock_pcex2
)
```

Now you can simulate shock scenario:
```{r}
# Simulate shock
model_pcex2 <- simulate_scenario(model_pcex2,
                                 scenario = "expansion", max_iter = 350, periods = 100,
                                 hidden_tol = 0.1, tol = 1e-08, method = "Newton"
)
```

Based on results you can plot national income, evolution of consumption, expected disposable income and lagged wealth:
```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_pcex2, scenario = c("expansion"), from = 1, to = 50,
                expressions = c("Y"))
```

```{r, out.width="100%"}
# Plot results
plot_simulation(model = model_pcex2, scenario = c("expansion"), from = 1, to = 50,
                expressions = c("Yd", "C", "V"))
```

### References
More information about models PC with expectations can be found in Chapter 4 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*