---
title: "Model Advanced OPEN"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model Advanced OPEN}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include = FALSE}
library(dplyr)
library(purrr)
library(plotly)
library(tidyverse)

plotly_style <- function(fig, title = NULL) {
  fig %>%
    plotly::layout(
      title = list(
        text = title %||% "",
        y = 0.95, x = 0.5,
        xanchor = "center", yanchor = "top"
      ),
      margin = list(t = 60, b = 70, l = 60, r = 120),
      hovermode = "spikers",
      spikedistance = 1000,
      xaxis = list(
        title = "time"
      ),
      yaxis = list(
        title = ""
      ),
      showlegend = TRUE,
      legend = list(
        xanchor = "left",
        x = 1.02,
        y = 0.95
      ),
      font = list(
        family = "Arial",
        size = 13
      )
    )
}

# Plot results
do_plotly <- function(m, scenario, variables, t0 = 1, start = NULL, end = NULL,
                      title = NULL, y_title = NULL,
                      mode = "lines") {
  
  m1 <- m[[scenario]][["result"]]
  
  start <- if (is.null(start)) t0 else start
  end   <- if (is.null(end)) max(m1[["time"]]) else end
  
  df_long <- m1 %>%
    mutate(tab_uk = x_uk - im_uk,
           tab_us = x_us - im_us,
           gab_uk = -psbr_uk,
           gab_us = -psbr_us,
           dres_uk = b_cb_ukus_d - lag(b_cb_ukus_d),
           db_cb_uk = b_cb_ukuk_d - lag(b_cb_ukuk_d),
           dh_uk = h_uk_s - lag(h_uk_s),
           dh_us = h_us_s - lag(h_us_s),
           by_uk = b_uk_s / y_uk,
           by_us = b_us_s / y_us,
           bukus_p = (b_ukus_d/xr_us)/v_uk,
           bukus_d = b_ukus_d / v_uk) %>%
    filter(time >= start & time <= end) %>%
    pivot_longer(cols = -time) %>%
    filter(name %in% variables)

  fig <- plotly::plot_ly()

  for (v in unique(df_long$name)) {
    dfv <- dplyr::filter(df_long, name == v)

    fig <- plotly::add_trace(
      fig,
      data = dfv,
      x = ~time,
      y = ~value,
      name = v,
      type = "scatter",
      mode = mode,
      line = list(width = 2, dash = "solid"),
      hovertemplate = paste(
        "<b>", v, "</b>",
        "<br>time=%{x}",
        "<br>value=%{y}<extra></extra>"
      )
    )
  }

  fig %>%
    plotly::layout(
      yaxis = list(title = y_title %||% "")
    ) %>%
    plotly_style(title)
}
```

This guide explains how to use the `godley` package to build more advanced open-economy SFC models, as described by *Wynne Godley* and *Marc Lavoie* in Chapter 12 of *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth*.  

We present **four variants** of the same open-economy model. All of them share the same core structure; what differs across them is the **closure**.

More specifically, they differ along the following dimensions:

* whether the exchange rate is endogenous or fixed,
* whether foreign reserves adjust or are kept constant,
* whether the adjustment happens through monetary policy (endogenous interest rate),
* whether the adjustment happens through fiscal policy (endogenous government expenditure).

```{r setup}
library(godley)
library(purrr)
library(plotly)
```

## 1. Model OPENFLEX
Following *Godley* and *Lavoie*, we first consider a **flexible exchange rate closure**. In this model, the central bank, faced with a step-up in imports, refuses to intervene in foreign exchange markets so that it does not lose any of its foreign exchange reserves. The exchange rate is endogenous, while the amount of foreign reserves, measured in dollars, becomes a constant. US Bills supplied to foreign bond holders are a residual. We refer to this description of the economy as the `OPENFLEX` model.

### Base scenario
We start with a baseline version of the model and build from there.

First, create an empty SFC (Stock-Flow Consistent) model:
```{r}
# Create empty model
model_open <- create_model(name = "SFC Advanced OPENFLEX")
```

Next come the model variables. We declare all variables used in the model. Additionally, if you wish, you can assign initial values — for some variables, for all of them, or for none.
```{r}
# Add variables
model_open <- model_open |>
  add_variable("alpha1_uk", init = 0.75) |>
  add_variable("alpha1_us", init = 0.75) |>
  add_variable("alpha2_uk", init = 0.13333) |>
  add_variable("alpha2_us", init = 0.13333) |>
  add_variable("eps0", init = -2.1) |>
  add_variable("eps1", init = 0.7) |>
  add_variable("eps2", init = 1) |>
  add_variable("lambda10", init = 0.7) |>
  add_variable("lambda11", init = 5) |>
  add_variable("lambda12", init = 5) |>
  add_variable("lambda20", init = 0.25) |>
  add_variable("lambda21", init = 5) |>
  add_variable("lambda22", init = 5) |>
  add_variable("lambda40", init = 0.7) |>
  add_variable("lambda41", init = 5) |>
  add_variable("lambda42", init = 5) |>
  add_variable("lambda50", init = 0.25) |>
  add_variable("lambda51", init = 5) |>
  add_variable("lambda52", init = 5) |>
  add_variable("mu0", init = -2.1) |>
  add_variable("mu1", init = 0.7) |>
  add_variable("mu2", init = 1) |>
  add_variable("nu0m", init = -0.00001) |>
  add_variable("nu0x", init = -0.00001) |>
  add_variable("nu1m", init = 0.7) |>
  add_variable("nu1x", init = 0.5) |>
  add_variable("phi_uk", init = 0.2381) |>
  add_variable("phi_us", init = 0.2381) |>
  add_variable("theta_uk", init = 0.2) |>
  add_variable("theta_us", init = 0.2) |>
  add_variable("b_cb_ukus_s", init = 0.02031) |>
  add_variable("dxre_us", init = 0) |>
  add_variable("g_k_uk", init = 16) |>
  add_variable("g_k_us", init = 16) |>
  add_variable("or_uk", init = 7) |>
  add_variable("pg_us", init = 1) |>
  add_variable("pr_uk", init = 1.3333) |>
  add_variable("pr_us", init = 1.3333) |>
  add_variable("r_uk", init = 0.03) |>
  add_variable("r_us", init = 0.03) |>
  add_variable("w_uk", init = 1) |>
  add_variable("w_us", init = 1) |>
  add_variable("b_cb_ukuk_d", init = 0.27984) |>
  add_variable("b_cb_ukuk_s", init = 0.27984) |>
  add_variable("b_cb_ukuk_sa", init = 0.27984) |>
  add_variable("b_cb_ukus_d", init = 0.0203) |>
  add_variable("b_cb_usus_d", init = 0.29843) |>
  add_variable("b_cb_usus_s", init = 0.29843) |>
  add_variable("b_uk_s", init = 138.94) |>
  add_variable("b_ukuk_d", init = 102.18) |>
  add_variable("b_ukuk_s", init = 102.18) |>
  add_variable("b_ukus_d", init = 36.493) |>
  add_variable("b_ukus_s", init = 36.504) |>
  add_variable("b_us_s", init = 139.02) |>
  add_variable("b_usuk_d", init = 36.497) |>
  add_variable("b_usuk_s", init = 36.487) |>
  add_variable("b_usus_d", init = 102.19) |>
  add_variable("b_usus_s", init = 102.19) |>
  add_variable("h_uk_d", init = 7.2987) |>
  add_variable("h_uk_s", init = 7.2987) |>
  add_variable("h_us_d", init = 7.2995) |>
  add_variable("h_us_s", init = 7.2995) |>
  add_variable("or_us", init = 7) |>
  add_variable("v_k_uk", init = 152.62) |>
  add_variable("v_k_us", init = 152.63) |>
  add_variable("v_uk", init = 145.97) |>
  add_variable("v_us", init = 145.99001) |>
  add_variable("c_k_uk", init = 81.393) |>
  add_variable("c_k_us", init = 81.401) |>
  add_variable("cab_uk", init = 0) |>
  add_variable("cab_us", init = 0) |>
  add_variable("cons_uk", init = 77.851) |>
  add_variable("cons_us", init = 77.86) |>
  add_variable("ds_k_uk", init = 97.393) |>
  add_variable("ds_k_us", init = 97.401) |>
  add_variable("ds_uk", init = 93.154) |>
  add_variable("ds_us", init = 93.164) |>
  add_variable("dxre_uk", init = 0) |>
  add_variable("f_cb_uk", init = 0.00869) |>
  add_variable("f_cb_us", init = 0.00895) |>
  add_variable("g_uk", init = 15.304) |>
  add_variable("g_us", init = 15.304) |>
  add_variable("im_k_uk", init = 11.928) |>
  add_variable("im_k_us", init = 11.926) |>
  add_variable("im_uk", init = 11.407) |>
  add_variable("im_us", init = 11.409) |>
  add_variable("kabp_uk", init = 0.00002) |>
  add_variable("kabp_us", init = -0.00002) |>
  add_variable("n_uk", init = 73.046) |>
  add_variable("n_us", init = 73.054) |>
  add_variable("pds_uk", init = 0.95648) |>
  add_variable("pds_us", init = 0.95649) |>
  add_variable("pg_uk", init = 0.99971) |>
  add_variable("pm_uk", init = 0.95628) |>
  add_variable("pm_us", init = 0.95661) |>
  add_variable("ps_uk", init = 0.95646) |>
  add_variable("ps_us", init = 0.9565) |>
  add_variable("px_uk", init = 0.95634) |>
  add_variable("px_us", init = 0.95656) |>
  add_variable("py_uk", init = 0.95648) |>
  add_variable("py_us", init = 0.95649) |>
  add_variable("s_k_uk", init = 109.32) |>
  add_variable("s_k_us", init = 109.33) |>
  add_variable("s_uk", init = 104.56) |>
  add_variable("s_us", init = 104.57) |>
  add_variable("t_uk", init = 19.463) |>
  add_variable("t_us", init = 19.465) |>
  add_variable("x_k_uk", init = 11.926) |>
  add_variable("x_k_us", init = 11.928) |>
  add_variable("x_uk", init = 11.406) |>
  add_variable("x_us", init = 11.41) |>
  add_variable("xr_uk", init = 1.0003) |>
  add_variable("xr_us", init = 0.99971) |>
  add_variable("xre_uk", init = 1.0003) |>
  add_variable("xre_us", init = 0.99971) |>
  add_variable("y_k_uk", init = 97.392) |>
  add_variable("y_k_us", init = 97.403) |>
  add_variable("y_uk", init = 93.154) |>
  add_variable("y_us", init = 93.164) |>
  add_variable("yd_uk", init = 77.851) |>
  add_variable("yd_us", init = 77.86) |>
  add_variable("ydhs_k_uk", init = 81.394) |>
  add_variable("ydhs_k_us", init = 81.402) |>
  add_variable("ydhse_k_uk", init = 81.394) |>
  add_variable("ydhse_k_us", init = 81.402) |>
  add_variable("yd_hs_uk") |>
  add_variable("yd_hs_us") |>
  add_variable("kab_uk") |>
  add_variable("kab_us") |>
  add_variable("psbr_uk") |>
  add_variable("psbr_us") |>
  add_variable("nafa_uk") |>
  add_variable("nafa_us") |>
  add_variable("nwcb_uk")
```

Then, add the behavioural equations and accounting identities that tie the system together. This block also includes the closure-specific equations for the fixed exchange-rate setup.
```{r}
# Add equations
model_open <- model_open |>
  add_equation("yd_uk = (y_uk + r_uk[-1]*b_ukuk_d[-1] + xr_us*r_us[-1]*b_ukus_s[-1])*(1 - theta_uk) + (xr_us - xr_us[-1])*b_ukus_s[-1]", desc = "12.1 : Disposable income in UK") |>
  add_equation("yd_hs_uk = yd_uk + (xr_us - xr_us[-1])*b_ukus_s[-1]", desc = "12.2 : Haig-Simons disposable income in UK") |>
  add_equation("v_uk = v_uk[-1] + yd_uk - cons_uk", desc = "12.3 : Wealth accumulation in UK") |>
  add_equation("yd_us = (y_us + r_us[-1]*b_usus_d[-1] + xr_uk*r_uk[-1]*b_usuk_s[-1])*(1 - theta_us) + (xr_uk - xr_uk[-1])*b_usuk_s[-1]", desc = "12.4 : Disposable income in US") |>
  add_equation("yd_hs_us = yd_us + d(xr_uk)*b_usuk_s[-1]", desc = "12.5 : Haig-Simons disposable income in US") |>
  add_equation("v_us = v_us[-1] + yd_us - cons_us", desc = "12.6 : Wealth accumulation in US") |>
  add_equation("t_uk = theta_uk*(y_uk + r_uk[-1]*b_ukuk_d[-1] + xr_us*r_us[-1]*b_ukus_s[-1])", desc = "12.7 : Taxes in UK") |>
  add_equation("t_us = theta_us*(y_us + r_us[-1]*b_usus_d[-1] + xr_uk*r_uk[-1]*b_usuk_s[-1])", desc = "12.8 : Taxes in US") |>
  add_equation("f_cb_uk = r_uk[-1]*b_cb_ukuk_d[-1] + r_us[-1]*b_cb_ukus_s[-1]*xr_us", desc = "12.11 : Profits of Central Bank in UK") |>
  add_equation("f_cb_us = r_us[-1]*b_cb_usus_d[-1]", desc = "12.12 : Profits of Central Bank in US") |>
  add_equation("b_uk_s = b_uk_s[-1] + g_uk + r_uk[-1]*b_uk_s[-1] - t_uk - f_cb_uk", desc = "12.13 : Government budget constraint - UK") |>
  add_equation("b_us_s = b_us_s[-1] + g_us + r_us[-1]*b_us_s[-1] - t_us - f_cb_us", desc = "12.14 : Government budget constraint - US") |>
  add_equation("cab_uk = x_uk - im_uk + xr_us*r_us[-1]*b_ukus_s[-1] - r_uk[-1]*b_usuk_s[-1] + r_us[-1]*b_cb_ukus_s[-1]*xr_us", desc = "12.15 : Current account balance - UK") |>
  add_equation("kab_uk = kabp_uk - (xr_us*(b_cb_ukus_s - b_cb_ukus_s[-1]) + pg_uk*(or_uk - or_uk[-1]))", desc = "12.16 : Capital account balance in UK") |>
  add_equation("cab_us = x_us - im_us + xr_uk*r_uk[-1]*b_usuk_s[-1] - r_us[-1]*b_ukus_s[-1] - r_us[-1]*b_cb_ukus_s[-1]", desc = "12.17 : Current account balance in US") |>
  add_equation("kab_us = kabp_us + (b_cb_ukus_s - b_cb_ukus_s[-1]) - pg_us*(or_us - or_us[-1])", desc = "12.18 : Capital account balance in US") |>
  add_equation("kabp_uk = -(b_ukus_s - b_ukus_s[-1])*xr_us + (b_usuk_s - b_usuk_s[-1])", desc = "12.19 : Capital account balance in UK, net of official transactions") |>
  add_equation("kabp_us = -(b_usuk_s - b_usuk_s[-1])*xr_uk + (b_ukus_s - b_ukus_s[-1])", desc = "12.20 : Capital account balance in US, net of official transactions") |>
  
  # TRADE
  add_equation("pm_uk = exp(nu0m + nu1m*log(py_us) + (1 - nu1m)*log(py_uk) - nu1m*log(xr_uk))", desc = "12.21 : Import prices in UK") |>
  add_equation("px_uk = exp(nu0x + nu1x*log(py_us) + (1 - nu1x)*log(py_uk) - nu1x*log(xr_uk))", desc = "12.22 : Export prices in UK") |>
  add_equation("px_us = pm_uk*xr_uk", desc = "12.23 : Export prices in US") |>
  add_equation("pm_us = px_uk*xr_uk", desc = "12.24 : Import prices in US") |>
  add_equation("x_k_uk = exp(eps0 - eps1*log(pm_us/py_us) + eps2*log(y_k_us))", desc = "12.25 : Real exports from UK - depends on current relative price") |>
  add_equation("im_k_uk = exp(mu0 - mu1*log(pm_uk[-1]/py_uk[-1]) + mu2*log(y_k_uk))", desc = "12.26 : Real imports of UK") |>
  add_equation("x_k_us = im_k_uk", desc = "12.27 : Real exports from US") |>
  add_equation("im_k_us = x_k_uk", desc = "12.28 : Real imports of US") |>
  add_equation("x_uk = x_k_uk*px_uk", desc = "12.29 : Exports of UK") |>
  add_equation("x_us = x_k_us*px_us", desc = "12.30 : Exports of US") |>
  add_equation("im_uk = im_k_uk*pm_uk", desc = "12.31 : Imports of UK") |>
  add_equation("im_us = im_k_us*pm_us", desc = "12.32 : Imports of US") |>
  
  # INCOME AND EXPENDITURE
  add_equation("v_k_uk = v_uk/pds_uk", desc = "12.33 : Real wealth in UK") |>
  add_equation("v_k_us = v_us/pds_us", desc = "12.34 : Real wealth in US") |>
  add_equation("ydhs_k_uk = yd_uk/pds_uk - v_k_uk[-1]*(pds_uk - pds_uk[-1])/pds_uk", desc = "12.35 : Real Haig-Simons disposable income in UK") |>
  add_equation("ydhs_k_us = yd_us/pds_us - v_k_us[-1]*(pds_us - pds_us[-1])/pds_us", desc = "12.36 : Real Haig-Simons disposable income in US") |>
  add_equation("c_k_uk = alpha1_uk*ydhse_k_uk + alpha2_uk*v_k_uk[-1]", desc = "12.37 : Real consumption in UK") |>
  add_equation("c_k_us = alpha1_us*ydhse_k_us + alpha2_us*v_k_us[-1]", desc = "12.38 : Real consumption in US") |>
  add_equation("ydhse_k_uk = (ydhs_k_uk + ydhs_k_uk[-1])/2", desc = "12.39 : Expected real Haig-Simons disposable income in UK") |>
  add_equation("ydhse_k_us = (ydhs_k_us + ydhs_k_us[-1])/2", desc = "12.40 : Expected real Haig-Simons disposable income in US") |>
  add_equation("s_k_uk = c_k_uk + g_k_uk + x_k_uk", desc = "12.41 : Real sales in UK") |>
  add_equation("s_k_us = c_k_us + g_k_us + x_k_us", desc = "12.42 : Real sales in US") |>
  add_equation("s_uk = s_k_uk*ps_uk", desc = "12.43 : Value of sales in UK") |>
  add_equation("s_us = s_k_us*ps_us", desc = "12.44 : Value of sales in US") |>
  add_equation("ps_uk = (1 + phi_uk)*(w_uk*n_uk + im_uk)/s_k_uk", desc = "12.45 : Price of sales in UK") |>
  add_equation("ps_us = (1 + phi_us)*(w_us*n_us + im_us)/s_k_us", desc = "12.46 : Price of sales in US") |>
  add_equation("pds_uk = (s_uk - x_uk)/(s_k_uk - x_k_uk)", desc = "12.47 : Price of domestic sales in UK") |>
  add_equation("pds_us = (s_us - x_us)/(s_k_us - x_k_us)", desc = "12.48 : Price of domestic sales in US") |>
  add_equation("ds_uk = s_uk - x_uk", desc = "12.49 : Domestic sales in UK") |>
  add_equation("ds_us = s_us - x_us", desc = "12.50 : Domestic sales in US") |>
  add_equation("ds_k_uk = c_k_uk + g_k_uk", desc = "12.51 : Real domestic sales in UK") |>
  add_equation("ds_k_us = c_k_us + g_k_us", desc = "12.52 : Real domestic sales in US") |>
  add_equation("y_uk = s_uk - im_uk", desc = "12.53 : Value of output in UK") |>
  add_equation("y_us = s_us - im_us", desc = "12.54 : Value of output in US") |>
  add_equation("y_k_uk = s_k_uk - im_k_uk", desc = "12.55 : Value of real output in UK") |>
  add_equation("y_k_us = s_k_us - im_k_us", desc = "12.56 : Value of real output in US") |>
  add_equation("py_uk = y_uk/y_k_uk", desc = "12.57 : Price of output in UK") |>
  add_equation("py_us = y_us/y_k_us", desc = "12.58 : Price of output in US") |>
  add_equation("cons_uk = c_k_uk*pds_uk", desc = "12.59 : Consumption in UK") |>
  add_equation("cons_us = c_k_us*pds_us", desc = "12.60 : Consumption in US") |>
  add_equation("g_uk = g_k_uk*pds_uk", desc = "12.61 : Government expenditure in UK") |>
  add_equation("g_us = g_k_us*pds_us", desc = "12.62 : Government expenditure in US") |>
  add_equation("n_uk = y_k_uk/pr_uk", desc = "12.65 : Employment in UK") |>
  add_equation("n_us = y_k_us/pr_us", desc = "12.66 : Employment in US") |>
  
  # ASSET DEMANDS
  add_equation("b_ukuk_d = v_uk*(lambda10 + lambda11*r_uk - lambda12*(r_us + dxre_us))", desc = "12.67 : Demand for UK bills in UK") |>
  add_equation("b_ukus_d = v_uk*(lambda20 - lambda21*r_uk + lambda22*(r_us + dxre_us))", desc = "12.68F : Demand for US bills in UK") |>
  add_equation("h_uk_d = v_uk - b_ukuk_d - b_ukus_d", desc = "12.69 : Demand for money in UK") |>
  add_equation("b_usus_d = v_us*(lambda40 + lambda41*r_us - lambda42*(r_uk + dxre_uk))", desc = "12.70 : Demand for US	bills in US") |>
  add_equation("b_usuk_d = v_us*(lambda50 - lambda51*r_us + lambda52*(r_uk + dxre_uk))", desc = "12.71 : Demand for UK bills in US") |>
  add_equation("h_us_d = v_us - b_usus_d - b_usuk_d", desc = "12.72 : Demand for money in US") |>
  
  # ASSET SUPPLIES
  add_equation("h_us_s = h_us_d", desc = "12.77 : Suply of cash in US") |>
  add_equation("b_usus_s = b_usus_d", desc = "12.78 : Supply of US bills to Country") |>
  add_equation("b_cb_usus_s = b_cb_usus_d", desc = "12.79 : Supply of US bills to US Central bank") |>
  add_equation("h_uk_s = h_uk_d", desc = "12.80 : Suply of cash in UK") |>
  add_equation("b_ukuk_s = b_ukuk_d", desc = "12.81 : Bills issued by US acquired by US") |>
  add_equation("b_cb_ukuk_s = b_cb_ukuk_d", desc = "12.82 : Supply of UK bills to UK Central bank") |>
  add_equation("b_cb_usus_d = b_cb_usus_d[-1] + (h_us_s - h_us_s[-1]) - (or_us - or_us[-1])*pg_us", desc = "12.83 : Balance sheet of US Central bank - expressed as changes") |>
  add_equation("b_cb_ukuk_d = b_cb_ukuk_d[-1] + (h_uk_s - h_uk_s[-1]) - (b_cb_ukus_s - b_cb_ukus_s[-1])*xr_us - (or_uk - or_uk[-1])*pg_uk", desc = "12.84 : Balance sheet of UK Central bank") |>
  add_equation("pg_uk = pg_us/xr_uk", desc = "12.85 : Price of gold is equal in the two countries") |>
  add_equation("xr_us = 1/xr_uk", desc = "12.86 : US exchange rate") |>
  add_equation("b_usuk_s = b_usuk_d*xr_us", desc = "12.87 : Equilibrium condition for bills issued by UK acquired by US") |>
  add_equation("b_cb_ukus_d = b_cb_ukus_s*xr_us", desc = "12.88 : Equilibrium condition for bills issued by US acquired by UK Central bank") |>
  add_equation("psbr_uk = g_uk + r_uk[-1]*b_uk_s[-1] - t_uk - f_cb_uk", desc = "Government deficit in the UK") |>
  add_equation("psbr_us = g_us + r_us[-1]*b_us_s[-1] - t_us - f_cb_us", desc = "Government deficit in the US") |>
  add_equation("nafa_uk = psbr_uk + cab_uk", desc = "Net accumulation of financial assets in the UK") |>
  add_equation("nafa_us = psbr_us + cab_us", desc = "Net accumulation of financial assets in the US") |>
  add_equation("b_cb_ukuk_sa = b_uk_s - b_ukuk_s - b_usuk_s", desc = "") |>
  add_equation("nwcb_uk = -h_uk_d + b_cb_ukuk_d + b_cb_ukus_d * xr_us + or_uk * pg_uk", desc = "Net wealth of CBUK") |>
  
  # Model OPEN FLEX Closure
  add_equation("xr_uk = b_ukus_s/b_ukus_d", desc = "12.89FL : UK Exchange rate") |>
  add_equation("b_ukus_s = b_us_s - b_usus_s - b_cb_usus_d - b_cb_ukus_s", desc = "12.90G : Supply of UK bills to US") |>
  
  # Hidden equation
  add_equation("b_cb_ukuk_s = b_cb_ukuk_sa", hidden = TRUE)
```

Now run the baseline simulation to make sure it converges and settles into a stable path.
```{r}
# Simulate model
model_open <- simulate_scenario(model_open, scenario = "baseline",
                                max_iter = 350, periods = 100, tol = 1e-15,
                                hidden_tol = 0.1, method = "Broyden")
```

### Shock scenario
With `godley` package we can simulate how shocks affect the economy (specifically, how they impact the base scenario).

#### Shock 1

This scenario explores how the model behaves when confronted with a step fall in the UK exports.

To start, define the shock by specifying which parameter changes and over what time window it applies.
```{r}
# Create empty shock and add shock equation
shock_open <- create_shock() |>
  add_shock(variable = "eps0", value = -2.2, start = 5, end = 100, desc = "")
```

Next, add the shock to the model by creating a dedicated scenario.
```{r}
# Create new scenario with this shock
model_open <- model_open |>
  add_scenario(name = "eps0_shock", origin = "baseline", shock = shock_open)
```

Simulate the scenario to capture its effects.
```{r}
# Simulate shock
model_open <- simulate_scenario(model_open, scenario = "eps0_shock", periods = 100,
                                max_iter = 350, tol = 1e-10, method = "Broyden",
                                hidden_tol = 0.1, rhtol = TRUE)
```

Finally, plot the simulation outcomes:
```{r, out.width="100%"}
# Figure 12.5A Effect of a decrease in the UK propensity to export, within a flexible exchange rate regime, on various UK variables: current account balance, trade balance and government budget deficit
do_plotly(m = model_open, scenario = "eps0_shock",
        variables = c("cab_uk", "tab_uk", "psbr_uk"),
        title = "Scenario eps0_shock", y_title = "")
```

```{r, out.width="100%"}
# Figure 12.5D Effect of a decrease in the UK propensity to export, within a flexible exchange rate regime, on the UK and US real GDP
do_plotly(m = model_open, scenario = "eps0_shock", variables = c("y_k_uk", "y_k_us"),
        title = "Scenario eps0_shock", y_title = "")
```

#### Shock 2

In the second example, we ask what changes when there is an increase in the desire to hold US treasury bills.

As before, start by creating an empty shock, define paths for `lambda20` and `lambda40`, and simulate the scenario.
```{r}
# Create empty shock and add shock equation
shock_open <- create_shock() |>
  add_shock(variable = "lambda20", value = 0.30, start = 5, end = 100, desc = "") |>
  add_shock(variable = "lambda40", value = 0.75, start = 5, end = 100, desc = "")

model_open <- model_open |>
  add_scenario(name = "lambda20-40_shock", origin = "baseline", shock = shock_open)

# Simulate shock
model_open <- simulate_scenario(model_open, scenario = "lambda20-40_shock", periods = 100,
                                max_iter = 350, tol = 1e-10, method = "Broyden",
                                hidden_tol = 0.1, rhtol = TRUE)
```

You can then plot the results.
```{r, out.width="100%"}
# Plot results
# Figure 12.7C Effect on UK and US real GDP of an increase in the desire to hold US Treasury bills, within a flexible exchange rate regime
do_plotly(m = model_open, scenario = "lambda20-40_shock", variables = c("y_k_us", "y_k_uk"),
        title = "Scenario lambda20, lambda40", y_title = "")
```

```{r, out.width="100%"}
# Figure 12.7D Effect of an increase in the desire to hold US Treasury bills, within a flexible exchange rate regime, on various US variables: net accumulation of financial assets, government budget deficit, current account balance and trade balance
do_plotly(m = model_open, scenario = "lambda20-40_shock", variables = c("psbr_us", "cab_us", "nafa_us", "tab_us"), title = "Scenario lambda20, lambda40", y_title = "")
```

## 2. Model OPENFIX
We next turn to **a fixed exchange rate closure, with endogenous foreign reserves**. 

In this case, the exchange rate is an exogenous variable, shifting the adjustment to foreign reserves. The second equation of the closure describes the evolution of UK foreign reserves. Alternatively expressed, it describes the purchases of US Treasury bills which the UK central bank must make in order to prevent its exchange rate from floating up or down. As to the first equation of the closure, it follows directly from the definition of the exchange rate and states that, for an exogenously given exchange rate, the level of the supply of foreign assets ( $B^{\$}_{£s}$ ) must equal the demand for foreign assets multiplied by the exchange rate ( $B^{\$}_{£d} \cdot xr^{£}$ ).

### Base scenario
All of the discussed versions of the Advanced OPEN model differ from each other in various places. For clarity, instead of reworking the previous model, we simply start over with a new one.

So, first create a new model.
```{r}
# Create empty model
model_open <- create_model(name = "SFC Advanced OPENFIX")
```

The next step is to define its variables.
```{r}
# Add variables
model_open <- model_open |>
  add_variable("alpha1_uk", init = 0.75) |>
  add_variable("alpha1_us", init = 0.75) |>
  add_variable("alpha2_uk", init = 0.13333) |>
  add_variable("alpha2_us", init = 0.13333) |>
  add_variable("eps0", init = -2.1) |>
  add_variable("eps1", init = 0.7) |>
  add_variable("eps2", init = 1) |>
  add_variable("lambda10", init = 0.7) |>
  add_variable("lambda11", init = 5) |>
  add_variable("lambda12", init = 5) |>
  add_variable("lambda20", init = 0.25) |>
  add_variable("lambda21", init = 5) |>
  add_variable("lambda22", init = 5) |>
  add_variable("lambda40", init = 0.7) |>
  add_variable("lambda41", init = 5) |>
  add_variable("lambda42", init = 5) |>
  add_variable("lambda50", init = 0.25) |>
  add_variable("lambda51", init = 5) |>
  add_variable("lambda52", init = 5) |>
  add_variable("mu0", init = -2.1) |>
  add_variable("mu1", init = 0.7) |>
  add_variable("mu2", init = 1) |>
  add_variable("nu0m", init = -0.00001) |>
  add_variable("nu0x", init = -0.00001) |>
  add_variable("nu1m", init = 0.7) |>
  add_variable("nu1x", init = 0.5) |>
  add_variable("phi_uk", init = 0.2381) |>
  add_variable("phi_us", init = 0.2381) |>
  add_variable("theta_uk", init = 0.2) |>
  add_variable("theta_us", init = 0.2) |>
  add_variable("b_cb_ukus_s", init = 0.02031) |>
  add_variable("dxre_us", init = 0) |>
  add_variable("g_k_uk", init = 16) |>
  add_variable("g_k_us", init = 16) |>
  add_variable("or_uk", init = 7) |>
  add_variable("pg_us", init = 1) |>
  add_variable("pr_uk", init = 1.3333) |>
  add_variable("pr_us", init = 1.3333) |>
  add_variable("r_uk", init = 0.03) |>
  add_variable("r_us", init = 0.03) |>
  add_variable("w_uk", init = 1) |>
  add_variable("w_us", init = 1) |>
  add_variable("b_cb_ukuk_d", init = 0.27984) |>
  add_variable("b_cb_ukuk_s", init = 0.27984) |>
  add_variable("b_cb_ukuk_sa", init = 0.27984) |>
  add_variable("b_cb_ukus_d", init = 0.0203) |>
  add_variable("b_cb_usus_d", init = 0.29843) |>
  add_variable("b_cb_usus_s", init = 0.29843) |>
  add_variable("b_uk_s", init = 138.94) |>
  add_variable("b_ukuk_d", init = 102.18) |>
  add_variable("b_ukuk_s", init = 102.18) |>
  add_variable("b_ukus_d", init = 36.493) |>
  add_variable("b_ukus_s", init = 36.504) |>
  add_variable("b_us_s", init = 139.02) |>
  add_variable("b_usuk_d", init = 36.497) |>
  add_variable("b_usuk_s", init = 36.487) |>
  add_variable("b_usus_d", init = 102.19) |>
  add_variable("b_usus_s", init = 102.19) |>
  add_variable("h_uk_d", init = 7.2987) |>
  add_variable("h_uk_s", init = 7.2987) |>
  add_variable("h_us_d", init = 7.2995) |>
  add_variable("h_us_s", init = 7.2995) |>
  add_variable("or_us", init = 7) |>
  add_variable("v_k_uk", init = 152.62) |>
  add_variable("v_k_us", init = 152.63) |>
  add_variable("v_uk", init = 145.97) |>
  add_variable("v_us", init = 145.99001) |>
  add_variable("c_k_uk", init = 81.393) |>
  add_variable("c_k_us", init = 81.401) |>
  add_variable("cab_uk", init = 0) |>
  add_variable("cab_us", init = 0) |>
  add_variable("cons_uk", init = 77.851) |>
  add_variable("cons_us", init = 77.86) |>
  add_variable("ds_k_uk", init = 97.393) |>
  add_variable("ds_k_us", init = 97.401) |>
  add_variable("ds_uk", init = 93.154) |>
  add_variable("ds_us", init = 93.164) |>
  add_variable("dxre_uk", init = 0) |>
  add_variable("f_cb_uk", init = 0.00869) |>
  add_variable("f_cb_us", init = 0.00895) |>
  add_variable("g_uk", init = 15.304) |>
  add_variable("g_us", init = 15.304) |>
  add_variable("im_k_uk", init = 11.928) |>
  add_variable("im_k_us", init = 11.926) |>
  add_variable("im_uk", init = 11.407) |>
  add_variable("im_us", init = 11.409) |>
  add_variable("kabp_uk", init = 0.00002) |>
  add_variable("kabp_us", init = -0.00002) |>
  add_variable("n_uk", init = 73.046) |>
  add_variable("n_us", init = 73.054) |>
  add_variable("pds_uk", init = 0.95648) |>
  add_variable("pds_us", init = 0.95649) |>
  add_variable("pg_uk", init = 0.99971) |>
  add_variable("pm_uk", init = 0.95628) |>
  add_variable("pm_us", init = 0.95661) |>
  add_variable("ps_uk", init = 0.95646) |>
  add_variable("ps_us", init = 0.9565) |>
  add_variable("px_uk", init = 0.95634) |>
  add_variable("px_us", init = 0.95656) |>
  add_variable("py_uk", init = 0.95648) |>
  add_variable("py_us", init = 0.95649) |>
  add_variable("s_k_uk", init = 109.32) |>
  add_variable("s_k_us", init = 109.33) |>
  add_variable("s_uk", init = 104.56) |>
  add_variable("s_us", init = 104.57) |>
  add_variable("t_uk", init = 19.463) |>
  add_variable("t_us", init = 19.465) |>
  add_variable("x_k_uk", init = 11.926) |>
  add_variable("x_k_us", init = 11.928) |>
  add_variable("x_uk", init = 11.406) |>
  add_variable("x_us", init = 11.41) |>
  add_variable("xr_uk", init = 1.0003) |>
  add_variable("xr_us", init = 0.99971) |>
  add_variable("xre_uk", init = 1.0003) |>
  add_variable("xre_us", init = 0.99971) |>
  add_variable("y_k_uk", init = 97.392) |>
  add_variable("y_k_us", init = 97.403) |>
  add_variable("y_uk", init = 93.154) |>
  add_variable("y_us", init = 93.164) |>
  add_variable("yd_uk", init = 77.851) |>
  add_variable("yd_us", init = 77.86) |>
  add_variable("ydhs_k_uk", init = 81.394) |>
  add_variable("ydhs_k_us", init = 81.402) |>
  add_variable("ydhse_k_uk", init = 81.394) |>
  add_variable("ydhse_k_us", init = 81.402) |>
  add_variable("yd_hs_uk") |>
  add_variable("yd_hs_us") |>
  add_variable("kab_uk") |>
  add_variable("kab_us") |>
  add_variable("psbr_uk") |>
  add_variable("psbr_us") |>
  add_variable("nafa_uk") |>
  add_variable("nafa_us") |>
  add_variable("nwcb_uk")
```

Then, add the equations that define the relationships between variables.
```{r}
# Add equations
model_open <- model_open |>
  add_equation("yd_uk = (y_uk + r_uk[-1]*b_ukuk_d[-1] + xr_us*r_us[-1]*b_ukus_s[-1])*(1 - theta_uk) + (xr_us - xr_us[-1])*b_ukus_s[-1]", desc = "12.1 : Disposable income in UK") |>
  add_equation("yd_hs_uk = yd_uk + (xr_us - xr_us[-1])*b_ukus_s[-1]", desc = "12.2 : Haig-Simons disposable income in UK") |>
  add_equation("v_uk = v_uk[-1] + yd_uk - cons_uk", desc = "12.3 : Wealth accumulation in UK") |>
  add_equation("yd_us = (y_us + r_us[-1]*b_usus_d[-1] + xr_uk*r_uk[-1]*b_usuk_s[-1])*(1 - theta_us) + (xr_uk - xr_uk[-1])*b_usuk_s[-1]", desc = "12.4 : Disposable income in US") |>
  add_equation("yd_hs_us = yd_us + d(xr_uk)*b_usuk_s[-1]", desc = "12.5 : Haig-Simons disposable income in US") |>
  add_equation("v_us = v_us[-1] + yd_us - cons_us", desc = "12.6 : Wealth accumulation in US") |>
  add_equation("t_uk = theta_uk*(y_uk + r_uk[-1]*b_ukuk_d[-1] + xr_us*r_us[-1]*b_ukus_s[-1])", desc = "12.7 : Taxes in UK") |>
  add_equation("t_us = theta_us*(y_us + r_us[-1]*b_usus_d[-1] + xr_uk*r_uk[-1]*b_usuk_s[-1])", desc = "12.8 : Taxes in US") |>
  add_equation("f_cb_uk = r_uk[-1]*b_cb_ukuk_d[-1] + r_us[-1]*b_cb_ukus_s[-1]*xr_us", desc = "12.11 : Profits of Central Bank in UK") |>
  add_equation("f_cb_us = r_us[-1]*b_cb_usus_d[-1]", desc = "12.12 : Profits of Central Bank in US") |>
  add_equation("b_uk_s = b_uk_s[-1] + g_uk + r_uk[-1]*b_uk_s[-1] - t_uk - f_cb_uk", desc = "12.13 : Government budget constraint - UK") |>
  add_equation("b_us_s = b_us_s[-1] + g_us + r_us[-1]*b_us_s[-1] - t_us - f_cb_us", desc = "12.14 : Government budget constraint - US") |>
  add_equation("cab_uk = x_uk - im_uk + xr_us*r_us[-1]*b_ukus_s[-1] - r_uk[-1]*b_usuk_s[-1] + r_us[-1]*b_cb_ukus_s[-1]*xr_us", desc = "12.15 : Current account balance") |>
  add_equation("kab_uk = kabp_uk - (xr_us*(b_cb_ukus_s - b_cb_ukus_s[-1]) + pg_uk*(or_uk - or_uk[-1]))", desc = "12.16 : Capital account balance in UK") |>
  add_equation("cab_us = x_us - im_us + xr_uk*r_uk[-1]*b_usuk_s[-1] - r_us[-1]*b_ukus_s[-1] - r_us[-1]*b_cb_ukus_s[-1]", desc = "12.17 : Current account balance in US") |>
  add_equation("kab_us = kabp_us + (b_cb_ukus_s - b_cb_ukus_s[-1]) - pg_us*(or_us - or_us[-1])", desc = "12.18 : Capital account balance in US") |>
  add_equation("kabp_uk = -(b_ukus_s - b_ukus_s[-1])*xr_us + (b_usuk_s - b_usuk_s[-1])", desc = "12.19 : Capital account balance in UK, net of official transactions") |>
  add_equation("kabp_us = -(b_usuk_s - b_usuk_s[-1])*xr_uk + (b_ukus_s - b_ukus_s[-1])", desc = "12.20 : Capital account balance in US, net of official transactions") |>
  
  # TRADE
  add_equation("pm_uk = exp(nu0m + nu1m*log(py_us) + (1 - nu1m)*log(py_uk) - nu1m*log(xr_uk))", desc = "12.21 : Import prices in UK") |>
  add_equation("px_uk = exp(nu0x + nu1x*log(py_us) + (1 - nu1x)*log(py_uk) - nu1x*log(xr_uk))", desc = "12.22 : Export prices in UK") |>
  add_equation("px_us = pm_uk*xr_uk", desc = "12.23 : Export prices in US") |>
  add_equation("pm_us = px_uk*xr_uk", desc = "12.24 : Import prices in US") |>
  add_equation("x_k_uk = exp(eps0 - eps1*log(pm_us/py_us) + eps2*log(y_k_us))", desc = "12.25 : Real exports from UK - depends on current relative price") |>
  add_equation("im_k_uk = exp(mu0 - mu1*log(pm_uk[-1]/py_uk[-1]) + mu2*log(y_k_uk))", desc = "12.26 : Real imports of UK") |>
  add_equation("x_k_us = im_k_uk", desc = "12.27 : Real exports from US") |>
  add_equation("im_k_us = x_k_uk", desc = "12.28 : Real imports of US") |>
  add_equation("x_uk = x_k_uk*px_uk", desc = "12.29 : Exports of UK") |>
  add_equation("x_us = x_k_us*px_us", desc = "12.30 : Exports of US") |>
  add_equation("im_uk = im_k_uk*pm_uk", desc = "12.31 : Imports of UK") |>
  add_equation("im_us = im_k_us*pm_us", desc = "12.32 : Imports of US") |>
  
  # INCOME AND EXPENDITURE
  add_equation("v_k_uk = v_uk/pds_uk", desc = "12.33 : Real wealth in UK") |>
  add_equation("v_k_us = v_us/pds_us", desc = "12.34 : Real wealth in US") |>
  add_equation("ydhs_k_uk = yd_uk/pds_uk - v_k_uk[-1]*(pds_uk - pds_uk[-1])/pds_uk", desc = "12.35 : Real Haig-Simons disposable income in UK") |>
  add_equation("ydhs_k_us = yd_us/pds_us - v_k_us[-1]*(pds_us - pds_us[-1])/pds_us", desc = "12.36 : Real Haig-Simons disposable income in US") |>
  add_equation("c_k_uk = alpha1_uk*ydhse_k_uk + alpha2_uk*v_k_uk[-1]", desc = "12.37 : Real consumption in UK") |>
  add_equation("c_k_us = alpha1_us*ydhse_k_us + alpha2_us*v_k_us[-1]", desc = "12.38 : Real consumption in US") |>
  add_equation("ydhse_k_uk = (ydhs_k_uk + ydhs_k_uk[-1])/2", desc = "12.39 : Expected real Haig-Simons disposable income in UK") |>
  add_equation("ydhse_k_us = (ydhs_k_us + ydhs_k_us[-1])/2", desc = "12.40 : Expected real Haig-Simons disposable income in US") |>
  add_equation("s_k_uk = c_k_uk + g_k_uk + x_k_uk", desc = "12.41 : Real sales in UK") |>
  add_equation("s_k_us = c_k_us + g_k_us + x_k_us", desc = "12.42 : Real sales in US") |>
  add_equation("s_uk = s_k_uk*ps_uk", desc = "12.43 : Value of sales in UK") |>
  add_equation("s_us = s_k_us*ps_us", desc = "12.44 : Value of sales in US") |>
  add_equation("ps_uk = (1 + phi_uk)*(w_uk*n_uk + im_uk)/s_k_uk", desc = "12.45 : Price of sales in UK") |>
  add_equation("ps_us = (1 + phi_us)*(w_us*n_us + im_us)/s_k_us", desc = "12.46 : Price of sales in US") |>
  add_equation("pds_uk = (s_uk - x_uk)/(s_k_uk - x_k_uk)", desc = "12.47 : Price of domestic sales in UK") |>
  add_equation("pds_us = (s_us - x_us)/(s_k_us - x_k_us)", desc = "12.48 : Price of domestic sales in US") |>
  add_equation("ds_uk = s_uk - x_uk", desc = "12.49 : Domestic sales in UK") |>
  add_equation("ds_us = s_us - x_us", desc = "12.50 : Domestic sales in US") |>
  add_equation("ds_k_uk = c_k_uk + g_k_uk", desc = "12.51 : Real domestic sales in UK") |>
  add_equation("ds_k_us = c_k_us + g_k_us", desc = "12.52 : Real domestic sales in US") |>
  add_equation("y_uk = s_uk - im_uk", desc = "12.53 : Value of output in UK") |>
  add_equation("y_us = s_us - im_us", desc = "12.54 : Value of output in US") |>
  add_equation("y_k_uk = s_k_uk - im_k_uk", desc = "12.55: Value of real output in UK") |>
  add_equation("y_k_us = s_k_us - im_k_us", desc = "12.56 : Value of real output in US") |>
  add_equation("py_uk = y_uk/y_k_uk", desc = "12.57 : Price of output in UK") |>
  add_equation("py_us = y_us/y_k_us", desc = "12.58 : Price of output in US") |>
  add_equation("cons_uk = c_k_uk*pds_uk", desc = "12.59 : Consumption in UK") |>
  add_equation("cons_us = c_k_us*pds_us", desc = "12.60 : Consumption in US") |>
  add_equation("g_uk = g_k_uk*pds_uk", desc = "12.61 : Government expenditure in UK") |>
  add_equation("g_us = g_k_us*pds_us", desc = "12.62 : Government expenditure in US") |>
  add_equation("n_uk = y_k_uk/pr_uk", desc = "12.65 : Employment in UK") |>
  add_equation("n_us = y_k_us/pr_us", desc = "12.66 : Employment in US") |>
  
  # ASSET DEMANDS
  add_equation("b_ukuk_d = v_uk*(lambda10 + lambda11*r_uk - lambda12*(r_us + dxre_us))", desc = "12.67 : Demand for UK bills in UK") |>
  add_equation("b_ukus_d = v_uk*(lambda20 - lambda21*r_uk + lambda22*(r_us + dxre_us))", desc = "12.68F: Demand for US bills in UK") |>
  add_equation("h_uk_d = v_uk - b_ukuk_d - b_ukus_d", desc = "12.69 : Demand for money in UK") |>
  add_equation("b_usus_d = v_us*(lambda40 + lambda41*r_us - lambda42*(r_uk + dxre_uk))", desc = "12.70 : Demand for US	bills in US") |>
  add_equation("b_usuk_d = v_us*(lambda50 - lambda51*r_us + lambda52*(r_uk + dxre_uk))", desc = "12.71 : Demand for UK bills in US") |>
  add_equation("h_us_d = v_us - b_usus_d - b_usuk_d", desc = "12.72 : Demand for money in US") |>
  
  # ASSET SUPPLIES
  add_equation("h_us_s = h_us_d", desc = "12.77 : Suply of cash in US") |>
  add_equation("b_usus_s = b_usus_d", desc = "12.78 : Supply of US bills to Country") |>
  add_equation("b_cb_usus_s = b_cb_usus_d", desc = "12.79 : Supply of US bills to US Central bank") |>
  add_equation("h_uk_s = h_uk_d", desc = "12.80 : Suply of cash in UK") |>
  add_equation("b_ukuk_s = b_ukuk_d", desc = "12.81 : Bills issued by US acquired by US") |>
  add_equation("b_cb_ukuk_s = b_cb_ukuk_d", desc = "12.82 : Supply of UK bills to UK Central bank") |>
  add_equation("b_cb_usus_d = b_cb_usus_d[-1] + (h_us_s - h_us_s[-1]) - (or_us - or_us[-1])*pg_us", desc = "12.83 : Balance sheet of US Central bank - expressed as changes") |>

  add_equation("b_cb_ukuk_d = b_cb_ukuk_d[-1] + (h_uk_s - h_uk_s[-1]) - (b_cb_ukus_s - b_cb_ukus_s[-1])*xr_us - (or_uk - or_uk[-1])*pg_uk", desc = "12.84 : Balance sheet of UK Central bank") |>
  add_equation("pg_uk = pg_us/xr_uk", desc = "12.85 : Price of gold is equal in the two countries") |>
  add_equation("xr_us = 1/xr_uk", desc = "12.86 : US exchange rate") |>
  add_equation("b_usuk_s = b_usuk_d*xr_us", desc = "12.87 : Equilibrium condition for bills issued by UK acquired by US") |>
  add_equation("b_cb_ukus_d = b_cb_ukus_s*xr_us", desc = "12.88 : Equilibrium condition for bills issued by US acquired by UK Central bank") |>
  add_equation("psbr_uk = g_uk + r_uk[-1]*b_uk_s[-1] - t_uk - f_cb_uk", desc = "Government deficit in the UK") |>
  add_equation("psbr_us = g_us + r_us[-1]*b_us_s[-1] - t_us - f_cb_us", desc = "Government deficit in the US") |>
  add_equation("nafa_uk = psbr_uk + cab_uk", desc = "Net accumulation of financial assets in the UK") |>
  add_equation("nafa_us = psbr_us + cab_us", desc = "Net accumulation of financial assets in the US") |>
  add_equation("b_cb_ukuk_sa = b_uk_s - b_ukuk_s - b_usuk_s", desc = "") |>
  add_equation("nwcb_uk = -h_uk_d + b_cb_ukuk_d + b_cb_ukus_d * xr_us + or_uk * pg_uk", desc = "Net wealth of CBUK") |>
  
  # Model OPEN FIX Closure
  add_equation("b_ukus_s = xr_uk*b_ukus_d", desc = "12.89R : Demand of UK Bills in US") |>
  add_equation("b_cb_ukus_s = b_us_s - b_usus_s - b_cb_usus_d - b_ukus_s", desc = "12.90R : Supply of UK bills to US") |>
  
  # Hidden equation
  add_equation("b_cb_ukuk_s = b_cb_ukuk_sa", hidden = TRUE)
```

Simulate the baseline scenario (in this example, over 100 periods using the Broyden method)...
```{r}
# Simulate model
model_open <- simulate_scenario(model_open, scenario = "baseline",
                                  max_iter = 350, periods = 100, tol = 1e-15,
                                  hidden_tol = 0.1, method = "Broyden")
```

... and move on to counterfactual simulations.

### Shock scenario
Here, we look at what happens when the US propensity to import increases.

Begin by defining the shock and its time window.
```{r}
# Create empty shock and add shock equation
shock_open <- create_shock() |>
  add_shock(variable = "eps0", value = -2, start = 5, end = 100, desc = "")
```

Add the shock to the model as a new scenario.
```{r}
# Create new scenario with this shock
model_open <- model_open |>
  add_scenario(name = "eps0_shock", origin = "baseline", shock = shock_open)
```

Run the simulation to see the effect of the shock.
```{r}
# Simulate shock
model_open <- simulate_scenario(model_open, scenario = "eps0_shock", periods = 100,
                                  max_iter = 350, tol = 1e-10, method = "Broyden",
                                  hidden_tol = 0.1, rhtol = TRUE)
```

And plot the results to show how the economy adjusts over time.
```{r, out.width="100%"}
# Figure 12.1A Effect of an increase in the US propensity to import on UK variables, within a fixed exchange rate regime with endogenous foreign reserves: net accumulation of financial assets, current account balance, trade balance and government budget balance
do_plotly(m = model_open, scenario = "eps0_shock",
        variables = c("nafa_uk", "cab_uk", "gab_uk", "tab_uk"),
        t0 = 1, end = 50, title = "Scenario eps0_shock", y_title = "")
```

```{r, out.width="100%"}
# Figure 12.1B Effect of an increase in the US propensity to import, within a fixed exchange rate regime with endogenous foreign reserves, on the UK current account balance and elements of the balance sheet of the Bank of England (the UK central bank): change in foreign reserves, stock of money, holdings of domestic Treasury bills
do_plotly(m = model_open, scenario = "eps0_shock",
        variables = c("cab_uk", "dres_uk", "db_cb_uk", "dh_uk"),
        t0 = 1, end = 50, title = "Scenario eps0_shock", y_title = "")
```

```{r, out.width="100%"}
# Figure 12.1C Effect of an increase in the US propensity to import on the US debt to GDP ratio and on the UK debt to income ratio, within a fixed exchange rate regime with endogenous foreign reserves
do_plotly(m = model_open, scenario = "eps0_shock",
        variables = c("by_uk", "by_us"),
        t0 = 1, end = 50, title = "Scenario eps0_shock", y_title = "")
```

## 3. Model OPENFIXR
In the third scenario, we consider **an alternative fixed exchange rate closure: the deficit country lets its interest rate become endogenous**.

An alternative way of fending off devaluation, in theory, is to raise the interest rate. To show
how this would have to happen, we start from the main fixed exchange rate closure, but impose upon it that the central bank declines to use its foreign reserves, as if the country were on a flexible exchange rate. Hence, the rate of interest on UK bills $r^{£}$, which was previously assumed to be constant, now needs to be made endogenous and is described in equation 3 of the model closure. The remaining two equations that make up the closure describe: the demand for UK bills in the US and the supply of UK bills to the US.

### Base scenario
Start by creating another model:
```{r}
# Create empty model
model_open <- create_model(name = "SFC Advanced OPENFIXR")
```

Next, declare all variables it will use. If you wish, you can also set initial values here.
```{r}
# Add variables
model_open <- model_open |>
  add_variable("alpha1_uk", init = 0.75) |>
  add_variable("alpha1_us", init = 0.75) |>
  add_variable("alpha2_uk", init = 0.13333) |>
  add_variable("alpha2_us", init = 0.13333) |>
  add_variable("eps0", init = -2.1) |>
  add_variable("eps1", init = 0.7) |>
  add_variable("eps2", init = 1) |>
  add_variable("lambda10", init = 0.7) |>
  add_variable("lambda11", init = 5) |>
  add_variable("lambda12", init = 5) |>
  add_variable("lambda20", init = 0.25) |>
  add_variable("lambda21", init = 5) |>
  add_variable("lambda22", init = 5) |>
  add_variable("lambda40", init = 0.7) |>
  add_variable("lambda41", init = 5) |>
  add_variable("lambda42", init = 5) |>
  add_variable("lambda50", init = 0.25) |>
  add_variable("lambda51", init = 5) |>
  add_variable("lambda52", init = 5) |>
  add_variable("mu0", init = -2.1) |>
  add_variable("mu1", init = 0.7) |>
  add_variable("mu2", init = 1) |>
  add_variable("nu0m", init = -0.00001) |>
  add_variable("nu0x", init = -0.00001) |>
  add_variable("nu1m", init = 0.7) |>
  add_variable("nu1x", init = 0.5) |>
  add_variable("phi_uk", init = 0.2381) |>
  add_variable("phi_us", init = 0.2381) |>
  add_variable("theta_uk", init = 0.2) |>
  add_variable("theta_us", init = 0.2) |>
  add_variable("b_cb_ukus_s", init = 0.02031) |>
  add_variable("dxre_us", init = 0) |>
  add_variable("g_k_uk", init = 16) |>
  add_variable("g_k_us", init = 16) |>
  add_variable("or_uk", init = 7) |>
  add_variable("pg_us", init = 1) |>
  add_variable("pr_uk", init = 1.3333) |>
  add_variable("pr_us", init = 1.3333) |>
  add_variable("r_uk", init = 0.03) |>
  add_variable("r_us", init = 0.03) |>
  add_variable("w_uk", init = 1) |>
  add_variable("w_us", init = 1) |>
  add_variable("b_cb_ukuk_d", init = 0.27984) |>
  add_variable("b_cb_ukuk_s", init = 0.27984) |>
  add_variable("b_cb_ukuk_sa", init = 0.27984) |>
  add_variable("b_cb_ukus_d", init = 0.0203) |>
  add_variable("b_cb_usus_d", init = 0.29843) |>
  add_variable("b_cb_usus_s", init = 0.29843) |>
  add_variable("b_uk_s", init = 138.94) |>
  add_variable("b_ukuk_d", init = 102.18) |>
  add_variable("b_ukuk_s", init = 102.18) |>
  add_variable("b_ukus_d", init = 36.493) |>
  add_variable("b_ukus_s", init = 36.504) |>
  add_variable("b_us_s", init = 139.02) |>
  add_variable("b_usuk_d", init = 36.497) |>
  add_variable("b_usuk_s", init = 36.487) |>
  add_variable("b_usus_d", init = 102.19) |>
  add_variable("b_usus_s", init = 102.19) |>
  add_variable("h_uk_d", init = 7.2987) |>
  add_variable("h_uk_s", init = 7.2987) |>
  add_variable("h_us_d", init = 7.2995) |>
  add_variable("h_us_s", init = 7.2995) |>
  add_variable("or_us", init = 7) |>
  add_variable("v_k_uk", init = 152.62) |>
  add_variable("v_k_us", init = 152.63) |>
  add_variable("v_uk", init = 145.97) |>
  add_variable("v_us", init = 145.99001) |>
  add_variable("c_k_uk", init = 81.393) |>
  add_variable("c_k_us", init = 81.401) |>
  add_variable("cab_uk", init = 0) |>
  add_variable("cab_us", init = 0) |>
  add_variable("cons_uk", init = 77.851) |>
  add_variable("cons_us", init = 77.86) |>
  add_variable("ds_k_uk", init = 97.393) |>
  add_variable("ds_k_us", init = 97.401) |>
  add_variable("ds_uk", init = 93.154) |>
  add_variable("ds_us", init = 93.164) |>
  add_variable("dxre_uk", init = 0) |>
  add_variable("f_cb_uk", init = 0.00869) |>
  add_variable("f_cb_us", init = 0.00895) |>
  add_variable("g_uk", init = 15.304) |>
  add_variable("g_us", init = 15.304) |>
  add_variable("im_k_uk", init = 11.928) |>
  add_variable("im_k_us", init = 11.926) |>
  add_variable("im_uk", init = 11.407) |>
  add_variable("im_us", init = 11.409) |>
  add_variable("kabp_uk", init = 0.00002) |>
  add_variable("kabp_us", init = -0.00002) |>
  add_variable("n_uk", init = 73.046) |>
  add_variable("n_us", init = 73.054) |>
  add_variable("pds_uk", init = 0.95648) |>
  add_variable("pds_us", init = 0.95649) |>
  add_variable("pg_uk", init = 0.99971) |>
  add_variable("pm_uk", init = 0.95628) |>
  add_variable("pm_us", init = 0.95661) |>
  add_variable("ps_uk", init = 0.95646) |>
  add_variable("ps_us", init = 0.9565) |>
  add_variable("px_uk", init = 0.95634) |>
  add_variable("px_us", init = 0.95656) |>
  add_variable("py_uk", init = 0.95648) |>
  add_variable("py_us", init = 0.95649) |>
  add_variable("s_k_uk", init = 109.32) |>
  add_variable("s_k_us", init = 109.33) |>
  add_variable("s_uk", init = 104.56) |>
  add_variable("s_us", init = 104.57) |>
  add_variable("t_uk", init = 19.463) |>
  add_variable("t_us", init = 19.465) |>
  add_variable("x_k_uk", init = 11.926) |>
  add_variable("x_k_us", init = 11.928) |>
  add_variable("x_uk", init = 11.406) |>
  add_variable("x_us", init = 11.41) |>
  add_variable("xr_uk", init = 1.0003) |>
  add_variable("xr_us", init = 0.99971) |>
  add_variable("xre_uk", init = 1.0003) |>
  add_variable("xre_us", init = 0.99971) |>
  add_variable("y_k_uk", init = 97.392) |>
  add_variable("y_k_us", init = 97.403) |>
  add_variable("y_uk", init = 93.154) |>
  add_variable("y_us", init = 93.164) |>
  add_variable("yd_uk", init = 77.851) |>
  add_variable("yd_us", init = 77.86) |>
  add_variable("ydhs_k_uk", init = 81.394) |>
  add_variable("ydhs_k_us", init = 81.402) |>
  add_variable("ydhse_k_uk", init = 81.394) |>
  add_variable("ydhse_k_us", init = 81.402) |>
  add_variable("yd_hs_uk") |>
  add_variable("yd_hs_us") |>
  add_variable("kab_uk") |>
  add_variable("kab_us") |>
  add_variable("psbr_uk") |>
  add_variable("psbr_us") |>
  add_variable("nafa_uk") |>
  add_variable("nafa_us") |>
  add_variable("nwcb_uk")
```

Then, add the equations that define how the variables relate to each other.
```{r}
# Add equations
model_open <- model_open |>
  add_equation("yd_uk = (y_uk + r_uk[-1]*b_ukuk_d[-1] + xr_us*r_us[-1]*b_ukus_s[-1])*(1 - theta_uk) + (xr_us - xr_us[-1])*b_ukus_s[-1]", desc = "12.1 : Disposable income in UK") |>
  add_equation("yd_hs_uk = yd_uk + (xr_us - xr_us[-1])*b_ukus_s[-1]", desc = "12.2 : Haig-Simons disposable income in UK") |>
  add_equation("v_uk = v_uk[-1] + yd_uk - cons_uk", desc = "12.3 : Wealth accumulation in UK") |>
  add_equation("yd_us = (y_us + r_us[-1]*b_usus_d[-1] + xr_uk*r_uk[-1]*b_usuk_s[-1])*(1 - theta_us) + (xr_uk - xr_uk[-1])*b_usuk_s[-1]", desc = "12.4 : Disposable income in US") |>
  add_equation("yd_hs_us = yd_us + d(xr_uk)*b_usuk_s[-1]", desc = "12.5 : Haig-Simons disposable income in US") |>
  add_equation("v_us = v_us[-1] + yd_us - cons_us", desc = "12.6 : Wealth accumulation in US") |>
  add_equation("t_uk = theta_uk*(y_uk + r_uk[-1]*b_ukuk_d[-1] + xr_us*r_us[-1]*b_ukus_s[-1])", desc = "12.7 : Taxes in UK") |>
  add_equation("t_us = theta_us*(y_us + r_us[-1]*b_usus_d[-1] + xr_uk*r_uk[-1]*b_usuk_s[-1])", desc = "12.8 : Taxes in US") |>
  add_equation("f_cb_uk = r_uk[-1]*b_cb_ukuk_d[-1] + r_us[-1]*b_cb_ukus_s[-1]*xr_us", desc = "12.11 : Profits of Central Bank in UK") |>
  add_equation("f_cb_us = r_us[-1]*b_cb_usus_d[-1]", desc = "12.12 : Profits of Central Bank in US") |>
  add_equation("b_uk_s = b_uk_s[-1] + g_uk + r_uk[-1]*b_uk_s[-1] - t_uk - f_cb_uk", desc = "12.13 : Government budget constraint - UK") |>
  add_equation("b_us_s = b_us_s[-1] + g_us + r_us[-1]*b_us_s[-1] - t_us - f_cb_us", desc = "12.14 : Government budget constraint - US") |>
  add_equation("cab_uk = x_uk - im_uk + xr_us*r_us[-1]*b_ukus_s[-1] - r_uk[-1]*b_usuk_s[-1] + r_us[-1]*b_cb_ukus_s[-1]*xr_us", desc = "12.15 : Current account balance - UK") |>
  add_equation("kab_uk = kabp_uk - (xr_us*(b_cb_ukus_s - b_cb_ukus_s[-1]) + pg_uk*(or_uk - or_uk[-1]))", desc = "12.16 : Capital account balance in UK") |>
  add_equation("cab_us = x_us - im_us + xr_uk*r_uk[-1]*b_usuk_s[-1] - r_us[-1]*b_ukus_s[-1] - r_us[-1]*b_cb_ukus_s[-1]", desc = "12.17 : Current account balance in US") |>
  add_equation("kab_us = kabp_us + (b_cb_ukus_s - b_cb_ukus_s[-1]) - pg_us*(or_us - or_us[-1])", desc = "12.18 : Capital account balance in US") |>
  add_equation("kabp_uk = -(b_ukus_s - b_ukus_s[-1])*xr_us + (b_usuk_s - b_usuk_s[-1])", desc = "12.19 : Capital account balance in UK, net of official transactions") |>
  add_equation("kabp_us = -(b_usuk_s - b_usuk_s[-1])*xr_uk + (b_ukus_s - b_ukus_s[-1])", desc = "12.20 : Capital account balance in US, net of official transactions") |>
  
  # TRADE
  add_equation("pm_uk = exp(nu0m + nu1m*log(py_us) + (1 - nu1m)*log(py_uk) - nu1m*log(xr_uk))", desc = "12.21 : Import prices in UK") |>
  add_equation("px_uk = exp(nu0x + nu1x*log(py_us) + (1 - nu1x)*log(py_uk) - nu1x*log(xr_uk))", desc = "12.22 : Export prices in UK") |>
  add_equation("px_us = pm_uk*xr_uk", desc = "12.23 : Export prices in US") |>
  add_equation("pm_us = px_uk*xr_uk", desc = "12.24 : Import prices in US") |>
  add_equation("x_k_uk = exp(eps0 - eps1*log(pm_us/py_us) + eps2*log(y_k_us))", desc = "12.25 : Real exports from UK - depends on current relative price") |>
  add_equation("im_k_uk = exp(mu0 - mu1*log(pm_uk[-1]/py_uk[-1]) + mu2*log(y_k_uk))", desc = "12.26 : Real imports of UK") |>
  add_equation("x_k_us = im_k_uk", desc = "12.27 : Real exports from US") |>
  add_equation("im_k_us = x_k_uk", desc = "12.28 : Real imports of US") |>
  add_equation("x_uk = x_k_uk*px_uk", desc = "12.29 : Exports of UK") |>
  add_equation("x_us = x_k_us*px_us", desc = "12.30 : Exports of US") |>
  add_equation("im_uk = im_k_uk*pm_uk", desc = "12.31 : Imports of UK") |>
  add_equation("im_us = im_k_us*pm_us", desc = "12.32 : Imports of US") |>
  
  # INCOME AND EXPENDITURE
  add_equation("v_k_uk = v_uk/pds_uk", desc = "12.33 : Real wealth in UK") |>
  add_equation("v_k_us = v_us/pds_us", desc = "12.34 : Real wealth in US") |>
  add_equation("ydhs_k_uk = yd_uk/pds_uk - v_k_uk[-1]*(pds_uk - pds_uk[-1])/pds_uk", desc = "12.35 : Real Haig-Simons disposable income in UK") |>
  add_equation("ydhs_k_us = yd_us/pds_us - v_k_us[-1]*(pds_us - pds_us[-1])/pds_us", desc = "12.36 : Real Haig-Simons disposable income in US") |>
  add_equation("c_k_uk = alpha1_uk*ydhse_k_uk + alpha2_uk*v_k_uk[-1]", desc = "12.37 : Real consumption in UK") |>
  add_equation("c_k_us = alpha1_us*ydhse_k_us + alpha2_us*v_k_us[-1]", desc = "12.38 : Real consumption in US") |>
  add_equation("ydhse_k_uk = (ydhs_k_uk + ydhs_k_uk[-1])/2", desc = "12.39 : Expected real Haig-Simons disposable income in UK") |>
  add_equation("ydhse_k_us = (ydhs_k_us + ydhs_k_us[-1])/2", desc = "12.40 : Expected real Haig-Simons disposable income in US") |>
  add_equation("s_k_uk = c_k_uk + g_k_uk + x_k_uk", desc = "12.41 : Real sales in UK") |>
  add_equation("s_k_us = c_k_us + g_k_us + x_k_us", desc = "12.42 : Real sales in US") |>
  add_equation("s_uk = s_k_uk*ps_uk", desc = "12.43 : Value of sales in UK") |>
  add_equation("s_us = s_k_us*ps_us", desc = "12.44 : Value of sales in US") |>
  add_equation("ps_uk = (1 + phi_uk)*(w_uk*n_uk + im_uk)/s_k_uk", desc = "12.45 : Price of sales in UK") |>
  add_equation("ps_us = (1 + phi_us)*(w_us*n_us + im_us)/s_k_us", desc = "12.46 : Price of sales in US") |>
  add_equation("pds_uk = (s_uk - x_uk)/(s_k_uk - x_k_uk)", desc = "12.47 : Price of domestic sales in UK") |>
  add_equation("pds_us = (s_us - x_us)/(s_k_us - x_k_us)", desc = "12.48 : Price of domestic sales in US") |>
  add_equation("ds_uk = s_uk - x_uk", desc = "12.49 : Domestic sales in UK") |>
  add_equation("ds_us = s_us - x_us", desc = "12.50 : Domestic sales in US") |>
  add_equation("ds_k_uk = c_k_uk + g_k_uk", desc = "12.51 : Real domestic sales in UK") |>
  add_equation("ds_k_us = c_k_us + g_k_us", desc = "12.52 : Real domestic sales in US") |>
  add_equation("y_uk = s_uk - im_uk", desc = "12.53 : Value of output in UK") |>
  add_equation("y_us = s_us - im_us", desc = "12.54 : Value of output in US") |>
  
  add_equation("y_k_uk = s_k_uk - im_k_uk", desc = "12.55 : Value of real output in UK") |>
  add_equation("y_k_us = s_k_us - im_k_us", desc = "12.56 : Value of real output in US") |>
  add_equation("py_uk = y_uk/y_k_uk", desc = "12.57 : Price of output in UK") |>
  add_equation("py_us = y_us/y_k_us", desc = "12.58 : Price of output in US") |>
  add_equation("cons_uk = c_k_uk*pds_uk", desc = "12.59 : Consumption in UK") |>
  add_equation("cons_us = c_k_us*pds_us", desc = "12.60 : Consumption in US") |>
  add_equation("g_uk = g_k_uk*pds_uk", desc = "12.61 : Government expenditure in UK") |>
  add_equation("g_us = g_k_us*pds_us", desc = "12.62 : Government expenditure in US") |>
  add_equation("n_uk = y_k_uk/pr_uk", desc = "12.65 : Employment in UK") |>
  add_equation("n_us = y_k_us/pr_us", desc = "12.66 : Employment in US") |>
  
  # ASSET DEMANDS
  add_equation("b_ukuk_d = v_uk*(lambda10 + lambda11*r_uk - lambda12*(r_us + dxre_us))", desc = "12.67 : Demand for UK bills in UK") |>
  add_equation("h_uk_d = v_uk - b_ukuk_d - b_ukus_d", desc = "12.69 : Demand for money in UK") |>
  add_equation("b_usus_d = v_us*(lambda40 + lambda41*r_us - lambda42*(r_uk + dxre_uk))", desc = "12.70 : Demand for US	bills in US") |>
  add_equation("b_usuk_d = v_us*(lambda50 - lambda51*r_us + lambda52*(r_uk + dxre_uk))", desc = "12.71 : Demand for UK bills in US") |>
  add_equation("h_us_d = v_us - b_usus_d - b_usuk_d", desc = "12.72 : Demand for money in US") |>
  
  # ASSET SUPPLIES
  add_equation("h_us_s = h_us_d", desc = "12.77 : Suply of cash in US") |>
  add_equation("b_usus_s = b_usus_d", desc = "12.78 : Supply of US bills to Country") |>
  add_equation("b_cb_usus_s = b_cb_usus_d", desc = "12.79 : Supply of US bills to US Central bank") |>
  add_equation("h_uk_s = h_uk_d", desc = "12.80 : Suply of cash in UK") |>
  add_equation("b_ukuk_s = b_ukuk_d", desc = "12.81 : Bills issued by US acquired by US") |>
  add_equation("b_cb_ukuk_s = b_cb_ukuk_d", desc = "12.82 : Supply of UK bills to UK Central bank") |>
  add_equation("b_cb_usus_d = b_cb_usus_d[-1] + (h_us_s - h_us_s[-1]) - (or_us - or_us[-1])*pg_us", desc = "12.83 : Balance sheet of US Central bank - expressed as changes") |>
  add_equation("b_cb_ukuk_d = b_cb_ukuk_d[-1] + (h_uk_s - h_uk_s[-1]) - (b_cb_ukus_s - b_cb_ukus_s[-1])*xr_us - (or_uk - or_uk[-1])*pg_uk", desc = "12.84 : Balance sheet of UK Central bank") |>
  add_equation("pg_uk = pg_us/xr_uk", desc = "12.85 : Price of gold is equal in the two countries") |>
  add_equation("xr_us = 1/xr_uk", desc = "12.86 : US exchange rate") |>
  add_equation("b_usuk_s = b_usuk_d*xr_us", desc = "12.87 : Equilibrium condition for bills issued by UK acquired by US") |>
  add_equation("b_cb_ukus_d = b_cb_ukus_s*xr_us", desc = "12.88 : Equilibrium condition for bills issued by US acquired by UK Central bank") |>
  add_equation("psbr_uk = g_uk + r_uk[-1]*b_uk_s[-1] - t_uk - f_cb_uk", desc = "Government deficit in the UK") |>
  add_equation("psbr_us = g_us + r_us[-1]*b_us_s[-1] - t_us - f_cb_us", desc = "Government deficit in the US") |>
  add_equation("nafa_uk = psbr_uk + cab_uk", desc = "Net accumulation of financial assets in the UK") |>
  add_equation("nafa_us = psbr_us + cab_us", desc = "Net accumulation of financial assets in the US") |>
  add_equation("b_cb_ukuk_sa = b_uk_s - b_ukuk_s - b_usuk_s", desc = "") |>
  add_equation("nwcb_uk = -h_uk_d + b_cb_ukuk_d + b_cb_ukus_d * xr_us + or_uk * pg_uk", desc = "Net wealth of CBUK") |>
  
  # Model OPEN FIX R closure
  add_equation("b_ukus_d = b_ukus_s*xr_us", desc = "12.89R : Demand of UK Bills in US") |>
  add_equation("b_ukus_s =  b_us_s - b_usus_s - b_cb_usus_d - b_cb_ukus_s", desc = "12.90R : Supply of UK bills to US") |>
  add_equation("r_uk = (lambda20 + lambda22*(r_us + dxre_us) - b_ukus_d/v_uk)/lambda21", desc = "12.68R: Endogenous interest rate")
```

You can then simulate the model; in this example, we run the baseline over 100 periods using Broyden’s method.
```{r}
# Simulate model
model_open <- simulate_scenario(model_open, scenario = "baseline",
                                max_iter = 350, periods = 100, tol = 1e-15,
                                method = "Broyden")
```

### Shock scenario
In this example, we observe the effect of an increase in the UK propensity to import.

Begin by creating an empty shock and setting the path for `mu0`:
```{r}
# Create empty shock and add shock equation
shock_open <- create_shock() |>
  add_shock(variable = "mu0", value = -2, start = 5, end = 100, desc = "")
```

Then add the shock to the model as a new scenario:
```{r}
# Create new scenario with this shock
model_open <- model_open |>
  add_scenario(name = "mu0_shock", origin = "baseline", shock = shock_open)
```

Run the model under this scenario:
```{r}
# Simulate shock
model_open <- simulate_scenario(model_open, scenario = "mu0_shock", periods = 100,
                                  max_iter = 350, tol = 1e-10, method = "Broyden",
                                  rhtol = TRUE)
```

You can now plot a few key variables to see how the model responds:
```{r, out.width="100%"}
# Figure 12.2A Effect of an increase in the UK propensity to import, within a fixed exchange rate regime with endogenous UK interest rates, on UK variables: capital account balance, trade balance, and current account balance
do_plotly(m = model_open, scenario = "mu0_shock",
        variables = c("cab_uk", "kabp_uk", "tab_uk"),
        t0 = 1, end = 50, title = "Scenario mu0_shock", y_title = "")
```

## 4. Model OPENFIXG

In this last specification (`OPENFIXG` model), we look at **another alternative fixed exchange rate closure: one country sets its government expenditures endogenously**.

Governments can also avoid devaluation by reducing public expenditures. In this case, fiscal policy is adjusted to preserve a fixed rate of exchange despite a deteriorated trade balance.
The new fiscal stance has to be such that after allowing for all lags and feedbacks, the level of domestic output is reduced to such an extent that the value of imports exactly equals that of exports. 

Closely following Godley and Lavoie, we again start from the main closure with a fixed exchange rate, but impose upon it that the central bank declines to use its foreign reserves, as if the country were on a flexible exchange rate. As a consequence, another variable, which was previously assumed to be constant, must become endogenous and that will be the (pure) government expenditures in the UK, $g^{£}$. In the resulting closure, nominal government expenditures are constrained by equation (12.13G). Real government expenditures are thus adjusted to accord with the nominal amount, through equation (12.61G). Equation (12.82AG) determines the supply of bills of the UK government. Equation (12.89G) follows from the fact that we remain within a fixed exchange rate regime and finally (12.90GA) is a redundant equation for our new closure.

### Base scenario
Create one more model — the last one in this chapter:
```{r}
# Create empty model
model_open <- create_model(name = "SFC Advanced OPENFIXG")
```

Declare all variables that appear in its specification.
```{r}
# Add variables
model_open <- model_open |>
  add_variable("alpha1_uk", init = 0.75) |>
  add_variable("alpha1_us", init = 0.75) |>
  add_variable("alpha2_uk", init = 0.13333) |>
  add_variable("alpha2_us", init = 0.13333) |>
  add_variable("eps0", init = -2.1) |>
  add_variable("eps1", init = 0.7) |>
  add_variable("eps2", init = 1) |>
  add_variable("lambda10", init = 0.7) |>
  add_variable("lambda11", init = 5) |>
  add_variable("lambda12", init = 5) |>
  add_variable("lambda20", init = 0.25) |>
  add_variable("lambda21", init = 5) |>
  add_variable("lambda22", init = 5) |>
  add_variable("lambda40", init = 0.7) |>
  add_variable("lambda41", init = 5) |>
  add_variable("lambda42", init = 5) |>
  add_variable("lambda50", init = 0.25) |>
  add_variable("lambda51", init = 5) |>
  add_variable("lambda52", init = 5) |>
  add_variable("mu0", init = -2.1) |>
  add_variable("mu1", init = 0.7) |>
  add_variable("mu2", init = 1) |>
  add_variable("nu0m", init = -0.00001) |>
  add_variable("nu0x", init = -0.00001) |>
  add_variable("nu1m", init = 0.7) |>
  add_variable("nu1x", init = 0.5) |>
  add_variable("phi_uk", init = 0.2381) |>
  add_variable("phi_us", init = 0.2381) |>
  add_variable("theta_uk", init = 0.2) |>
  add_variable("theta_us", init = 0.2) |>
  add_variable("b_cb_ukus_s", init = 0.02031) |>
  add_variable("dxre_us", init = 0) |>
  add_variable("g_k_uk") |>
  add_variable("g_k_us", init = 16) |>
  add_variable("or_uk", init = 7) |>
  add_variable("pg_us", init = 1) |>
  add_variable("pr_uk", init = 1.3333) |>
  add_variable("pr_us", init = 1.3333) |>
  add_variable("r_uk", init = 0.03) |>
  add_variable("r_us", init = 0.03) |>
  add_variable("w_uk", init = 1) |>
  add_variable("w_us", init = 1) |>
  add_variable("b_cb_ukuk_d", init = 0.27984) |>
  add_variable("b_cb_ukuk_s", init = 0.27984) |>
  add_variable("b_cb_ukuk_sa", init = 0.27984) |>
  add_variable("b_cb_ukus_d", init = 0.0203) |>
  add_variable("b_cb_usus_d", init = 0.29843) |>
  add_variable("b_cb_usus_s", init = 0.29843) |>
  add_variable("b_uk_s", init = 138.94) |>
  add_variable("b_ukuk_d", init = 102.18) |>
  add_variable("b_ukuk_s", init = 102.18) |>
  add_variable("b_ukus_d", init = 36.493) |>
  add_variable("b_ukus_s", init = 36.504) |>
  add_variable("b_us_s", init = 139.02) |>
  add_variable("b_usuk_d", init = 36.497) |>
  add_variable("b_usuk_s", init = 36.487) |>
  add_variable("b_usus_d", init = 102.19) |>
  add_variable("b_usus_s", init = 102.19) |>
  add_variable("h_uk_d", init = 7.2987) |>
  add_variable("h_uk_s", init = 7.2987) |>
  add_variable("h_us_d", init = 7.2995) |>
  add_variable("h_us_s", init = 7.2995) |>
  add_variable("or_us", init = 7) |>
  add_variable("v_k_uk", init = 152.62) |>
  add_variable("v_k_us", init = 152.63) |>
  add_variable("v_uk", init = 145.97) |>
  add_variable("v_us", init = 145.99001) |>
  add_variable("c_k_uk", init = 81.393) |>
  add_variable("c_k_us", init = 81.401) |>
  add_variable("cab_uk", init = 0) |>
  add_variable("cab_us", init = 0) |>
  add_variable("cons_uk", init = 77.851) |>
  add_variable("cons_us", init = 77.86) |>
  add_variable("ds_k_uk", init = 97.393) |>
  add_variable("ds_k_us", init = 97.401) |>
  add_variable("ds_uk", init = 93.154) |>
  add_variable("ds_us", init = 93.164) |>
  add_variable("dxre_uk", init = 0) |>
  add_variable("f_cb_uk", init = 0.00869) |>
  add_variable("f_cb_us", init = 0.00895) |>
  add_variable("g_uk", init = 15.304) |>
  add_variable("g_us", init = 15.304) |>
  add_variable("im_k_uk", init = 11.928) |>
  add_variable("im_k_us", init = 11.926) |>
  add_variable("im_uk", init = 11.407) |>
  add_variable("im_us", init = 11.409) |>
  add_variable("kabp_uk", init = 0.00002) |>
  add_variable("kabp_us", init = -0.00002) |>
  add_variable("n_uk", init = 73.046) |>
  add_variable("n_us", init = 73.054) |>
  add_variable("pds_uk", init = 0.95648) |>
  add_variable("pds_us", init = 0.95649) |>
  add_variable("pg_uk", init = 0.99971) |>
  add_variable("pm_uk", init = 0.95628) |>
  add_variable("pm_us", init = 0.95661) |>
  add_variable("ps_uk", init = 0.95646) |>
  add_variable("ps_us", init = 0.9565) |>
  add_variable("px_uk", init = 0.95634) |>
  add_variable("px_us", init = 0.95656) |>
  add_variable("py_uk", init = 0.95648) |>
  add_variable("py_us", init = 0.95649) |>
  add_variable("s_k_uk", init = 109.32) |>
  add_variable("s_k_us", init = 109.33) |>
  add_variable("s_uk", init = 104.56) |>
  add_variable("s_us", init = 104.57) |>
  add_variable("t_uk", init = 19.463) |>
  add_variable("t_us", init = 19.465) |>
  add_variable("x_k_uk", init = 11.926) |>
  add_variable("x_k_us", init = 11.928) |>
  add_variable("x_uk", init = 11.406) |>
  add_variable("x_us", init = 11.41) |>
  add_variable("xr_uk", init = 1.0003) |>
  add_variable("xr_us", init = 0.99971) |>
  add_variable("xre_uk", init = 1.0003) |>
  add_variable("xre_us", init = 0.99971) |>
  add_variable("y_k_uk", init = 97.392) |>
  add_variable("y_k_us", init = 97.403) |>
  add_variable("y_uk", init = 93.154) |>
  add_variable("y_us", init = 93.164) |>
  add_variable("yd_uk", init = 77.851) |>
  add_variable("yd_us", init = 77.86) |>
  add_variable("ydhs_k_uk", init = 81.394) |>
  add_variable("ydhs_k_us", init = 81.402) |>
  add_variable("ydhse_k_uk", init = 81.394) |>
  add_variable("ydhse_k_us", init = 81.402) |>
  add_variable("yd_hs_uk") |>
  add_variable("yd_hs_us") |>
  add_variable("kab_uk") |>
  add_variable("kab_us") |>
  add_variable("psbr_uk") |>
  add_variable("psbr_us") |>
  add_variable("nafa_uk") |>
  add_variable("nafa_us") |>
  add_variable("nwcb_uk")
```

Add the equations that describe how the variables interact.
```{r}
# Add equations
model_open <- model_open |>
  add_equation("yd_uk = (y_uk + r_uk[-1]*b_ukuk_d[-1] + xr_us*r_us[-1]*b_ukus_s[-1])*(1 - theta_uk) + (xr_us - xr_us[-1])*b_ukus_s[-1]", desc = "12.1 : Disposable income in UK") |>
  add_equation("yd_hs_uk = yd_uk + (xr_us - xr_us[-1])*b_ukus_s[-1]", desc = "12.2 : Haig-Simons disposable income in UK") |>
  add_equation("v_uk = v_uk[-1] + yd_uk - cons_uk", desc = "12.3 : Wealth accumulation in UK") |>
  add_equation("yd_us = (y_us + r_us[-1]*b_usus_d[-1] + xr_uk*r_uk[-1]*b_usuk_s[-1])*(1 - theta_us) + (xr_uk - xr_uk[-1])*b_usuk_s[-1]", desc = "12.4 : Disposable income in US") |>
  add_equation("yd_hs_us = yd_us + d(xr_uk)*b_usuk_s[-1]", desc = "12.5 : Haig-Simons disposable income in US") |>
  add_equation("v_us = v_us[-1] + yd_us - cons_us", desc = "12.6 : Wealth accumulation in US") |>
  add_equation("t_uk = theta_uk*(y_uk + r_uk[-1]*b_ukuk_d[-1] + xr_us*r_us[-1]*b_ukus_s[-1])", desc = "12.7 : Taxes in UK") |>
  add_equation("t_us = theta_us*(y_us + r_us[-1]*b_usus_d[-1] + xr_uk*r_uk[-1]*b_usuk_s[-1])", desc = "12.8 : Taxes in US") |>
  add_equation("f_cb_uk = r_uk[-1]*b_cb_ukuk_d[-1] + r_us[-1]*b_cb_ukus_s[-1]*xr_us", desc = "12.11 : Profits of Central Bank in UK") |>
  add_equation("f_cb_us = r_us[-1]*b_cb_usus_d[-1]", desc = "12.12 : Profits of Central Bank in US") |>
  add_equation("b_us_s = b_us_s[-1] + g_us + r_us[-1]*b_us_s[-1] - t_us - f_cb_us", desc = "12.14 : Government budget constraint - US") |>
  add_equation("cab_uk = x_uk - im_uk + xr_us*r_us[-1]*b_ukus_s[-1] - r_uk[-1]*b_usuk_s[-1] + r_us[-1]*b_cb_ukus_s[-1]*xr_us", desc = "12.15 : Current account balance - UK") |>
  add_equation("kab_uk = kabp_uk - (xr_us*(b_cb_ukus_s - b_cb_ukus_s[-1]) + pg_uk*(or_uk - or_uk[-1]))", desc = "12.16 : Capital account balance in UK") |>
  add_equation("cab_us = x_us - im_us + xr_uk*r_uk[-1]*b_usuk_s[-1] - r_us[-1]*b_ukus_s[-1] - r_us[-1]*b_cb_ukus_s[-1]", desc = "12.17 : Current account balance in US") |>
  add_equation("kab_us = kabp_us + (b_cb_ukus_s - b_cb_ukus_s[-1]) - pg_us*(or_us - or_us[-1])", desc = "12.18 : Capital account balance in US") |>
  add_equation("kabp_uk = -(b_ukus_s - b_ukus_s[-1])*xr_us + (b_usuk_s - b_usuk_s[-1])", desc = "12.19 : Capital account balance in UK, net of official transactions") |>
  add_equation("kabp_us = -(b_usuk_s - b_usuk_s[-1])*xr_uk + (b_ukus_s - b_ukus_s[-1])", desc = "12.20 : Capital account balance in US, net of official transactions") |>
  
  # TRADE
  add_equation("pm_uk = exp(nu0m + nu1m*log(py_us) + (1 - nu1m)*log(py_uk) - nu1m*log(xr_uk))", desc = "12.21 : Import prices in UK") |>
  add_equation("px_uk = exp(nu0x + nu1x*log(py_us) + (1 - nu1x)*log(py_uk) - nu1x*log(xr_uk))", desc = "12.22 : Export prices in UK") |>
  add_equation("px_us = pm_uk*xr_uk", desc = "12.23 : Export prices in US") |>
  add_equation("pm_us = px_uk*xr_uk", desc = "12.24 : Import prices in US") |>
  add_equation("x_k_uk = exp(eps0 - eps1*log(pm_us/py_us) + eps2*log(y_k_us))", desc = "12.25 : Real exports from UK - depends on current relative price") |>
  add_equation("im_k_uk = exp(mu0 - mu1*log(pm_uk[-1]/py_uk[-1]) + mu2*log(y_k_uk))", desc = "12.26 : Real imports of UK") |>
  add_equation("x_k_us = im_k_uk", desc = "12.27 : Real exports from US") |>
  add_equation("im_k_us = x_k_uk", desc = "12.28 : Real imports of US") |>
  add_equation("x_uk = x_k_uk*px_uk", desc = "12.29 : Exports of UK") |>
  add_equation("x_us = x_k_us*px_us", desc = "12.30 : Exports of US") |>
  add_equation("im_uk = im_k_uk*pm_uk", desc = "12.31 : Imports of UK") |>
  add_equation("im_us = im_k_us*pm_us", desc = "12.32 : Imports of US") |>
  
  # INCOME AND EXPENDITURE
  add_equation("v_k_uk = v_uk/pds_uk", desc = "12.33 : Real wealth in UK") |>
  add_equation("v_k_us = v_us/pds_us", desc = "12.34 : Real wealth in US") |>
  add_equation("ydhs_k_uk = yd_uk/pds_uk - v_k_uk[-1]*(pds_uk - pds_uk[-1])/pds_uk", desc = "12.35 : Real Haig-Simons disposable income in UK") |>
  add_equation("ydhs_k_us = yd_us/pds_us - v_k_us[-1]*(pds_us - pds_us[-1])/pds_us", desc = "12.36 : Real Haig-Simons disposable income in US") |>
  add_equation("c_k_uk = alpha1_uk*ydhse_k_uk + alpha2_uk*v_k_uk[-1]", desc = "12.37 : Real consumption in UK") |>
  add_equation("c_k_us = alpha1_us*ydhse_k_us + alpha2_us*v_k_us[-1]", desc = "12.38 : Real consumption in US") |>
  add_equation("ydhse_k_uk = (ydhs_k_uk + ydhs_k_uk[-1])/2", desc = "12.39 : Expected real Haig-Simons disposable income in UK") |>
  add_equation("ydhse_k_us = (ydhs_k_us + ydhs_k_us[-1])/2", desc = "12.40 : Expected real Haig-Simons disposable income in US") |>
  add_equation("s_k_uk = c_k_uk + g_k_uk + x_k_uk", desc = "12.41 : Real sales in UK") |>
  add_equation("s_k_us = c_k_us + g_k_us + x_k_us", desc = "12.42 : Real sales in US") |>
  add_equation("s_uk = s_k_uk*ps_uk", desc = "12.43 : Value of sales in UK") |>
  add_equation("s_us = s_k_us*ps_us", desc = "12.44 : Value of sales in US") |>
  add_equation("ps_uk = (1 + phi_uk)*(w_uk*n_uk + im_uk)/s_k_uk", desc = "12.45 : Price of sales in UK") |>
  add_equation("ps_us = (1 + phi_us)*(w_us*n_us + im_us)/s_k_us", desc = "12.46 : Price of sales in US") |>
  add_equation("pds_uk = (s_uk - x_uk)/(s_k_uk - x_k_uk)", desc = "12.47 : Price of domestic sales in UK") |>
  add_equation("pds_us = (s_us - x_us)/(s_k_us - x_k_us)", desc = "12.48 : Price of domestic sales in US") |>
  add_equation("ds_uk = s_uk - x_uk", desc = "12.49 : Domestic sales in UK") |>
  add_equation("ds_us = s_us - x_us", desc = "12.50 : Domestic sales in US") |>
  add_equation("ds_k_uk = c_k_uk + g_k_uk", desc = "12.51 : Real domestic sales in UK") |>
  add_equation("ds_k_us = c_k_us + g_k_us", desc = "12.52 : Real domestic sales in US") |>
  add_equation("y_uk = s_uk - im_uk", desc = "12.53 : Value of output in UK") |>
  add_equation("y_us = s_us - im_us", desc = "12.54 : Value of output in US") |>
  add_equation("y_k_uk = s_k_uk - im_k_uk", desc = "12.55 : Value of real output in UK") |>
  add_equation("y_k_us = s_k_us - im_k_us", desc = "12.56 : Value of real output in US") |>
  add_equation("py_uk = y_uk/y_k_uk", desc = "12.57 : Price of output in UK") |>
  add_equation("py_us = y_us/y_k_us", desc = "12.58 : Price of output in US") |>
  add_equation("cons_uk = c_k_uk*pds_uk", desc = "12.59 : Consumption in UK") |>
  add_equation("cons_us = c_k_us*pds_us", desc = "12.60 : Consumption in US") |>
  add_equation("g_us = g_k_us*pds_us", desc = "12.62 : Government expenditure in US") |>
  add_equation("n_uk = y_k_uk/pr_uk", desc = "12.65 : Employment in UK") |>
  add_equation("n_us = y_k_us/pr_us", desc = "12.66 : Employment in US") |>
  
  # ASSET DEMANDS
  add_equation("b_ukuk_d = v_uk*(lambda10 + lambda11*r_uk - lambda12*(r_us + dxre_us))", desc = "12.67 : Demand for UK bills in UK") |>
  add_equation("b_ukus_d = v_uk*(lambda20 - lambda21*r_uk + lambda22*(r_us + dxre_us))", desc = "12.68F : Demand for US bills in UK") |>
  add_equation("h_uk_d = v_uk - b_ukuk_d - b_ukus_d", desc = "12.69 : Demand for money in UK") |>
  add_equation("b_usus_d = v_us*(lambda40 + lambda41*r_us - lambda42*(r_uk + dxre_uk))", desc = "12.70 : Demand for US	bills in US") |>
  add_equation("b_usuk_d = v_us*(lambda50 - lambda51*r_us + lambda52*(r_uk + dxre_uk))", desc = "12.71 : Demand for UK bills in US") |>
  add_equation("h_us_d = v_us - b_usus_d - b_usuk_d", desc = "12.72 : Demand for money in US") |>
  
  # ASSET SUPPLIES
  add_equation("h_us_s = h_us_d", desc = "12.77 : Suply of cash in US") |>
  add_equation("b_usus_s = b_usus_d", desc = "12.78 : Supply of US bills to Country") |>
  add_equation("b_cb_usus_s = b_cb_usus_d", desc = "12.79 : Supply of US bills to US Central bank") |>
  add_equation("h_uk_s = h_uk_d", desc = "12.80 : Suply of cash in UK") |>
  add_equation("b_ukuk_s = b_ukuk_d", desc = "12.81 : Bills issued by US acquired by US") |>
  add_equation("b_cb_usus_d = b_cb_usus_d[-1] + (h_us_s - h_us_s[-1]) - (or_us - or_us[-1])*pg_us", desc = "12.83 : Balance sheet of US Central bank - expressed as changes") |>
  add_equation("b_cb_ukuk_d = b_cb_ukuk_d[-1] + (h_uk_s - h_uk_s[-1]) - (b_cb_ukus_s - b_cb_ukus_s[-1])*xr_us - (or_uk - or_uk[-1])*pg_uk", desc = "12.84 : Balance sheet of UK Central bank") |>
  add_equation("pg_uk = pg_us/xr_uk", desc = "12.85 : Price of gold is equal in the two countries") |>
  add_equation("xr_us = 1/xr_uk", desc = "12.86 : US exchange rate") |>
  add_equation("b_usuk_s = b_usuk_d*xr_us", desc = "12.87 : Equilibrium condition for bills issued by UK acquired by US") |>
  add_equation("b_cb_ukus_d = b_cb_ukus_s*xr_us", desc = "12.88 : Equilibrium condition for bills issued by US acquired by UK Central bank") |>
  add_equation("psbr_uk = g_uk + r_uk[-1]*b_uk_s[-1] - t_uk - f_cb_uk", desc = "Government deficit in the UK") |>
  add_equation("psbr_us = g_us + r_us[-1]*b_us_s[-1] - t_us - f_cb_us", desc = "Government deficit in the US") |>
  add_equation("nafa_uk = psbr_uk + cab_uk", desc = "Net accumulation of financial assets in the UK") |>
  add_equation("nafa_us = psbr_us + cab_us", desc = "Net accumulation of financial assets in the US") |>
  add_equation("b_cb_ukuk_sa = b_uk_s - b_ukuk_s - b_usuk_s", desc = "") |>
  add_equation("nwcb_uk = -h_uk_d + b_cb_ukuk_d + b_cb_ukus_d * xr_us + or_uk * pg_uk", desc = "Net wealth of CBUK") |>
  
  # Model OPEN FIX G Closure
  add_equation("b_ukus_s = xr_uk*b_ukus_d", desc = "12.89G : Bills supply from UK to US") |>
  add_equation("b_cb_ukus_s = b_us_s - b_usus_s - b_cb_usus_d - b_ukus_s", desc = "12.90G : Supply of UK bills to US") |>
  add_equation("g_uk = b_uk_s - b_uk_s[-1] -(r_uk[-1]*b_uk_s[-1] - t_uk - f_cb_uk)", desc = "12.13G: Endogenous govt. expenditures in UK") |>
  add_equation("g_k_uk = g_uk / pds_uk", desc = "12.61G: Real government expenditures") |>
  add_equation("b_uk_s = b_usuk_s + b_ukuk_s + b_cb_ukuk_s", desc = "12.82AG: Supply of UK Bills")
```

And run the baseline simulation:
```{r}
# Simulate model
model_open <- simulate_scenario(model_open, scenario = "baseline",
                                max_iter = 350, periods = 100, tol = 1e-15,
                                method = "Broyden")
```

### Shock scenario
We close with one more experiment. We take another look at the effects of an increase in the UK propensity to import.

As before, the first step is to define the shock by setting a new path for `mu0`.
```{r}
# Create empty shock and add shock equation
shock_open <- create_shock() |>
  add_shock(variable = "mu0", value = -2, start = 5, end = 100, desc = "")
```

We add it to the model as a separate scenario.
```{r}
# Create new scenario with this shock
model_open <- model_open |>
  add_scenario(name = "mu0_shock", origin = "baseline", shock = shock_open)
```

We then run the simulation to see how the economy adjusts under this change.
```{r}
# Simulate shock
model_open <- simulate_scenario(model_open, scenario = "mu0_shock",
                                max_iter = 350, periods = 100, tol = 1e-15,
                                method = "Broyden")
```

To round things off, we look at the longer-run effects by plotting a few variables discussed by *Godley* and *Lavoie* in their book.
```{r, out.width="100%"}
# Figure 12.3A Effect of an increase in the UK propensity to import, within a fixed exchange rate regime with endogenous UK government expenditures, on the US and UK real GDP
do_plotly(m = model_open, scenario = "mu0_shock",
        variables = c("y_k_us", "y_k_uk"),
        t0 = 1, end = 50, title = "Scenario mu0_shock", y_title = "")
```

```{r, out.width="100%"}
# Figure 12.3B Effect of an increase in the UK propensity to import, within a fixed exchange rate regime with endogenous UK government expenditures, on various UK variables: the capital account balance, the current account balance and the trade balance
do_plotly(m = model_open, scenario = "mu0_shock",
        variables = c("cab_uk", "tab_uk", "kabp_uk"),
        t0 = 1, end = 50, title = "Scenario mu0_shock", y_title = "")
```

```{r, out.width="100%"}
# Figure 12.3C Effect of an increase in the UK propensity to import, within a fixed exchange rate regime with endogenous UK government expenditures, on various UK variables: the current account balance, the government budget balance, and the net accumulation of financial assets
do_plotly(m = model_open, scenario = "mu0_shock",
        variables = c("cab_uk", "gab_uk", "nafa_uk"),
        t0 = 1, end = 50, title = "Scenario mu0_shock", y_title = "")
```

```{r, out.width="100%"}
# Figure 12.3D Effect of an increase in the UK propensity to import, within a fixed exchange rate regime with endogenous UK government expenditures, on the UK and US debt to GDP ratios
do_plotly(m = model_open, scenario = "mu0_shock",
        variables = c("by_us", "by_uk"),
        t0 = 1, end = 50, title = "Scenario mu0_shock", y_title = "")
```

### References
For the full model discussion and interpretation of the GROWTH model, refer to Chapter 12 of *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*
