---
title: "Model INSOUT"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model INSOUT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include = FALSE}
library(dplyr)
library(tidyr)
library(plotly)

`%||%` <- function(a, b) if (!is.null(a) && length(a) > 0) a else b

plotly_style <- function(fig, title = NULL) {
  fig %>%
    plotly::layout(
      title = list(
        text = title %||% "",
        y = 0.95, x = 0.5,
        xanchor = "center", yanchor = "top"
      ),
      margin = list(t = 60, b = 70, l = 60, r = 120),
      hovermode = "spikers",
      spikedistance = 1000,
      xaxis = list(
        title = "time"
      ),
      yaxis = list(
        title = ""
      ),
      showlegend = TRUE,
      legend = list(
        xanchor = "left",
        x = 1.02,
        y = 0.95
      ),
      font = list(
        family = "Arial",
        size = 13
      )
    )
}

do_plotly <- function(m, scenario, variables, t0 = 1,
                      start = NULL, end = NULL,
                      title = NULL,
                      y_title = NULL) {
  
  m1 <- m[[scenario]][["result"]]
  
  start <- if (is.null(start)) t0 else start
  end   <- if (is.null(end)) max(m1[["time"]]) else end
  
  # The lookup_names table has the code of the variables
  # and the name I wanted displayed in the Figures.
  lookup_names <- tribble(
    ~name, ~Var,
    "dBhh", "HH. Bills",
    "dV", "Wealth",
    "dM1s", "Checking deposits",
    "dM2s", "Time deposits",
    "dHhs", "Cash",
    "dBLh", "HH. Bonds",
    "yr", "Real output",
    "cr", "Real consumption",
    "ydr", "Real disposable income",
    "c", "Real consumption",
    "v", "Real wealth",
    "rm", "Interest on term deposits",
    "rb", "Interest on bills",
    "rl", "Interest on loans",
    "rbl", "Interest on bonds",
    "dLs", "Loans",
    "dM", "All money deposits",
    "M", "All money deposits",
    "dAs", "CB advances",
    "dHbs", "Reserves",
    "dBbd", "B. Bills",
    "BLRN", "Net bank liq. ratio",
    "BPM", "Bank profit margin",
    "dGb", "Government budget balance",
    "Bcb", "CB. Bills",
    "dPSBR", "Deflated PSBR",
    "pi", "Inflation",
    "gr", "Real govt. budget balance - Book",
    "gr2", "Real govt. budget balance - tav",
    "gr3", "Real govt. budget balance - cb adj.",
    "BYR", "Debt to GDP ratio",
    "dCB", "Change in the CB' stock of bills",
    "ddAs", "Change in Advances to banks",
    "y", "Real output",
    "s", "Real sales"
  )
  
  df_long <- m1 %>%
    mutate(yr = y / y[t0],
           cr = c / c[t0],
           dV = V - V[t0],
           dM1s = M1s - M1s[t0],
           dM2s = M2s - M2s[t0],
           dBhh = Bhh - Bhh[t0],
           dHhs = Hhs - Hhs[t0],
           dBLh = (pbl * BLh) - (pbl[t0] * BLh[t0]),
           dLs = Ls - Ls[t0],
           M = M1s + M2s,
           dM = M - M[t0],
           dAs = As - As[t0],
           dHbs = Hbs - Hbs[t0],
           dBbd = Bbd - Bbd[t0],
           dGb = -1 * (PSBR - PSBR[t0]),
           dPSBR = PSBR / p,
           tav = (p - lag(p)) / p,
           gr = ((p - lag(p)) * (lag(Bs) + lag(BLs) * lag(pbl)) - PSBR) / p,
           gr2 = (tav * (lag(Bs) + lag(BLs) * lag(pbl)) - PSBR) / p,
           gr3 = (tav * (lag(Bs) - lag(Bcb) + lag(BLs) * lag(pbl) - lag(As)) - PSBR)/p,
           BYR = (Bs + pbl * BLs) / Y,
           dCB = Bcb - lag(Bcb),
           ddAs = As - lag(As)
    ) %>%
    filter(time >= start & time <= end) %>%
    select(time, all_of(variables)) %>%
    pivot_longer(cols = -time) %>%
    # Left-join the lookup_names table
    left_join(lookup_names, by = "name") %>%
    mutate(Var = if_else(is.na(Var), name, Var))
  
  fig <- plotly::plot_ly()
  
  for (v in unique(df_long$Var)) {
    dfi <- df_long %>% filter(Var == v)
    
    fig <- plotly::add_trace(
      fig,
      data = dfi,
      x = ~time,
      y = ~value,
      name = v,
      type = "scatter",
      mode = "lines",
      hovertemplate = paste(
        "<b>", v, "</b>",
        "<br>time=%{x}",
        "<br>value=%{y}<extra></extra>"
      )
    )
  }
  
  fig %>%
    plotly::layout(
      yaxis = list(title = y_title %||% "")
    ) %>%
    plotly_style(title)
}
```

This guide explains how to use the `godley` package to create a model with both inside and outside money, as described by *Wynne Godley* and *Marc Lavoie* in Chapter 10 of *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth*.

```{r setup}
library(godley)
library(purrr)
library(plotly)
```

### Base scenario
Create an empty SFC model to use as our starting point:
```{r}
# Create empty model
model_insout <- create_model(name = "SFC INSOUT")
```

Next, declare all the variables used in the model. You can also set initial values upfront — for some variables, all of them, or none.
```{r}
# Add variables
model_insout <- model_insout |>
  add_variable("rbl", init = 0.027) |>
  add_variable("rb", init = 0.023) |>
  add_variable("pr", init = 1) |>
  add_variable("g", init = 25) |>
  add_variable("Nfe", init = 133.28) |>
  add_variable("alpha0", init = 0) |>
  add_variable("alpha1", init = 0.95) |>
  add_variable("alpha2", init = 0.05) |>
  add_variable("beta", init = 0.5) |>
  add_variable("bot", init = 0.02) |>
  add_variable("botpm", init = 0.003) |>
  add_variable("epsilon", init = 0.5) |>
  add_variable("gamma", init = 0.5) |>
  add_variable("lambdac", init = 0.1) |>
  add_variable("phi", init = 0.1) |>
  add_variable("ro1", init = 0.1) |>
  add_variable("ro2", init = 0.1) |>
  add_variable("sigma0", init = 0.3612) |>
  add_variable("sigma1", init = 3) |>
  add_variable("tau", init = 0.25) |>
  add_variable("zetab", init = 0.9) |>
  add_variable("zetal", init = 0.0002) |>
  add_variable("zetam", init = 0.0002) |>
  add_variable("Omega0", init = -0.32549) |>
  add_variable("Omega1", init = 1) |>
  add_variable("Omega2", init = 1.5) |>
  add_variable("Omega3", init = 0.1) |>
  add_variable("top", init = 0.06) |>
  add_variable("toppm", init = 0.005) |>
  add_variable("lambda10", init = -0.17071) |>
  add_variable("lambda11", init = 0) |>
  add_variable("lambda12", init = 0) |>
  add_variable("lambda13", init = 0) |>
  add_variable("lambda14", init = 0) |>
  add_variable("lambda15", init = 0.18) |>
  add_variable("lambda20", init = 0.52245) |>
  add_variable("lambda21", init = 0) |>
  add_variable("lambda22", init = 30) |>
  add_variable("lambda23", init = -15) |>
  add_variable("lambda24", init = -15) |>
  add_variable("lambda25", init = -0.06) |>
  add_variable("lambda30", init = 0.47311) |>
  add_variable("lambda31", init = 0) |>
  add_variable("lambda32", init = -15) |>
  add_variable("lambda33", init = 30) |>
  add_variable("lambda34", init = -15) |>
  add_variable("lambda35", init = -0.06) |>
  add_variable("lambda40", init = 0.17515) |>
  add_variable("lambda41", init = 0) |>
  add_variable("lambda42", init = -15) |>
  add_variable("lambda43", init = -15) |>
  add_variable("lambda44", init = 30) |>
  add_variable("lambda45", init = -0.06) |>
  add_variable("y") |>
  add_variable("sE") |>
  add_variable("invE") |>
  add_variable("inv") |>
  add_variable("N") |>
  add_variable("WB") |>
  add_variable("W", init = 1) |>
  add_variable("UC", init = 1) |>
  add_variable("s") |>
  add_variable("invT") |>
  add_variable("sigmaT") |>
  add_variable("rl", init = 0.02) |>
  add_variable("p", init = 1) |>
  add_variable("NHUC") |>
  add_variable("c") |>
  add_variable("S") |>
  add_variable("sigmas") |>
  add_variable("INV") |>
  add_variable("Ld") |>
  add_variable("FXf") |>
  add_variable("TX") |>
  add_variable("pi") |>
  add_variable("YDr") |>
  add_variable("FX") |>
  add_variable("rm", init = 0.02) |>
  add_variable("M2h") |>
  add_variable("Bhh") |>
  add_variable("BLh") |>
  add_variable("CG") |>
  add_variable("pbl") |>
  add_variable("YDhs") |>
  add_variable("FXb") |>
  add_variable("V") |>
  add_variable("C") |>
  add_variable("Vnc") |>
  add_variable("Hhh") |>
  add_variable("ydr") |>
  add_variable("ydhs") |>
  add_variable("v") |>
  add_variable("ydrE") |>
  add_variable("YDrE") |>
  add_variable("VE") |>
  add_variable("Hhd") |>
  add_variable("VncE") |>
  add_variable("ERrbl") |>
  add_variable("M2d") |>
  add_variable("Bhd") |>
  add_variable("BLd") |>
  add_variable("M1d") |>
  add_variable("M1d2") |>
  add_variable("M1hN") |>
  add_variable("z1") |>
  add_variable("z2") |>
  add_variable("M1h") |>
  add_variable("M2hN") |>
  add_variable("G") |>
  add_variable("PSBR") |>
  add_variable("Bs") |>
  add_variable("BLs") |>
  add_variable("FXcb") |>
  add_variable("GD") |>
  add_variable("Hs") |>
  add_variable("Bcb") |>
  add_variable("As") |>
  add_variable("Hbs") |>
  add_variable("Hhs") |>
  add_variable("Bbd") |>
  add_variable("Ad") |>
  add_variable("ra") |>
  add_variable("M1s") |>
  add_variable("M2s") |>
  add_variable("Ls") |>
  add_variable("Hbd") |>
  add_variable("BbdN") |>
  add_variable("BLRN") |>
  add_variable("z3") |>
  add_variable("BLR") |>
  add_variable("z4") |>
  add_variable("z5") |>
  add_variable("z6") |>
  add_variable("z7") |>
  add_variable("BPM", init = 0.0035) |>
  add_variable("lM1s") |>
  add_variable("lM2s") |>
  add_variable("omegaT") |>
  add_variable("Y")
```

Then, add the equations that define the relationships between variables.
```{r}
# Add equations
model_insout <- model_insout |>
  # Firm's behavioral equations
  add_equation("y = sE + (invE - inv[-1])", desc = "") |>
  add_equation("N = y / pr", desc = "") |>
  add_equation("WB = N * W", desc = "") |>
  add_equation("UC = WB / y", desc = "") |>
  add_equation("sE = beta * s[-1] + (1 - beta) * sE[-1]", desc = "") |>
  add_equation("invT = sigmaT * sE", desc = "") |>
  add_equation("sigmaT = sigma0 - sigma1 * rl", desc = "") |>
  add_equation("invE = inv[-1] + gamma * (invT - inv[-1])", desc = "") |>
  add_equation("p = (1 + tau) * (1 + phi) * NHUC", desc = "") |>
  add_equation("NHUC = (1 - sigmaT) * UC + sigmaT * (1 + rl[-1]) * UC[-1]", desc = "") |>
  
  # Firm's realized outcomes
  add_equation("s = c + g", desc = "") |>
  add_equation("S = s * p", desc = "") |>
  add_equation("inv = inv[-1] + y - s", desc = "") |>
  add_equation("sigmas = inv[-1] / s", desc = "") |>
  add_equation("INV = inv * UC", desc = "") |>
  add_equation("Ld = INV", desc = "") |>
  add_equation("FXf = S - TX - WB + (INV - INV[-1]) - rl[-1] * INV[-1]", desc = "") |>
  add_equation("pi = (p / p[-1]) - 1", desc = "") |>
  
  # Households realized outcomes
  add_equation("YDr = FX + WB + rm[-1] * M2h[-1] + rb[-1] * Bhh[-1] + BLh[-1]", desc = "") |>
  add_equation("CG = (pbl - pbl[-1]) * BLh[-1]", desc = "") |>
  add_equation("YDhs = YDr + CG", desc = "") |>
  add_equation("FX = FXf + FXb", desc = "") |>
  add_equation("V = V[-1] + YDhs - C", desc = "") |>
  add_equation("Vnc = V - Hhh", desc = "") |>
  add_equation("ydr = YDr/p - pi * (V[-1]/p)", desc = "") |>
  add_equation("ydhs = (YDr - pi * V[-1] + CG) / p", desc = "") |>
  add_equation("v = V/p", desc = "") |>
  
  # Households behavioral
  add_equation("c = alpha0 + alpha1 * ydrE + alpha2 * v[-1]", desc = "") |>
  add_equation("ydrE = epsilon * ydr[-1] + (1 - epsilon) * ydrE[-1]", desc = "") |>
  add_equation("C = p * c", desc = "") |>
  add_equation("YDrE = p * ydrE + pi * (V[-1]/p)", desc = "") |>
  add_equation("VE = V[-1] + (YDrE - C)", desc = "") |>
  add_equation("Hhd = lambdac * C", desc = "") |>
  add_equation("VncE = VE - Hhd", desc = "") |>
  add_equation("ERrbl = rbl", desc = "") |>
  
  # Households' portfolio equations
  add_equation("M2d = VncE * (lambda20 + lambda22 * rm + lambda23 * rb + lambda24 * ERrbl + lambda25 * (YDrE / VncE))", desc = "") |>
  add_equation("Bhd = VncE * (lambda30 + lambda32 * rm + lambda33 * rb + lambda34 * ERrbl + lambda35 * (YDrE / VncE))", desc = "") |>
  add_equation("BLd = (VncE / pbl) * (lambda40 + lambda42 * rm + lambda43 * rb + lambda44 * ERrbl + lambda45 * (YDrE / VncE))", desc = "") |>
  add_equation("M1d = VncE * (lambda10 + lambda12 * rm + lambda13 * rb + lambda14 * ERrbl + lambda15 * (YDrE / VncE))", desc = "") |>
  add_equation("M1d2 = VncE - M2d - Bhd - pbl * BLd", desc = "") |>
  
  # Realized portfolio asset holdings
  add_equation("Hhh = Hhd", desc = "") |>
  add_equation("Bhh = Bhd", desc = "") |>
  add_equation("BLh = BLd", desc = "") |>
  add_equation("M1hN = Vnc - M2d - Bhd - pbl * BLd", desc = "") |>
  add_equation("z1 = as.numeric(M1hN > 0)", desc = "") |>
  add_equation("z2 = 1 - z1", desc = "") |>
  add_equation("M1h = M1hN * z1", desc = "") |>
  add_equation("M2hN = M2d", desc = "") |>
  add_equation("M2h = M2d * z1 + (Vnc - Bhh - pbl * BLd) * z2", desc = "") |>
  
  # Government's equations
  add_equation("TX = S * (tau / (1 + tau))", desc = "") |>
  add_equation("G = p * g", desc = "") |>
  add_equation("PSBR = G + rb[-1] * Bs[-1] + BLs[-1] - (TX + FXcb)", desc = "") |>
  add_equation("Bs = Bs[-1] + PSBR - (BLs - BLs[-1]) * pbl", desc = "") |>
  add_equation("BLs = BLd", desc = "") |>
  add_equation("pbl = 1 / rbl", desc = "") |>
  add_equation("GD = GD[-1] + PSBR", desc = "") |>
  
  # Central bank's equations
  add_equation("Hs = Bcb + As", desc = "") |>
  add_equation("Hbs = Hs - Hhs", desc = "") |>
  add_equation("Bcb = Bs - Bhh - Bbd", desc = "") |>
  add_equation("As = Ad", desc = "") |>
  add_equation("ra = rb", desc = "") |>
  add_equation("FXcb = rb[-1] * Bcb[-1] + ra[-1] * As[-1]", desc = "") |>
  
  # Bank's realized (supply) equations
  add_equation("Hhs = Hhd", desc = "") |>
  add_equation("M1s = M1h", desc = "") |>
  add_equation("M2s = M2h", desc = "") |>
  add_equation("Ls = Ld", desc = "") |>
  add_equation("Hbd = ro1 * M1s + ro2 * M2s", desc = "") |>
  
  # Bank's balance sheet constraints
  add_equation("BbdN = M1s + M2s - Ls - Hbd", desc = "") |>
  add_equation("BLRN = BbdN / (M1s + M2s)", desc = "") |>
  add_equation("Ad = (bot * (M1s + M2s) - BbdN) * z3", desc = "") |>  # z3 instead of z4
  add_equation("z3 = as.numeric(BLRN < bot)", desc = "") |>
  add_equation("Bbd = Ad + M1s + M2s - Ls - Hbd", desc = "") |>
  add_equation("BLR = Bbd / (M1s + M2s)", desc = "") |>
  
  # Determination of interest rates by banks
  add_equation("rm = rm[-1] + zetam * (z4 - z5) + zetab * (rb - rb[-1])", desc = "") |>
  add_equation("z4 = as.numeric(BLRN[-1] < bot)", desc = "") |>
  add_equation("z5 = as.numeric(BLRN[-1] > top)", desc = "") |>
  add_equation("FXb = rl[-1] * Ls[-1] + rb[-1] * Bbd[-1] - rm[-1] * M2s[-1] - ra[-1] * Ad[-1]", desc = "") |>
  add_equation("rl = rl[-1] + zetal * (z6 - z7) + (rb - rb[-1])", desc = "") |>
  add_equation("z6 = as.numeric(BPM < botpm)", desc = "") |>
  add_equation("z7 = as.numeric(BPM > toppm)", desc = "") |>
  add_equation("lM1s = M1s[-1]", desc = "") |>
  add_equation("lM2s = M2s[-1]", desc = "") |>
  add_equation("BPM = (FXb + FXb[-1]) / (lM1s + lM1s[-1] + lM2s + lM2s[-1])", desc = "") |>
  # Inflationary forces
  add_equation("omegaT = exp(Omega0 + Omega1 * log(pr) + Omega2 * log(N / Nfe))", desc = "") |>
  add_equation("W = W[-1] * (1 + Omega3 * (omegaT[-1] - (W[-1] / p[-1])))", desc = "") |>
  add_equation("Y = p * s + UC * (inv - inv[-1])", desc = "") |>
  
  # Hidden equation
  add_equation("Hbd = Hbs", hidden = TRUE)
```

Now, you can simulate the model (in this example, the baseline scenario over 210 periods using the Broyden method).
```{r}
# Simulate model
model_insout <- simulate_scenario(model_insout, scenario = "baseline",
                                  max_iter = 350, periods = 210, tol = 1e-15,
                                  hidden_tol = 0.1, method = "Broyden")
```

After the simulation, plot a few variables to check if the model stabilises.
```{r, out.width="100%", out.height="100%"}
# Steady state
exprs <- c("Y", "y", "s", "inv", "pi", "Bs", "M1s", "M2s", "V", "INV", "FXf", "FXb")  # Bs, M2s, FXb look different
plots <- map(exprs,
             ~ plot_simulation(model = model_insout, scenario = "baseline",
                               from = 1, to = 210, expressions = .x
             )
)
subplot(plots, nrows = 3, shareX = TRUE, titleX = TRUE)
```

*Note:* The above example uses the new pipe operator (`|>`), which requires R 4.1 or later.

The model reaches a steady state from around `t = 100` onward.
```{r}
t0 <- 100
```

### Shock scenario
With `godley` package we can simulate how shocks affect the economy (specifically, how they impact the base scenario).

#### Shock 1
This scenario explores how the model behaves when confronted with an increase in the targeted inventories-to-sales ratio.

To implement this change, define a shock that specifies its size and timing.
```{r}
# Create empty shock and add shock equation
shock_insout <- create_shock() |>
  add_shock(variable = "sigma0", value = 0.4, start = t0+5, end = t0+70, desc = "")
```

Next, link the shock to the baseline model by defining a new scenario.
```{r}
# Create new scenario with this shock
model_insout <- model_insout |>
  add_scenario(name = "sigma0_shock", origin = "baseline", shock = shock_insout)
```

From here, the model is ready to be solved forward to examine the adjustment dynamics.
```{r}
# Simulate shock
model_insout <- simulate_scenario(model_insout, scenario = "sigma0_shock", periods = t0+70, max_iter = 350, tol = 1e-10, method = "Broyden", hidden_tol = 0.1, rhtol = TRUE)
```

To conclude, plot selected variables to see how they evolve following the change in the target ratio.
```{r, out.width="100%"}
# Figure 10.1A Evolution of inventories (and hence bank loans), following an increase in the target inventories to sales ratio.
plot_simulation(model = model_insout, scenario = c("sigma0_shock"),
                from = t0+1, to = t0+70, expressions = c("INV", "Ls"))
```

#### Shock 2
In another example, we look at the effect of an increase in pure government expenditure.
```{r}
# Create empty shock and add shock equation
shock_insout <- create_shock() |>
  add_shock(variable = "g", value = 30, start = t0+5, end = t0+55, desc = "")
```

Start by creating an empty shock and specifying the strength, onset and duration of the intervention.
```{r}
# Create new scenario with this shock
model_insout <- model_insout |>
  add_scenario(name = "g_shock", origin = "baseline", shock = shock_insout)
```

With the scenario in place, run the simulation.
```{r}
# Simulate shock
model_insout <- simulate_scenario(model_insout, scenario = "g_shock", periods = t0+70,
                                  max_iter = 350, tol = 1e-15, method = "Broyden",
                                  hidden_tol = 0.1, rhtol = TRUE)
```

And finally, plot a few key variables to see how the economy responds to the shock.
```{r, out.width="100%"}
# Figure 10.2A Evolution of household real wealth, real disposable income and real consumption, following a one-step permanent increase in real government expenditures
do_plotly(
    m = model_insout, scenario = "g_shock",
    variables = c("c", "v", "ydr"), t0 = t0, end = t0+55,
    title = "Scenario g_shock", y_title = ""
)
```

### References
For a more detailed discussion of the INSOUT model and its economic interpretation, refer to Chapter 10 of *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth*.
