---
title: "Model SIMEX"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model SIMEX}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Please find below guidelines how godley package can be used to create model SIMEX — model with government money and expectations described by *Wynne Godley* and *Marc Lavoie* in Chapter 3 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*

```{r setup}
library(godley)
```

### Base scenario
First you need to create empty SFC model:
```{r}
# Create empty model
model_simex <- create_model(name = "SFC SIMEX")
```

Next you have to add variables:
```{r}
# Add variables
model_simex <- model_simex |>
  add_variable("C_d", desc = "Consumption demand by households") |>
  add_variable("C_s", desc = "Consumption supply") |>
  add_variable("G_s", desc = "Government supply") |>
  add_variable("T_d", desc = "Taxes, demand") |>
  add_variable("T_s", desc = "Taxes, supply") |>
  add_variable("N_d", desc = "Demand for labor") |>
  add_variable("N_s", desc = "Supply of labor") |>
  add_variable("H_h", desc = "Cash money held by households") |>
  add_variable("H_s", desc = "Cash money supplied by the government") |>
  add_variable("H_d", desc = "Cash money demanded by the government") |>
  add_variable("Y", desc = "Income = GDP") |>
  add_variable("Yd", desc = "Disposable income of households") |>
  add_variable("Yd_e", desc = "Expected disposable income of households") |>
  add_variable("alpha1", init = 0.6, desc = "Propensity to consume out of income") |>
  add_variable("alpha2", init = 0.4, desc = "Propensity to consume out of wealth") |>
  add_variable("theta", init = 0.2, desc = "Tax rate") |>
  add_variable("G_d", init = 20, desc = "Government demand") |>
  add_variable("W", init = 1, desc = "Wage rate")
```

You also have to add equations:
```{r}
# Add equations
model_simex <- model_simex |>
  add_equation("C_s = C_d", desc = "Consumption") |>
  add_equation("G_s = G_d") |>
  add_equation("T_s = T_d") |>
  add_equation("N_s = N_d") |>
  add_equation("Yd = W * N_s - T_s") |>
  add_equation("T_d = theta * W * N_s") |>
  add_equation("C_d = alpha1 * Yd_e + alpha2 * H_h[-1]") |>
  add_equation("H_s = G_d - T_d + H_s[-1]") |>
  add_equation("H_h = Yd - C_d + H_h[-1]") |>
  add_equation("Y = C_s + G_s") |>
  add_equation("N_d = Y/W") |>
  add_equation("H_d = Yd_e - C_d + H_h[-1]") |>
  add_equation("Yd_e = Yd[-1]") |>
  add_equation("H_s = H_h", desc = "Money equilibrium", hidden = TRUE)
```

Now you can simulate model (in this example for the next 100 periods using Newton method):
```{r}
# Simulate model
model_simex <- simulate_scenario(model_simex,
  scenario = "baseline", max_iter = 350, periods = 100,
  hidden_tol = 0.1, tol = 1e-08, method = "Newton"
)
```

Based on calculated simulation you can display results for desired set of variables:
```{r, out.width="100%"}
# Plot results
plot_simulation(
  model = model_simex, scenario = c("baseline"), from = 1, to = 50,
  expressions = c("Y", "C_d", "C_s / alpha1")
)
```

Note: In above example new pipe operator was applied. It does not work in older R versions (before R 4.1). 

### Shock scenario
You can simulate how a shock will affect the base scenario. In this example you will implement an increased government expenditures. \
\
First you need to create empty shock:
```{r}
# Create empty shock
shock_simex <- create_shock()
```

Next you need to add appropriate equation:
```{r}
# Add shock equation with increased government expenditures
shock_simex <- add_shock(shock_simex,
  variable = "G_d",
  value = 25,
  desc = "Increase in government expenditures", 
  start = 5, end = 50
)
```

Then you need to create new shock scenario:
```{r}
# Create new scenario with this shock
model_simex <- add_scenario(model_simex,
  name = "expansion", origin = "baseline",
  origin_start = 1,
  origin_end = 100, 
  shock = shock_simex
)
```

Now you can simulate new shock scenario:
```{r}
# Simulate shock
model_simex <- simulate_scenario(model_simex,
  scenario = "expansion", max_iter = 350, periods = 100,
  hidden_tol = 0.1, tol = 1e-08, method = "Newton"
)
```

You can also display results from the shock scenario:
```{r, out.width="100%"}
# Plot results
plot_simulation(
  model = model_simex, scenario = c("expansion"), from = 1, to = 50,
  expressions = c("Y", "C_d", "C_s / alpha1")
)
```

### References
More information about model SIMEX can be found in Chapter 3 in *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*
