---
title: "Model GROWTH"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model GROWTH}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include = FALSE}
plotly_style <- function(fig, title = NULL) {
  fig %>%
    plotly::layout(
      title = list(
        text = title %||% "",
        y = 0.95, x = 0.5,
        xanchor = "center", yanchor = "top"
      ),
      margin = list(t = 60, b = 70, l = 60, r = 120),
      hovermode = "spikers",
      spikedistance = 1000,
      xaxis = list(
        title = "time"
      ),
      yaxis = list(
        title = ""
      ),
      showlegend = TRUE,
      legend = list(
        xanchor = "left",
        x = 1.02,
        y = 0.95
      ),
      font = list(
        family = "Arial",
        size = 13
      )
    )
}

do_plotly <- function(m, variables, plot = NULL,
                      title = NULL,
                      y_title = NULL) {

  variables <- unique(c("time", variables))

  m1 <- m %>%
    dplyr::mutate(
      uk      = Y / K,
      Bsk     = Bs / K,
      VK      = V / K,
      wi      = (W - lag(W)) / lag(W),
      DefY    = PSBR / Y,
      GDY     = GD / Y,
      LIN     = Lfd / IN,
      GRy     = (Yk / lag(Yk)) - 1,
      LhYDr   = Lhd / YDr,
      FUfInv  = FUf / INV,
      EqFMA   = (Pe * Ekd) / Vfma,
      GRFfk   = -1 + (Ff / P) / (lag(Ff) / lag(P)),
      GRPek   = -1 + (Pe / P) / (lag(Pe) / lag(P))
    ) %>%
    tidyr::pivot_longer(cols = -time, names_to = "name", values_to = "value")

  if (is.null(plot)) {
    m2 <- m1 %>% dplyr::filter(name %in% setdiff(variables, "time"))
  } else {
    m2 <- m1 %>% dplyr::filter(name %in% plot)
  }

  fig <- plotly::plot_ly()

  for (v in unique(m2$name)) {
    df <- dplyr::filter(m2, name == v)

    fig <- plotly::add_trace(
      fig,
      data = df,
      x = ~time,
      y = ~value,
      name = v,
      type = "scatter",
      mode = "lines+markers",
      line = list(width = 2),
      marker = list(size = 3),
      hovertemplate = paste(
        "<b>", v, "</b>",
        "<br>time=%{x}",
        "<br>value=%{y}<extra></extra>"
      )
    )
  }

  fig %>%
    plotly::layout(
      yaxis = list(title = y_title %||% "")
    ) %>%
    plotly_style(title)
}


`%||%` <- function(a, b) if (!is.null(a) && length(a) > 0) a else b

do_cbind <- function(m, scenario, variables) {
  variables <- unique(c(variables, "time"))
  
  merge(
    m[[scenario]][["result"]],
    dplyr::select(m[["baseline"]][["result"]], dplyr::all_of(variables)) %>%
      setNames(paste0(names(.), "_bl")),
    by.x = "time",
    by.y = "time_bl"
  )
}

do_cplotly <- function(m, scenario, variables, plot = NULL, from, to,
                       title = NULL,
                       y_title = "scenario / baseline") {

  vars_bl <- paste0(variables, "_bl")
  ntbl <- do_cbind(m, scenario, variables)

  for (i in seq_along(variables)) {
    v  <- variables[[i]]
    vb <- vars_bl[[i]]
    ntbl[[v]] <- ntbl[[v]] / ntbl[[vb]]
  }

  ntbl %>%
    dplyr::filter(time >= from & time <= to) %>%
    do_plotly(
      variables = variables,
      plot = plot,
      title = title %||% paste("Scenario", scenario),
      y_title = y_title
    )
}

```

This guide explains how to use the `godley` package to create a Growth model prototype — by far the most ambitious — with a model of the financial system comprising banks, loans, credit money and equities, together with a model of inflation, as described by *Wynne Godley* and *Marc Lavoie* in Chapter 11 of *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*

```{r setup}
library(godley)
```

### Base scenario
Start by creating an empty SFC model:
```{r}
# Create empty model
model_growth <- create_model(name = "SFC GROWTH")
```

Next, declare all relevant variables. You can also set initial values for all variables, some of them (as in this example), or none at all.
```{r}
# Add variables
model_growth <- model_growth |>
  add_variable("alpha1", init = 0.75) |>
  add_variable("alpha2", init = 0.064) |>
  add_variable("beta", init = 0.5) |>
  add_variable("betab", init = 0.4) |>
  add_variable("gamma", init = 0.15) |>
  add_variable("gamma0", init = 0.00122) |>
  add_variable("gammar", init = 0.1) |>
  add_variable("gammau", init = 0.05) |>
  add_variable("delta", init = 0.10667) |>
  add_variable("deltarep", init = 0.1) |>
  add_variable("eps", init = 0.5) |>
  add_variable("eps2", init = 0.8) |>
  add_variable("epsb", init = 0.25) |>
  add_variable("epsrb", init = 0.9) |>
  add_variable("eta0", init = 0.07416) |>
  add_variable("etan", init = 0.6) |>
  add_variable("etar", init = 0.4) |>
  add_variable("theta", init = 0.22844) |>
  add_variable("lambda20", init = 0.25) |>
  add_variable("lambda21", init = 2.2) |>
  add_variable("lambda22", init = 6.6) |>
  add_variable("lambda23", init = 2.2) |>
  add_variable("lambda24", init = 2.2) |>
  add_variable("lambda25", init = 0.1) |>
  add_variable("lambda30", init = -0.04341) |>
  add_variable("lambda31", init = 2.2) |>
  add_variable("lambda32", init = 2.2) |>
  add_variable("lambda33", init = 6.6) |>
  add_variable("lambda34", init = 2.2) |>
  add_variable("lambda35", init = 0.1) |>
  add_variable("lambda40", init = 0.67132) |>
  add_variable("lambda41", init = 2.2) |>
  add_variable("lambda42", init = 2.2) |>
  add_variable("lambda43", init = 2.2) |>
  add_variable("lambda44", init = 6.6) |>
  add_variable("lambda45", init = 0.1) |>
  add_variable("lambdab", init = 0.0153) |>
  add_variable("lambdac", init = 0.05) |>
  add_variable("xim1", init = 0.0008) |>
  add_variable("xim2", init = 0.0007) |>
  add_variable("ro", init = 0.05) |>
  add_variable("sigman", init = 0.1666) |>
  add_variable("sigmat", init = 0.2) |>
  add_variable("psid", init = 0.15255) |>
  add_variable("psiu", init = 0.92) |>
  add_variable("omega0", init = -0.20594) |>
  add_variable("omega1", init = 1) |>
  add_variable("omega2", init = 2) |>
  add_variable("omega3", init = 0.45621) |>
  add_variable("ADDbl", init = 0.02) |>
  add_variable("BANDt", init = 0.01) |>
  add_variable("BANDb", init = 0.01) |>
  add_variable("bot", init = 0.05) |>
  add_variable("GRg", init = 0.03) |>
  add_variable("GRpr", init = 0.03) |>
  add_variable("Nfe", init = 87.181) |>
  add_variable("NCAR", init = 0.1) |>
  add_variable("NPLk", init = 0.02) |>
  add_variable("Rbbar", init = 0.035) |>
  add_variable("Rln", init = 0.07) |>
  add_variable("RA", init = 0) |>
  add_variable("top", init = 0.12) |>
  add_variable("sigmase", init = 0.16667) |>
  add_variable("eta", init = 0.04918) |>
  add_variable("phi", init = 0.26417) |>
  add_variable("phit", init = 0.26417) |>
  add_variable("ADDl", init = 0.04592) |>
  add_variable("BLR", init = 0.1091) |>
  add_variable("BUR", init = 0.06324) |>
  add_variable("Ck", init = 7334240) |>
  add_variable("CAR", init = 0.09245) |>
  add_variable("CONS", init = 52603100) |>
  add_variable("ER", init = 1) |>
  add_variable("Fb", init = 1744130) |>
  add_variable("Fbt", init = 1744140) |>
  add_variable("Ff", init = 18081100) |>
  add_variable("Fft", init = 18013600) |>
  add_variable("FDb", init = 1325090) |>
  add_variable("FDf", init = 2670970) |>
  add_variable("FUb", init = 419039) |>
  add_variable("FUf", init = 15153800) |>
  add_variable("FUft", init = 15066200) |>
  add_variable("G", init = 16755600) |>
  add_variable("Gk", init = 2336160) |>
  add_variable("GL", init = 2775900) |>
  add_variable("GRk", init = 0.03001) |>
  add_variable("INV", init = 16911600) |>
  add_variable("Ik", init = 2357910) |>
  add_variable("N", init = 87.181) |>
  add_variable("Nt", init = 87.181) |>
  add_variable("NHUC", init = 5.6735) |>
  add_variable("NL", init = 683593) |>
  add_variable("NLk", init = 95311) |>
  add_variable("NPL", init = 309158) |>
  add_variable("NPLke", init = 0.02) |>
  add_variable("NUC", init = 5.6106) |>
  add_variable("omegat", init = 112852) |>
  add_variable("P", init = 7.1723) |>
  add_variable("Pbl", init = 18.182) |>
  add_variable("Pe", init = 17937) |>
  add_variable("PE", init = 5.07185) |>
  add_variable("PI", init = 0.0026) |>
  add_variable("PR", init = 138659) |>
  add_variable("PSBR", init = 1894780) |>
  add_variable("Q", init = 0.77443) |>
  add_variable("Rb", init = 0.035) |>
  add_variable("Rbl", init = 0.055) |>
  add_variable("Rk", init = 0.03008) |>
  add_variable("Rl", init = 0.06522) |>
  add_variable("Rm", init = 0.0193) |>
  add_variable("REP", init = 2092310) |>
  add_variable("RRl", init = 0.06246) |>
  add_variable("S", init = 86270300) |>
  add_variable("Sk", init = 12028300) |>
  add_variable("Ske", init = 12028300) |>
  add_variable("TX", init = 17024100) |>
  add_variable("U", init = 0.70073) |>
  add_variable("UC", init = 5.6106) |>
  add_variable("W", init = 777968) |>
  add_variable("WB", init = 67824000) |>
  add_variable("Y", init = 86607700) |>  
  add_variable("Yk", init = 12088400) |>
  add_variable("YDr", init = 56446400) |>
  add_variable("YDkr", init = 7813270) |>
  add_variable("YDkre", init = 7813290) |>
  add_variable("YP", init = 73158700) |>
  add_variable("z1a", init = 0) |>
  add_variable("z1b", init = 0) |>
  add_variable("z2a", init = 0) |>  
  add_variable("z2b", init = 0) |>
  add_variable("Bbd", init = 4389790) |>
  add_variable("Bbs", init = 4389790) |>
  add_variable("Bcbd", init = 4655690) |>
  add_variable("Bcbs", init = 4655690) |>
  add_variable("Bhd", init = 33439320) |>
  add_variable("Bhs", init = 33439320) |>
  add_variable("Bs", init = 42484800) |>
  add_variable("BLd", init = 840742) |>
  add_variable("BLs", init = 840742) |>
  add_variable("GD", init = 57728700) |>
  add_variable("Ekd", init = 5112.6001) |>
  add_variable("Eks", init = 5112.6001) |>
  add_variable("Hbd", init = 2025540) |>
  add_variable("Hbs", init = 2025540) |>
  add_variable("Hhd", init = 2630150) |>
  add_variable("Hhs", init = 2630150) |>
  add_variable("Hs", init = 4655690) |>
  add_variable("IN", init = 11585400) |>
  add_variable("INk", init = 2064890) |>
  add_variable("INke", init = 2405660) |>
  add_variable("INkt", init = 2064890) |>
  add_variable("K", init = 127486471) |>
  add_variable("Kk", init = 17774838) |>
  add_variable("Lfd", init = 15962900) |>
  add_variable("Lfs", init = 15962900) |>
  add_variable("Lhd", init = 21606600) |>
  add_variable("Lhs", init = 21606600) |>
  add_variable("Ls", init = 37569500) |>
  add_variable("Mh", init = 40510800) |>
  add_variable("Ms", init = 40510800) |>
  add_variable("OFb", init = 3474030) |>
  add_variable("OFbe", init = 3474030) |>
  add_variable("OFbt", init = 3638100) |>
  add_variable("V", init = 165438779) |>
  add_variable("Vfma", init = 159334599) |>
  add_variable("Vk", init = 23066350) |>
  add_variable("Vf", init = 31361792) |>
  add_variable("z3") |>
  add_variable("z4") |>
  add_variable("z5") |>
  add_variable("z3a") |>
  add_variable("z3b") |>
  add_variable("HCe") |>
  add_variable("YDhs") |>
  add_variable("CG") |>
  add_variable("VfmaA") |>
  add_variable("Fcb") |>
  add_variable("FUbt")
```

Then, add the equations — this is where you define how the variables relate to each other.
```{r}
# Add equations
model_growth <- model_growth |>
  add_equation("Yk = Ske + INke - INk[-1]", desc = "11.1 : Real output") |>
  add_equation("Ske = beta*Sk + (1-beta)*Sk[-1]*(1 + (GRpr + RA))", desc = "11.2 : Expected real sales") |>
  add_equation("INke = INk[-1] + gamma*(INkt - INk[-1])", desc = "11.3 : Long-run inventory target") |>
  add_equation("INkt = sigmat*Ske", desc = "11.4 : Short-run inventory target") |>
  add_equation("INk = INk[-1] + Yk - Sk - NPL/UC", desc = "11.5 : Actual real inventories") |>
  add_equation("Kk = Kk[-1]*(1 + GRk)", desc = "11.6 : Real capital stock") |>
  add_equation("GRk = gamma0 + gammau*U[-1] - gammar*RRl", desc = "11.7 : Growth of real capital stock") |>
  add_equation("U = Yk/Kk[-1]", desc = "11.8 : Capital utilization proxy") |>
  add_equation("RRl = ((1 + Rl)/(1 + PI)) - 1", desc = "11.9 : Real interest rate on loans") |>
  add_equation("PI = (P - P[-1])/P[-1]", desc = "11.10 : Rate of price inflation") |>
  add_equation("Ik = (Kk - Kk[-1]) + delta*Kk[-1]", desc = "11.11 : Real gross investment") |>
  
  # Firms equations
  add_equation("Sk = Ck + Gk + Ik", desc = "11.12 : Actual real sales") |>
  add_equation("S = Sk*P", desc = "11.13 : Value of realized sales") |>
  add_equation("IN = INk*UC", desc = "11.14 : Inventories valued at current cost") |>
  add_equation("INV = Ik*P", desc = "11.15 : Nominal gross investment") |>
  add_equation("K = Kk*P", desc = "11.16 : Nomincal value of fixed capital") |>
  add_equation("Y = Sk*P + (INk - INk[-1])*UC", desc = "11.17 : Nomincal GDP") |>
  add_equation("omegat = exp(omega0 + omega1*log(PR) + omega2*log(ER + z3*(1 - ER) - z4*BANDt + z5*BANDb))", desc = "11.18 : Real wage aspirations") |>
  add_equation("ER = N[-1]/Nfe[-1]", desc = "11.19 : Employment rate") |>
  # 11.20 : Switch variables
  add_equation("z3a = (ER > (1 - BANDb))", desc = "") |>
  add_equation("z3b = 1 - as.numeric(ER > (1 + BANDt))", desc = "") |>
  add_equation("z3 =  z3a * z3b", desc = "") |>
  add_equation("z4  = (ER >  (1 + BANDt))", desc = "") |>
  add_equation("z5  = (ER <  (1 - BANDb))", desc = "") |>
  add_equation("W = W[-1] + omega3*(omegat*P[-1] - W[-1])", desc = "11.21 : Nominal wage") |>
  add_equation("PR = PR[-1]*(1 + GRpr)", desc = "11.22 : Labor productivity") |>
  add_equation("Nt = Yk/PR", desc = "11.23 : Desired employment") |>
  add_equation("N = N[-1] + etan*(Nt - N[-1])", desc = "11.24 : Actual employment") |>
  add_equation("WB = N*W", desc = "11.25 : Nominal wage bill") |>
  add_equation("UC = WB/Yk", desc = "11.26 : Actual unit cost") |>
  add_equation("NUC = W/PR", desc = "11.27 : Normal unit cost") |>
  add_equation("NHUC = (1 - sigman)*NUC + sigman*(1 + Rln[-1])*NUC[-1]", desc = "11.28 : Normal historic unit cost") |>
  add_equation("P = (1 + phi)*NHUC", desc = "11.29 : Normal-cost pricing") |>
  add_equation("phi = phi[-1] + eps2*(phit[-1] - phi[-1])", desc = "11.30 : Actual mark-up") |>
  add_equation("phit = (FUft + FDf + Rl[-1]*(Lfd[-1] - IN[-1])) / ((1 - sigmase)*Ske*UC + (1 + Rl[-1])*sigmase*Ske*UC[-1])", desc = "11.31 : Ideal mark-up") |>
  add_equation("HCe = (1 - sigmase)*Ske*UC + (1 + Rl[-1])*sigmase*Ske*UC[-1]", desc = "11.32 : Expected historical costs") |>
  add_equation("sigmase = INk[-1]/Ske", desc = "11.33 : Opening inventories to expected sales ratio") |>
  add_equation("Fft = FUft + FDf + Rl[-1]*(Lfd[-1] - IN[-1])", desc = "11.34 : Planned entrepeneurial profits of firms") |>
  add_equation("FUft = psiu*INV[-1]", desc = "11.35 : Planned retained earnings of firms") |>
  add_equation("FDf = psid*Ff[-1]", desc = "11.36 : Dividends of firms") |>
  add_equation("Ff = S - WB + (IN - IN[-1]) - Rl[-1]*IN[-1]", desc = "11.37 : Realized entrepeneurial profits") |>
  add_equation("FUf = Ff - FDf - Rl[-1]*(Lfd[-1] - IN[-1]) + Rl[-1]*NPL", desc = "11.38 : Retained earnings of firms") |>
  add_equation("Lfd = Lfd[-1] + INV + (IN - IN[-1]) - FUf - (Eks - Eks[-1])*Pe - NPL", desc = "11.39 : Demand for loans by firms") |>
  add_equation("NPL = NPLk * Lfs[-1]", desc = "11.40 : Defaulted loans") |>
  add_equation("Eks = Eks[-1] + ((1 - psiu)*INV[-1])/Pe", desc = "11.41 : Supply of equities issued by firms") |>
  add_equation("Rk = FDf/(Pe[-1]*Ekd[-1])", desc = "11.42 : Dividend yield of firms") |>
  add_equation("PE = Pe/(Ff/Eks[-1])", desc = "11.43 : Price earnings ratio") |>
  add_equation("Q = (Eks*Pe + Lfd)/(K + IN)", desc = "11.44 : Tobins Q ratio") |>
  
  # Households equations
  add_equation("YP = WB + FDf + FDb + Rm[-1]*Mh[-1] + Rb[-1]*Bhd[-1] + BLs[-1]", desc = "") |>
  add_equation("TX = theta*YP", desc = "11.46 : Income taxes") |>
  add_equation("YDr = YP - TX - Rl[-1]*Lhd[-1]", desc = "11.47 : Regular disposable income") |>
  add_equation("YDhs = YDr + CG", desc = "11.48 : Haig-Simons disposable income") |>
  add_equation("CG = (Pbl - Pbl[-1])*BLd[-1] + (Pe - Pe[-1])*Ekd[-1] + (OFb - OFb[-1])", desc = "11.49 : Capital gains") |>
  add_equation("V = V[-1] + YDr - CONS + (Pbl - Pbl[-1])*BLd[-1] + (Pe - Pe[-1])*Ekd[-1] + (OFb - OFb[-1])", desc = "11.50 : Wealth") |>
  add_equation("Vk = V/P", desc = "11.51 : Real stock of wealth") |>
  add_equation("CONS = Ck*P", desc = "11.52 : Consumption") |>
  add_equation("Ck = alpha1*(YDkre + NLk) + alpha2*Vk[-1]", desc = "11.53 : Real consumption") |>
  add_equation("YDkre = eps*YDkr + (1 - eps)*(YDkr[-1]*(1 + GRpr))", desc = "11.54 : Expected real regular disposable income") |>
  add_equation("YDkr = YDr/P - ((P - P[-1]) * Vk[-1])/P", desc = "11.55 : Real regular disposable income") |>
  add_equation("GL = eta*YDr", desc = "11.56 : Gross amount of new personal loans") |>
  add_equation("eta = eta0 - etar*RRl", desc = "11.57 : New loans to personal income ratio") |>
  add_equation("NL = GL - REP", desc = "11.58 : Net amount of new personal loans") |>
  add_equation("REP = deltarep*Lhd[-1]", desc = "11.59 : Personal loans repayments") |>
  add_equation("Lhd = Lhd[-1] + GL - REP", desc = "11.60 : Demand for personal loans") |>
  add_equation("NLk = NL/P", desc = "11.61 : Real amount of new personal loans") |>
  add_equation("BUR = (REP + Rl[-1] * Lhd[-1]) / YDr[-1]", desc = "11.62 : Burden of personal debt") |>
  
  # Households equations - portfolio decisions
  add_equation("Bhd = Vfma[-1]*(lambda20 + lambda22*Rb[-1] - lambda21*Rm[-1] - lambda24*Rk[-1] - lambda23*Rbl[-1] - lambda25*(YDr/V))", desc = "11.64 : Demand for bills") |>
  add_equation("BLd = Vfma[-1]*(lambda30 - lambda32*Rb[-1] - lambda31*Rm[-1] - lambda34*Rk[-1] + lambda33*Rbl[-1] - lambda35*(YDr/V))/Pbl", desc = "11.65 : Demand for bonds") |>
  add_equation("Pe = Vfma[-1]*(lambda40 - lambda42*Rb[-1] - lambda41*Rm[-1] + lambda44*Rk[-1] - lambda43*Rbl[-1] - lambda45*(YDr/V))/Ekd", desc = "11.66 : Demand for equities - normalized to get the price of equitities") |>
  add_equation("Mh = Vfma - Bhd - Pe*Ekd - Pbl*BLd + Lhd", desc = "11.67 : Money deposits - as a residual") |>
  add_equation("Vfma = V - Hhd - OFb", desc = "11.68 : Investible wealth") |>
  add_equation("VfmaA = Mh + Bhd + Pbl * BLd + Pe * Ekd", desc = "") |>
  add_equation("Hhd = lambdac*CONS", desc = "11.69 : Households demand for cash") |>
  add_equation("Ekd = Eks", desc = "11.70 : Stock market equilibrium") |>
  
  # Governments equations
  add_equation("G = Gk*P", desc = "11.71 : Pure government expenditures") |>
  add_equation("Gk = Gk[-1]*(1 + GRg)", desc = "11.72 : Real government expenditures") |>
  add_equation("PSBR = G + BLs[-1] + Rb[-1]*(Bbs[-1] + Bhs[-1]) - TX", desc = "11.73 : Government deficit") |>
  add_equation("Bs = Bs[-1] + G - TX - (BLs - BLs[-1])*Pbl + Rb[-1]*(Bhs[-1] + Bbs[-1]) + BLs[-1]", desc = "11.74 : New issues of bills") |>
  add_equation("GD = Bbs + Bhs + BLs*Pbl + Hs", desc = "11.75 : Government debt") |>
  
  # The Central banks equations
  add_equation("Fcb = Rb[-1]*Bcbd[-1]", desc = "11.76 : Central bank profits") |>
  add_equation("BLs = BLd", desc = "11.77 : Bonds are supplied on demand") |>
  add_equation("Bhs = Bhd", desc = "11.78 : Household bills supplied on demand") |>
  add_equation("Hhs = Hhd", desc = "11.79 : Cash supplied on demand") |>
  add_equation("Hbs = Hbd", desc = "11.80 : Reserves supplied on demand") |>
  add_equation("Hs = Hbs + Hhs", desc = "11.81 : Total supply of cash") |>
  add_equation("Bcbd = Hs", desc = "11.82 : Central bankd ") |>
  add_equation("Bcbs = Bcbd", desc = "11.83 : Supply of bills to Central bank") |>
  add_equation("Rb = Rbbar", desc = "11.84 : Interest rate on bills set exogenously") |>
  add_equation("Rbl = Rb + ADDbl", desc = "11.85 : Long term interest rate") |>
  add_equation("Pbl = 1/Rbl", desc = "11.86 : Price of long-term bonds") |>
  
  # Commercial Banks equations
  add_equation("Ms = Mh", desc = "11.87 : Bank deposits supplied on demand") |>
  add_equation("Lfs = Lfd", desc = "11.88 : Loans to firms supplied on demand") |>
  add_equation("Lhs = Lhd", desc = "11.89 : Personal loans supplied on demand") |>
  add_equation("Hbd = ro*Ms", desc = "11.90 Reserve requirements of banks") |>
  add_equation("Bbs = Bbs[-1] + (Bs - Bs[-1]) - (Bhs - Bhs[-1]) - (Bcbs - Bcbs[-1])", desc = "11.91 : Bills supplied to banks") |>
  add_equation("Bbd = Ms + OFb - Lfs - Lhs - Hbd", desc = "11.92 : Balance sheet constraint of banks") |>
  add_equation("BLR = Bbd/Ms", desc = "11.93 : Bank liquidity ratio") |>
  add_equation("Rm = Rm[-1] + z1a*xim1 + z1b*xim2 - z2a*xim1 - z2b*xim2", desc = "11.94 : Deposit interest rate") |>
  # 11.95-97 : Mechanism for determining changes to the interest rate on deposits,
  # i.e., logical functions dependent on whether the bank liquidity ratio is within its bot and top range.
  add_equation("z2a = (BLR[-1] >  (top + 0.05))", desc = "") |>
  add_equation("z2b = (BLR[-1] >  top)", desc = "") |>
  add_equation("z1a = 1 - as.numeric(BLR[-1] > bot)", desc = "") |>
  add_equation("z1b = 1 - as.numeric(BLR[-1] > (bot - 0.05))", desc = "") |>
  
  # Commercial banks equations
  add_equation("Rl = Rm + ADDl", desc = "11.98 : Loan interest rate") |>
  add_equation("OFbt = NCAR*(Lfs[-1] + Lhs[-1])", desc = "11.99 : Long-run own funds target") |>
  add_equation("OFbe = OFb[-1] + betab*(OFbt - OFb[-1])", desc = "11.100 : Short-run own funds target") |>
  add_equation("FUbt = OFbe - OFb[-1] + NPLke*Lfs[-1]", desc = "11.101 : Target retained earnings of banks") |>
  add_equation("NPLke = epsb*NPLke[-1] + (1 - epsb)*NPLk[-1]", desc = "11.102 : Expected proportion of non-performaing loans") |>
  add_equation("FDb = Fb - FUb", desc = "11.103 : Dividends of banks") |>
  add_equation("Fbt = lambdab*Y[-1] + (OFbe - OFb[-1] + NPLke*Lfs[-1])", desc = "11.104 : Target profits of banks") |>
  add_equation("Fb = Rl[-1]*(Lfs[-1] + Lhs[-1] - NPL) + Rb[-1]*Bbd[-1] - Rm[-1]*Ms[-1]", desc = "11.105 : Actual profits of banks") |>
  add_equation("ADDl = (Fbt - Rb[-1]*Bbd[-1] + Rm[-1]*(Ms[-1] - (1 - NPLke)*Lfs[-1] - Lhs[-1]))/((1 - NPLke)*Lfs[-1] + Lhs[-1])", desc = "11.106 : Lending mark-up over deposit rate") |>
  add_equation("FUb = Fb - lambdab*Y[-1]", desc = "11.107 : Actual retained earnings") |>
  add_equation("OFb = OFb[-1] + FUb - NPL", desc = "11.108 : Own funds of banks") |>
  add_equation("CAR = OFb/(Lfs + Lhs)", desc = "11.109 : Actual capital adequacy ratio") |>
  add_equation("Vf = IN + K - Lfd - Ekd * Pe", desc = "Firm's wealth (memo for matrices)") |>
  add_equation("Ls = Lfs + Lhs", desc = "Loans supply (memo for matrices)") |>
  
  # Hidden equation
  add_equation("Bbs = Bbd", desc = "11.110A : Redundant equation", hidden = TRUE)

```

Now, you can simulate the model (in this example, the baseline scenario over 500 periods using the Broyden method).
```{r}
# Simulate model
model_growth <- simulate_scenario(model_growth, scenario = "baseline", 
                                  max_iter = 350, periods = 500, tol = 1e-15,
                                  hidden_tol = 1e-6, rhtol = TRUE, method = "Broyden")
```

With the simulation estimated, plot a few variables to quickly check whether the model settles into a steady state.
```{r, out.width="100%"}
# Plot results
exprs <- c("Bsk = Bs / K", "VK = V / K", "GRk", "PI")
plots <- purrr::map(exprs,
             ~ plot_simulation(model = model_growth, scenario = "baseline",
                               from = 1, to = 350, expressions = .x
             )
)
plotly::subplot(plots, nrows = 2, shareX = TRUE, titleX = TRUE)

# A steady state from about t = 300 onward.
t0 <- 300
```

*Note:* The above example uses the new pipe operator (`|>`), which requires R 4.1 or later.

### Shock scenario
With `godley` package we can simulate how shocks affect the economy (specifically, how they impact the base scenario).

#### Shock 1
In this example, we observe the effect of an autonomous increase in the target real wage.

First, initialize an empty shock object and define the shock path for `omega0`.
```{r}
# Create empty shock and add shock equation
shock_growth <- create_shock() |>
  add_shock(variable = "omega0", value = -0.1, start = 5, end = 150, desc = "")
```

Then, integrate the shock into the model by creating a new scenario:
```{r}
# Create new scenario with this shock
model_growth <- model_growth |>
  add_scenario(name = "omega0_shock", origin = "baseline", shock = shock_growth)
```

Simulate the scenario with the shock applied:
```{r}
# Simulate shock
model_growth <- simulate_scenario(model_growth, scenario = "omega0_shock", periods = 150, 
                                  max_iter = 350, tol = 1e-10, method = "Broyden",
                                  hidden_tol = 1e-6, rhtol = TRUE)
```

Finally, plot the shock scenario against the baseline to see what changed.
```{r, out.width="100%"}
# Figure 11.2A: Evolution of wage inflation and price inflation, following an autonomous increase in the target real wage
plot_simulation(model = model_growth, scenario = c("baseline", "omega0_shock"),
                from = 1, to = 80, expressions = c("PI", "wi = (W - dplyr::lag(W)) / dplyr::lag(W)"))
```

```{r, out.width="100%"}
# Figure 11.2B: Evolution of real gross fixed investment, real output and real consumption, all relative to the base line solution, following an autonomous increase in the target real wage
do_cplotly(m = model_growth, scenario = "omega0_shock",
         from = 1, to = 80, variables = c("Ik", "Ck", "Yk"))
```

#### Shock 2
In the second example, we introduce an increase in the growth rate of pure government expenditures for a single period only.

To start, create an empty shock, define the one-time change in `GRg`, and run the scenario.
```{r}
# Create empty shock and add shock equation
shock_growth <- create_shock() |>
  add_shock(variable = "GRg", value = 0.035, start = t0+10, end = t0+11, desc = "")

model_growth <- model_growth |>
  add_scenario(name = "GRg_onetime_shock", origin = "baseline", shock = shock_growth)

# Simulate shock
model_growth <- simulate_scenario(model_growth, scenario = "GRg_onetime_shock", periods = t0+150,
                                  max_iter = 350, tol = 1e-10, method = "Broyden",
                                  hidden_tol = 1e-6, rhtol = TRUE)
```

Once you have the simulation, you can plot the results:
```{r, out.width="100%"}
# Figure 11.3A: Evolution of real output and real pure government expenditure, relative to the base line solution, following an increase in the rate of growth of real pure government expenditures for only one year
do_cplotly(m = model_growth, scenario = "GRg_onetime_shock",
         variables = c("Yk", "Gk"), plot = c("Yk", "Gk"),
         from = t0, to = t0+80)
```

```{r, out.width="100%"}
# Figure 11.3C: Evolution of the government deficit to GDP ratio and of the government debt to GDP ratio, relative to the base line solution, following an increase in the rate of growth of real pure government expenditures for only one year
do_cplotly(m = model_growth, scenario = "GRg_onetime_shock",
         variables = c("PSBR", "Y", "GD"), plot = c("DefY", "GDY"),
         from = t0, to = t0+80)
```

```{r, out.width="100%"}
# Figure 11.3D: Evolution of the bank liquidity ratio and of the loans to inventories ratio of firms, relative to the base line solution, following an increase in the rate of growth of real pure government expenditures for only one year
do_cplotly(m = model_growth, scenario = "GRg_onetime_shock",
         variables = c("Lfd", "IN", "BLR"), plot = c("LIN", "BLR"),
         from = t0, to = t0+80)

```


#### Shock 3
Finally, we consider a permanent increase in the growth rate of pure government expenditures.
 
Again, create a shock and specify the new path for `GRg`:
```{r}
# Create empty shock and add shock equation
shock_growth <- create_shock() |>
  add_shock(variable = "GRg", value = 0.035, start = t0+10, end = t0+150, desc = "")
```

Add the shock as a new scenario based on the baseline:
```{r}
# Create new scenario with this shock
model_growth <- model_growth |>
  add_scenario(name = "GRg_permanent_shock", origin = "baseline", shock = shock_growth)
```

Run the simulation for this scenario:
```{r}
# Simulate shock
model_growth <- simulate_scenario(model_growth, scenario = "GRg_permanent_shock", periods = t0+150, max_iter = 350, tol = 1e-10, method = "Broyden", hidden_tol = 1e-6, rhtol = TRUE)
```

Plot the long-run effects, for example on output growth and capital accumulation.
```{r, out.width="100%"}
# Figure 11.4B: Evolution of the real rate of capital accumulation and of the growth rate of real output, with the growth rate of real pure government expenditures being forever higher than in the base line solution
plot_simulation(model = model_growth, scenario = "GRg_permanent_shock",
                from = t0, to = t0+80, expressions = c("GRg", "GRk", "GRy = (Yk / dplyr::lag(Yk)) - 1")
)
```

### Model GROWTH2: Making monetary policy endogenous
In **GROWTH2**, we make monetary policy endogenous:
the central bank actively conducts monetary policy by adjusting the nominal bill rate so that the real rate moves toward a target that depends on inflation.

To begin, create an empty SFC model:
```{r}
# Create empty model
model_growth2 <- create_model(name = "SFC GROWTH2")
```

Next, declare the variables (same set as in **GROWTH**, plus the extra policy-rate variables).
```{r}
# Add variables
model_growth2 <- model_growth2 |>
  add_variable("alpha1", init = 0.75) |>
  add_variable("alpha2", init = 0.064) |>
  add_variable("beta", init = 0.5) |>
  add_variable("betab", init = 0.4) |>
  add_variable("gamma", init = 0.15) |>
  add_variable("gamma0", init = 0.00122) |>
  add_variable("gammar", init = 0.1) |>
  add_variable("gammau", init = 0.05) |>
  add_variable("delta", init = 0.10667) |>
  add_variable("deltarep", init = 0.1) |>
  add_variable("eps", init = 0.5) |>
  add_variable("eps2", init = 0.8) |>
  add_variable("epsb", init = 0.25) |>
  add_variable("epsrb", init = 0.9) |>
  add_variable("eta0", init = 0.07416) |>
  add_variable("etan", init = 0.6) |>
  add_variable("etar", init = 0.4) |>
  add_variable("theta", init = 0.22844) |>
  add_variable("lambda20", init = 0.25) |>
  add_variable("lambda21", init = 2.2) |>
  add_variable("lambda22", init = 6.6) |>
  add_variable("lambda23", init = 2.2) |>
  add_variable("lambda24", init = 2.2) |>
  add_variable("lambda25", init = 0.1) |>
  add_variable("lambda30", init = -0.04341) |>
  add_variable("lambda31", init = 2.2) |>
  add_variable("lambda32", init = 2.2) |>
  add_variable("lambda33", init = 6.6) |>
  add_variable("lambda34", init = 2.2) |>
  add_variable("lambda35", init = 0.1) |>
  add_variable("lambda40", init = 0.67132) |>
  add_variable("lambda41", init = 2.2) |>
  add_variable("lambda42", init = 2.2) |>
  add_variable("lambda43", init = 2.2) |>
  add_variable("lambda44", init = 6.6) |>
  add_variable("lambda45", init = 0.1) |>
  add_variable("lambdab", init = 0.0153) |>
  add_variable("lambdac", init = 0.05) |>
  add_variable("xim1", init = 0.0008) |>
  add_variable("xim2", init = 0.0007) |>
  add_variable("ro", init = 0.05) |>
  add_variable("sigman", init = 0.1666) |>
  add_variable("sigmat", init = 0.2) |>
  add_variable("psid", init = 0.15255) |>
  add_variable("psiu", init = 0.92) |>
  add_variable("omega0", init = -0.20594) |>
  add_variable("omega1", init = 1) |>
  add_variable("omega2", init = 2) |>
  add_variable("omega3", init = 0.45621) |>
  add_variable("ADDbl", init = 0.02) |>
  add_variable("BANDt", init = 0.01) |>
  add_variable("BANDb", init = 0.01) |>
  add_variable("bot", init = 0.05) |>
  add_variable("GRg", init = 0.03) |>
  add_variable("GRpr", init = 0.03) |>
  add_variable("Nfe", init = 87.181) |>
  add_variable("NCAR", init = 0.1) |>
  add_variable("NPLk", init = 0.02) |>
  add_variable("Rbbar", init = 0.035) |>
  add_variable("Rln", init = 0.07) |>
  add_variable("RA", init = 0) |>
  add_variable("top", init = 0.12) |>
  add_variable("sigmase", init = 0.16667) |>
  add_variable("eta", init = 0.04918) |>
  add_variable("phi", init = 0.26417) |>
  add_variable("phit", init = 0.26417) |>
  add_variable("ADDl", init = 0.04592) |>
  add_variable("BLR", init = 0.1091) |>
  add_variable("BUR", init = 0.06324) |>
  add_variable("Ck", init = 7334240) |>
  add_variable("CAR", init = 0.09245) |>
  add_variable("CONS", init = 52603100) |>
  add_variable("ER", init = 1) |>
  add_variable("Fb", init = 1744130) |>
  add_variable("Fbt", init = 1744140) |>
  add_variable("Ff", init = 18081100) |>
  add_variable("Fft", init = 18013600) |>
  add_variable("FDb", init = 1325090) |>
  add_variable("FDf", init = 2670970) |>
  add_variable("FUb", init = 419039) |>
  add_variable("FUf", init = 15153800) |>
  add_variable("FUft", init = 15066200) |>
  add_variable("G", init = 16755600) |>
  add_variable("Gk", init = 2336160) |>
  add_variable("GL", init = 2775900) |>
  add_variable("GRk", init = 0.03001) |>
  add_variable("INV", init = 16911600) |>
  add_variable("Ik", init = 2357910) |>
  add_variable("N", init = 87.181) |>
  add_variable("Nt", init = 87.181) |>
  add_variable("NHUC", init = 5.6735) |>
  add_variable("NL", init = 683593) |>
  add_variable("NLk", init = 95311) |>
  add_variable("NPL", init = 309158) |>
  add_variable("NPLke", init = 0.02) |>
  add_variable("NUC", init = 5.6106) |>
  add_variable("omegat", init = 112852) |>
  add_variable("P", init = 7.1723) |>
  add_variable("Pbl", init = 18.182) |>
  add_variable("Pe", init = 17937) |>
  add_variable("PE", init = 5.07185) |>
  add_variable("PI", init = 0.0026) |>
  add_variable("PR", init = 138659) |>
  add_variable("PSBR", init = 1894780) |>
  add_variable("Q", init = 0.77443) |>
  add_variable("Rb", init = 0.035) |>
  add_variable("Rbl", init = 0.055) |>
  add_variable("Rk", init = 0.03008) |>
  add_variable("Rl", init = 0.06522) |>
  add_variable("Rm", init = 0.0193) |>
  add_variable("REP", init = 2092310) |>
  add_variable("RRl", init = 0.06246) |>
  add_variable("S", init = 86270300) |>
  add_variable("Sk", init = 12028300) |>
  add_variable("Ske", init = 12028300) |>
  add_variable("TX", init = 17024100) |>
  add_variable("U", init = 0.70073) |>
  add_variable("UC", init = 5.6106) |>
  add_variable("W", init = 777968) |>
  add_variable("WB", init = 67824000) |>
  add_variable("Y", init = 86607700) |>  
  add_variable("Yk", init = 12088400) |>
  add_variable("YDr", init = 56446400) |>
  add_variable("YDkr", init = 7813270) |>
  add_variable("YDkre", init = 7813290) |>
  add_variable("YP", init = 73158700) |>
  add_variable("z1a", init = 0) |>
  add_variable("z1b", init = 0) |>
  add_variable("z2a", init = 0) |>  
  add_variable("z2b", init = 0) |>
  add_variable("Bbd", init = 4389790) |>
  add_variable("Bbs", init = 4389790) |>
  add_variable("Bcbd", init = 4655690) |>
  add_variable("Bcbs", init = 4655690) |>
  add_variable("Bhd", init = 33439320) |>
  add_variable("Bhs", init = 33439320) |>
  add_variable("Bs", init = 42484800) |>
  add_variable("BLd", init = 840742) |>
  add_variable("BLs", init = 840742) |>
  add_variable("GD", init = 57728700) |>
  add_variable("Ekd", init = 5112.6001) |>
  add_variable("Eks", init = 5112.6001) |>
  add_variable("Hbd", init = 2025540) |>
  add_variable("Hbs", init = 2025540) |>
  add_variable("Hhd", init = 2630150) |>
  add_variable("Hhs", init = 2630150) |>
  add_variable("Hs", init = 4655690) |>
  add_variable("IN", init = 11585400) |>
  add_variable("INk", init = 2064890) |>
  add_variable("INke", init = 2405660) |>
  add_variable("INkt", init = 2064890) |>
  add_variable("K", init = 127486471) |>
  add_variable("Kk", init = 17774838) |>
  add_variable("Lfd", init = 15962900) |>
  add_variable("Lfs", init = 15962900) |>
  add_variable("Lhd", init = 21606600) |>
  add_variable("Lhs", init = 21606600) |>
  add_variable("Ls", init = 37569500) |>
  add_variable("Mh", init = 40510800) |>
  add_variable("Ms", init = 40510800) |>
  add_variable("OFb", init = 3474030) |>
  add_variable("OFbe", init = 3474030) |>
  add_variable("OFbt", init = 3638100) |>
  add_variable("V", init = 165438779) |>
  add_variable("Vfma", init = 159334599) |>
  add_variable("Vk", init = 23066350) |>
  add_variable("Vf", init = 31361792) |>
  add_variable("Rrbt", init = 0.03232) |>
  add_variable("Rrb", init = 0.03232) |>
  add_variable("z3") |>
  add_variable("z4") |>
  add_variable("z5") |>
  add_variable("z3a") |>
  add_variable("z3b") |>
  add_variable("HCe") |>
  add_variable("YDhs") |>
  add_variable("CG") |>
  add_variable("VfmaA") |>
  add_variable("Fcb") |>
  add_variable("FUbt")
```

Following that, replace an exogenous interest rate with an endogenous monetary-policy block. To do so, replace Eq. 11.84:

```
Rb ~ Rbbar
```

*(interest rate on bills set exogenously)* from the previous example with the policy rule for `Rb`:

```
Rb ~ (1 + Rrb) * (1 + PI) - 1
```

and add the two supporting equations for the target real rate and partial adjustment:

```
Rrbt ~ (1 + Rb)/(1 + PI) - 1
Rrb  ~ Rrb[-1] + epsrb * (Rrbt - Rrb[-1])
```

```{r}
# Add equations
model_growth2 <- model_growth2 |>
  add_equation("Yk = Ske + INke - INk[-1]", desc = "11.1 : Real output") |>
  add_equation("Ske = beta*Sk + (1-beta)*Sk[-1]*(1 + (GRpr + RA))", desc = "11.2 : Expected real sales") |>
  add_equation("INke = INk[-1] + gamma*(INkt - INk[-1])", desc = "11.3 : Long-run inventory target") |>
  add_equation("INkt = sigmat*Ske", desc = "11.4 : Short-run inventory target") |>
  add_equation("INk = INk[-1] + Yk - Sk - NPL/UC", desc = "11.5 : Actual real inventories") |>
  add_equation("Kk = Kk[-1]*(1 + GRk)", desc = "11.6 : Real capital stock") |>
  add_equation("GRk = gamma0 + gammau*U[-1] - gammar*RRl", desc = "11.7 : Growth of real capital stock") |>
  add_equation("U = Yk/Kk[-1]", desc = "11.8 : Capital utilization proxy") |>
  add_equation("RRl = ((1 + Rl)/(1 + PI)) - 1", desc = "11.9 : Real interest rate on loans") |>
  add_equation("PI = (P - P[-1])/P[-1]", desc = "11.10 : Rate of price inflation") |>
  add_equation("Ik = (Kk - Kk[-1]) + delta*Kk[-1]", desc = "11.11 : Real gross investment") |>
  
  # Box 11.2 : Firms equations
  # ---------------------------
  add_equation("Sk = Ck + Gk + Ik", desc = "11.12 : Actual real sales") |>
  add_equation("S = Sk*P", desc = "11.13 : Value of realized sales") |>
  add_equation("IN = INk*UC", desc = "11.14 : Inventories valued at current cost") |>
  add_equation("INV = Ik*P", desc = "11.15 : Nominal gross investment") |>
  add_equation("K = Kk*P", desc = "11.16 : Nomincal value of fixed capital") |>
  add_equation("Y = Sk*P + (INk - INk[-1])*UC", desc = "11.17 : Nomincal GDP") |>
  
  # Box 11.3 : Firms equations
  # ---------------------------
  # 11.18 : Real wage aspirations
  add_equation("omegat = exp(omega0 + omega1*log(PR) + omega2*log(ER + z3*(1 - ER) - z4*BANDt + z5*BANDb))", desc = "") |>
  add_equation("ER = N[-1]/Nfe[-1]", desc = "11.19 : Employment rate") |>
  # 11.20 : Switch variables
  add_equation("z3a = (ER > (1 - BANDb))", desc = "") |>
  add_equation("z3b = 1 - as.numeric(ER > (1 + BANDt))", desc = "") |>
  add_equation("z3 =  z3a * z3b", desc = "") |>
  add_equation("z4  = (ER >  (1 + BANDt))", desc = "") |>
  add_equation("z5  = (ER <  (1 - BANDb))", desc = "") |>
  add_equation("W = W[-1] + omega3*(omegat*P[-1] - W[-1])", desc = "11.21 : Nominal wage") |>
  add_equation("PR = PR[-1]*(1 + GRpr)", desc = "11.22 : Labor productivity") |>
  add_equation("Nt = Yk/PR", desc = "11.23 : Desired employment") |>
  add_equation("N = N[-1] + etan*(Nt - N[-1])", desc = "11.24 : Actual employment") |>
  add_equation("WB = N*W", desc = "11.25 : Nominal wage bill") |>
  add_equation("UC = WB/Yk", desc = "11.26 : Actual unit cost") |>
  add_equation("NUC = W/PR", desc = "11.27 : Normal unit cost") |>
  add_equation("NHUC = (1 - sigman)*NUC + sigman*(1 + Rln[-1])*NUC[-1]", desc = "11.28 : Normal historic unit cost") |>
  
  # Box 11.4 : Firms equations
  # ---------------------------
  add_equation("P = (1 + phi)*NHUC", desc = "11.29 : Normal-cost pricing") |>
  add_equation("phi = phi[-1] + eps2*(phit[-1] - phi[-1])", desc = "11.30 : Actual mark-up") |>
  # 11.31 : Ideal mark-up
  add_equation("phit = (FUft + FDf + Rl[-1]*(Lfd[-1] - IN[-1])) / ((1 - sigmase)*Ske*UC + (1 + Rl[-1])*sigmase*Ske*UC[-1])", desc = "") |>
  add_equation("HCe = (1 - sigmase)*Ske*UC + (1 + Rl[-1])*sigmase*Ske*UC[-1]", desc = "11.32 : Expected historical costs") |>
  add_equation("sigmase = INk[-1]/Ske", desc = "11.33 : Opening inventories to expected sales ratio") |>
  add_equation("Fft = FUft + FDf + Rl[-1]*(Lfd[-1] - IN[-1])", desc = "11.34 : Planned entrepeneurial profits of firms") |>
  add_equation("FUft = psiu*INV[-1]", desc = "11.35 : Planned retained earnings of firms") |>
  add_equation("FDf = psid*Ff[-1]", desc = "11.36 : Dividends of firms") |>
  
  # Box 11.5 : Firms equations
  # ---------------------------
  add_equation("Ff = S - WB + (IN - IN[-1]) - Rl[-1]*IN[-1]", desc = "11.37 : Realized entrepeneurial profits") |>
  add_equation("FUf = Ff - FDf - Rl[-1]*(Lfd[-1] - IN[-1]) + Rl[-1]*NPL", desc = "11.38 : Retained earnings of firms") |>
  # 11.39 : Demand for loans by firms
  add_equation("Lfd = Lfd[-1] + INV + (IN - IN[-1]) - FUf - (Eks - Eks[-1])*Pe - NPL", desc = "") |>
  add_equation("NPL = NPLk * Lfs[-1]", desc = "11.40 : Defaulted loans") |>
  add_equation("Eks = Eks[-1] + ((1 - psiu)*INV[-1])/Pe", desc = "11.41 : Supply of equities issued by firms") |>
  add_equation("Rk = FDf/(Pe[-1]*Ekd[-1])", desc = "11.42 : Dividend yield of firms") |>
  add_equation("PE = Pe/(Ff/Eks[-1])", desc = "11.43 : Price earnings ratio") |>
  add_equation("Q = (Eks*Pe + Lfd)/(K + IN)", desc = "11.44 : Tobins Q ratio") |>
  
  # Box 11.6 : Households equations
  # --------------------------------
  add_equation("YP = WB + FDf + FDb + Rm[-1]*Mh[-1] + Rb[-1]*Bhd[-1] + BLs[-1]", desc = "") |>
  add_equation("TX = theta*YP", desc = "11.46 : Income taxes") |>
  add_equation("YDr = YP - TX - Rl[-1]*Lhd[-1]", desc = "11.47 : Regular disposable income") |>
  add_equation("YDhs = YDr + CG", desc = "11.48 : Haig-Simons disposable income") |>
  # 11.49 : Capital gains
  add_equation("CG = (Pbl - Pbl[-1])*BLd[-1] + (Pe - Pe[-1])*Ekd[-1] + (OFb - OFb[-1])", desc = "") |>
  # 11.50 : Wealth
  add_equation("V = V[-1] + YDr - CONS + (Pbl - Pbl[-1])*BLd[-1] + (Pe - Pe[-1])*Ekd[-1] + (OFb - OFb[-1])", desc = "") |>
  add_equation("Vk = V/P", desc = "11.51 : Real stock of wealth") |>
  add_equation("CONS = Ck*P", desc = "11.52 : Consumption") |>
  add_equation("Ck = alpha1*(YDkre + NLk) + alpha2*Vk[-1]", desc = "11.53 : Real consumption") |>
  add_equation("YDkre = eps*YDkr + (1 - eps)*(YDkr[-1]*(1 + GRpr))", desc = "11.54 : Expected real regular disposable income") |>
  add_equation("YDkr = YDr/P - ((P - P[-1]) * Vk[-1])/P", desc = "11.55 : Real regular disposable income") |>
  
  # Box 11.7 : Households equations
  # --------------------------------
  add_equation("GL = eta*YDr", desc = "11.56 : Gross amount of new personal loans") |>
  add_equation("eta = eta0 - etar*RRl", desc = "11.57 : New loans to personal income ratio") |>
  add_equation("NL = GL - REP", desc = "11.58 : Net amount of new personal loans") |>
  add_equation("REP = deltarep*Lhd[-1]", desc = "11.59 : Personal loans repayments") |>
  add_equation("Lhd = Lhd[-1] + GL - REP", desc = "11.60 : Demand for personal loans") |>
  add_equation("NLk = NL/P", desc = "11.61 : Real amount of new personal loans") |>
  add_equation("BUR = (REP + Rl[-1] * Lhd[-1]) / YDr[-1]", desc = "11.62 : Burden of personal debt") |>
  
  # Box 11.8 : Households equations - portfolio decisions
  # -----------------------------------------------------
  # 11.64 : Demand for bills
  add_equation("Bhd = Vfma[-1]*(lambda20 + lambda22*Rb[-1] - lambda21*Rm[-1] - lambda24*Rk[-1] - lambda23*Rbl[-1] - lambda25*(YDr/V))", desc = "") |>
  # 11.65 : Demand for bonds
  add_equation("BLd = Vfma[-1]*(lambda30 - lambda32*Rb[-1] - lambda31*Rm[-1] - lambda34*Rk[-1] + lambda33*Rbl[-1] - lambda35*(YDr/V))/Pbl", desc = "") |>
  # 11.66 : Demand for equities - normalized to get the price of equitities
  add_equation("Pe = Vfma[-1]*(lambda40 - lambda42*Rb[-1] - lambda41*Rm[-1] + lambda44*Rk[-1] - lambda43*Rbl[-1] - lambda45*(YDr/V))/Ekd", desc = "") |>
  add_equation("Mh = Vfma - Bhd - Pe*Ekd - Pbl*BLd + Lhd", desc = "11.67 : Money deposits - as a residual") |>
  add_equation("Vfma = V - Hhd - OFb", desc = "11.68 : Investible wealth") |>
  add_equation("VfmaA = Mh + Bhd + Pbl * BLd + Pe * Ekd", desc = "") |>
  add_equation("Hhd = lambdac*CONS", desc = "11.69 : Households demand for cash") |>
  add_equation("Ekd = Eks", desc = "11.70 : Stock market equilibrium") |>
  
  # Box 11.9 : Governments equations
  # ---------------------------------
  add_equation("G = Gk*P", desc = "11.71 : Pure government expenditures") |>
  add_equation("Gk = Gk[-1]*(1 + GRg)", desc = "11.72 : Real government expenditures") |>
  add_equation("PSBR = G + BLs[-1] + Rb[-1]*(Bbs[-1] + Bhs[-1]) - TX", desc = "11.73 : Government deficit") |>
  # 11.74 : New issues of bills
  add_equation("Bs = Bs[-1] + G - TX - (BLs - BLs[-1])*Pbl + Rb[-1]*(Bhs[-1] + Bbs[-1]) + BLs[-1]", desc = "") |>
  add_equation("GD = Bbs + Bhs + BLs*Pbl + Hs", desc = "11.75 : Government debt") |>
  
  # Box 11.10 : The Central banks equations
  # ----------------------------------------
  add_equation("Fcb = Rb[-1]*Bcbd[-1]", desc = "11.76 : Central bank profits") |>
  add_equation("BLs = BLd", desc = "11.77 : Bonds are supplied on demand") |>
  add_equation("Bhs = Bhd", desc = "11.78 : Household bills supplied on demand") |>
  add_equation("Hhs = Hhd", desc = "11.79 : Cash supplied on demand") |>
  add_equation("Hbs = Hbd", desc = "11.80 : Reserves supplied on demand") |>
  add_equation("Hs = Hbs + Hhs", desc = "11.81 : Total supply of cash") |>
  add_equation("Bcbd = Hs", desc = "11.82 : Central bankd ") |>
  add_equation("Bcbs = Bcbd", desc = "11.83 : Supply of bills to Central bank") |>
  # Making monetary policy endogenous
  # Replace eq. 11.84 : "Rb = Rbbar" (Interest rate on bills set exogenously) with: "Rb ~  (1 + Rrb) * (1 + PI) - 1"
  add_equation("Rb = (1 + Rrb) * (1 + PI) - 1", desc = "") |>
  add_equation("Rrbt = (1 + Rb)/(1 + PI) - 1", desc = "") |>
  add_equation("Rrb = Rrb[-1] + epsrb * (Rrbt - Rrb[-1])", desc = "") |>
  add_equation("Rbl = Rb + ADDbl", desc = "11.85 : Long term interest rate") |>
  add_equation("Pbl = 1/Rbl", desc = "11.86 : Price of long-term bonds") |>
  
  # Box 11.11 : Commercial Banks equations
  # ---------------------------------------
  add_equation("Ms = Mh", desc = "11.87 : Bank deposits supplied on demand") |>
  add_equation("Lfs = Lfd", desc = "11.88 : Loans to firms supplied on demand") |>
  add_equation("Lhs = Lhd", desc = "11.89 : Personal loans supplied on demand") |>
  add_equation("Hbd = ro*Ms", desc = "11.90 Reserve requirements of banks") |>
  # 11.91 : Bills supplied to banks
  add_equation("Bbs = Bbs[-1] + (Bs - Bs[-1]) - (Bhs - Bhs[-1]) - (Bcbs - Bcbs[-1])", desc = "") |>
  # 11.92 : Balance sheet constraint of banks
  add_equation("Bbd = Ms + OFb - Lfs - Lhs - Hbd", desc = "") |>
  add_equation("BLR = Bbd/Ms", desc = "11.93 : Bank liquidity ratio") |>
  # 11.94 : Deposit interest rate
  add_equation("Rm = Rm[-1] + z1a*xim1 + z1b*xim2 - z2a*xim1 - z2b*xim2", desc = "") |>
  # 11.95-97 : Mechanism for determining changes to the interest rate on deposits
  add_equation("z2a = (BLR[-1] >  (top + 0.05))", desc = "") |>
  add_equation("z2b = (BLR[-1] >  top)", desc = "") |>
  add_equation("z1a = 1 - as.numeric(BLR[-1] > bot)", desc = "") |>
  add_equation("z1b = 1 - as.numeric(BLR[-1] > (bot - 0.05))", desc = "") |>
  
  # Box 11.12 : Commercial banks equations
  # ---------------------------------------
  add_equation("Rl = Rm + ADDl", desc = "11.98 : Loan interest rate") |>
  add_equation("OFbt = NCAR*(Lfs[-1] + Lhs[-1])", desc = "11.99 : Long-run own funds target") |>
  add_equation("OFbe = OFb[-1] + betab*(OFbt - OFb[-1])", desc = "11.100 : Short-run own funds target") |>
  add_equation("FUbt = OFbe - OFb[-1] + NPLke*Lfs[-1]", desc = "11.101 : Target retained earnings of banks") |>
  add_equation("NPLke = epsb*NPLke[-1] + (1 - epsb)*NPLk[-1]", desc = "11.102 : Expected proportion of non-performaing loans") |>
  add_equation("FDb = Fb - FUb", desc = "11.103 : Dividends of banks") |>
  add_equation("Fbt = lambdab*Y[-1] + (OFbe - OFb[-1] + NPLke*Lfs[-1])", desc = "11.104 : Target profits of banks") |>
  # 11.105 : Actual profits of banks
  add_equation("Fb = Rl[-1]*(Lfs[-1] + Lhs[-1] - NPL) + Rb[-1]*Bbd[-1] - Rm[-1]*Ms[-1]", desc = "") |>
  # 11.106 : Lending mark-up over deposit rate
  add_equation("ADDl = (Fbt - Rb[-1]*Bbd[-1] + Rm[-1]*(Ms[-1] - (1 - NPLke)*Lfs[-1] - Lhs[-1]))/((1 - NPLke)*Lfs[-1] + Lhs[-1])", desc = "") |>
  add_equation("FUb = Fb - lambdab*Y[-1]", desc = "11.107 : Actual retained earnings") |>
  add_equation("OFb = OFb[-1] + FUb - NPL", desc = "11.108 : Own funds of banks") |>
  add_equation("CAR = OFb/(Lfs + Lhs)", desc = "") |>
  add_equation("Vf = IN + K - Lfd - Ekd * Pe", desc = "Firm's wealth (memo for matrices)") |>
  add_equation("Ls = Lfs + Lhs", desc = "Loans supply (memo for matrices)") |>
  
  # Hidden equation
  # ---------------------------------------
  add_equation("Bbs = Bbd", desc = "", hidden = TRUE)

```

You can now simulate the baseline (again, 500 periods with Broyden in this example) and plot a few headline indicators:
```{r, out.width="100%"}
# Simulate model
model_growth2 <- simulate_scenario(model_growth2, scenario = "baseline", 
                                  max_iter = 350, periods = 500, tol = 1e-15,
                                  hidden_tol = 1e-6, rhtol = TRUE, method = "Broyden")

# Plot results
exprs <- c("Bsk = Bs / K", "GRk", "PI", "VK = V / K")
plots <- purrr::map(exprs,
             ~ plot_simulation(model = model_growth2, scenario = "baseline",
                               from = 1, to = 350, expressions = .x
             )
)
plotly::subplot(plots, nrows = 2, shareX = TRUE, titleX = TRUE)
```

#### Shock - permanent increase in government expenditures

Next, run a shock scenario where government expenditures grow faster for an extended period.

First, create an empty shock and set the new path for `GRg`:  
```{r}
# Create empty shock and add shock equation with increased government expenditures
shock_growth <- create_shock() |>
  add_shock(variable = "GRg", value = 0.033, start = 10, end = 100, desc = "")

# Create new scenario with this shock
model_growth2 <- model_growth2 |>
  add_scenario(name = "GRg_shock", origin = "baseline", shock = shock_growth)
```

Afterward, run the scenario:
```{r}
# Simulate shock
model_growth2 <- simulate_scenario(model_growth2, scenario = "GRg_shock", periods = 100, 
                                  max_iter = 350, tol = 1e-10, method = "Broyden",
                                  hidden_tol = 0.1, rhtol = TRUE)
```

With these results, you can visualize the growth rate of real output...
```{r, out.width="100%"}
# Figure 11.6A: Evolution of the growth rate of real output, with the growth rate of real pure government expenditures being forever higher than in the base line solution, when the central bank attempts to keep the real interest rate on bills at a constant level, but with a partial adjustment function
plot_simulation(model = model_growth2, scenario = c("GRg_shock"),
                from = 1, to = 80, expressions = c("GRy = (Yk / dplyr::lag(Yk)) - 1"))
```

...or, for example, check how interest rates move: the lending rate, the bill rate, and the deposit rate.
```{r, out.width="100%"}
# Figure 11.6D: Evolution of the lending rate, the bill rate and the deposit rate, with the growth rate of real pure government expenditures being forever higher than in the base line solution, when the central bank attempts to keep the employment rate at a constant level in a forward-looking manner
plot_simulation(model = model_growth2, scenario = c("GRg_shock"),
                from = 1, to = 80, expressions = c("Rl", "Rb", "Rm"))
```

### References
For the full model discussion and interpretation of the GROWTH model, refer to Chapter 11 of *Monetary Economics. An Integrated Approach to Credit, Money, Income, Production and Wealth.*
