---
title: "godley — package outline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{package outline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Scenario simulation
There are few steps that should be process in order to create and simulate baseline scenario. \
Function ```create_model()``` creates a SFC object. \
Functions ```add_variable()``` and ```add_equation()``` add attributes about variables and equations to the model as a new lines in SFC tibble. \
Function ```simulate_scenario()``` prepares model for simulation and calculates final matrix for provided scenarios. \

### 1. Model creation
To create SFC model (SFC object) please use ```create_model()``` function as in example below:
```{r}
library(godley)
model_sim <- create_model(name = "SFC model")
```
where: \
name — model name (required) \
template — name of model template to upload (optional) \
Note: Instead of manually adding variables and equations to the model (described below in steps 2. and 3.) user can upload model's attributes from templates available in the package. Available model templates are: SIM, SIMEX, PC, LP, REG, OPEN, BMW, BMWK, DIS and DISINF. 

### 2. Adding variables
To add variables to the SFC model please use ```add_variable()``` function as in example below:
```{r}
# add parameters
model_sim <- add_variable(model = model_sim, name = "C_d", desc = "Consumption demand by households")
model_sim <- add_variable(model = model_sim, name = "C_s", desc = "Consumption supply")
model_sim <- add_variable(model = model_sim, name = "G_s", desc = "Government supply")
model_sim <- add_variable(model = model_sim, name = "H_h", desc = "Cash money held by households")
model_sim <- add_variable(
  model = model_sim, name = "H_s",
  desc = "Cash money supplied by the government"
)
model_sim <- add_variable(model = model_sim, name = "N_d", desc = "Demand for labor")
model_sim <- add_variable(model = model_sim, name = "N_s", desc = "Supply of labor")
model_sim <- add_variable(model = model_sim, name = "T_d", desc = "Taxes, demand")
model_sim <- add_variable(model = model_sim, name = "T_s", desc = "Taxes, supply")
model_sim <- add_variable(model = model_sim, name = "Y", desc = "Income = GDP")
model_sim <- add_variable(model = model_sim, name = "Yd", desc = "Disposable income of households")
model_sim <- add_variable(
  model = model_sim, name = "alpha1", init = 0.6,
  desc = "Propensity to consume out of income"
)
model_sim <- add_variable(
  model = model_sim, name = "alpha2", init = 0.4,
  desc = "Propensity to consume out of wealth"
)
model_sim <- add_variable(model = model_sim, name = "theta", init = 0.2, desc = "Tax rate")
model_sim <- add_variable(model = model_sim, name = "G_d", init = 20, desc = "Government demand")
model_sim <- add_variable(model = model_sim, name = "W", init = 1, desc = "Wage rate")
```
where: \
model — model name (required) \
name — variable name (required) \
init — initial variable value (optional) \
desc — variable description (optional) 

### 3. Adding equations
To add equation to the SFC model please use ```add_equation()``` function as in example below:
```{r}
# add equations
model_sim <- add_equation(model = model_sim, equation = "C_s = C_d", desc = "Consumption")
model_sim <- add_equation(model = model_sim, equation = "G_s = G_d")
model_sim <- add_equation(model = model_sim, equation = "T_s = T_d")
model_sim <- add_equation(model = model_sim, equation = "N_s = N_d")
model_sim <- add_equation(model = model_sim, equation = "Yd = W * N_s - T_s")
model_sim <- add_equation(model = model_sim, equation = "T_d = theta * W * N_s")
model_sim <- add_equation(model = model_sim, equation = "C_d = alpha1 * Yd + alpha2 * H_h[-1]")
model_sim <- add_equation(model = model_sim, equation = "H_s = G_d - T_d + H_s[-1]")
model_sim <- add_equation(model = model_sim, equation = "H_h = Yd - C_d + H_h[-1]")
model_sim <- add_equation(model = model_sim, equation = "Y = C_s + G_s")
model_sim <- add_equation(model = model_sim, equation = "N_d = Y/W")
model_sim <- add_equation(
  model = model_sim, equation = "H_s = H_h", desc = "Money equilibrium",
  hidden = TRUE
)
```
where: \
model — model name (required) \
equation —  equation formula (required) \
desc — equation description (optional)\
hidden — boolean argument that inform if equation should be hidden (optional) \
Note: Hidden equation is an equation that is not simulate.

### 4. Change initial values (optional)
To change initial values of parameters please use ```change_init()``` function as in example below:
```{r}
# change initial value for alpha1 parameter
model_sim <- change_init(model = model_sim, name = "alpha1", value = 0.5)
```
where: \
model — model name (required) \
name —  name of parameter that will be changed (required) \
value — new initial value for provided parameter (required)

### 5. Simulating scenario(s)
To simulate model please use ```simulate_scenario()``` function as in example below:
```{r}
# simulate baseline scenario
model_sim <- simulate_scenario(
  model = model_sim, scenario = "baseline", max_iter = 350, periods = 100,
  hidden_tol = 0.1, tol = 1e-08, method = "Gauss"
)
```
where: \
model — name of the SFC model (required) \
scenario — name of scenario that will be calculated (optional) \
max_iter — number of iteration in simulation (optional) \
periods — number of simulating periods in simulation (optional) \
hidden_tol — maximum error tolerated for hidden equations in simulation (optional) \
tol — maximum error tolerated for algorithm until algorithm converge in simulation (optional) \
method — 'Newton' or 'Gauss' calculation method in simulation (optional) \
Note: If scenario name is not provided all not simulated scenarios will be calculated.

## Model — plots
To visualize model simulation please use ```plot_simulation()``` function as in example below:
```{r, out.width="100%"}
# plot results
plot_simulation(
  model = model_sim, scenario = "baseline", from = 1, to = 100,
  expressions = c("Y", "C_s / alpha1")
)
```
where: \
model — name of SFC model that will be displayed on the chart (required) \
scenario — name of scenario(s) that will be displayed on the chart (optional) \
take_all — boolean indicating whether all scenarios containing the given scenario name strings should be displayed on the chart (optional) \
from — first period that will be displayed on the chart (optional) \
to — last period that will be displayed on the chart (optional) \
expressions — variable names or equations that will be displayed on the chart (optional) 

## Model — shock
There are few steps that should be process in order to create and simulate shock scenario. \
Function ```create_shock()``` creates a SFC shock tibble object. \
Function ```add_shock()``` adds shock equations to the SFC shock object. \
Function ```add_scenario()```  adds shock scenario to the SFC model. \
Function ```simulate_scenario()``` calculates shock scenario(s). \

### 1. Creating shock object
To create SFC shock object please use ```create_shock()``` function as in example below:
```{r}
# create shock
sim_shock <- create_shock()
```

### 2. Adding shock equations
To add shock equation to SFC shock object please use ```add_shock()``` function as in example below:
```{r}
# add shock equation for increased government expenditures
sim_shock <- add_shock(
  shock = sim_shock,
  variable = "G_d",
  value = 25,
  desc = "Increase in government expenditures", start = 5, end = 10
)
```
where: \
shock — name of the shock object to which new equation will be added (required) \
equation — new equation describing the applied shock (required) \
desc — description of the applied shock (optional) \
start — first period when shock will be applied (optional) \
end — last period when shock will be applied (optional) 
Note: 'start' and 'end' parameters are set by default to NA, so by default the shock will be applied to the whole simulation period

### 3. Adding shock scenario to the SFC model
To add new shock scenario to SFC model please use ```add_scenario()``` function as in example below:

```{r}
# add shock scenario for increased government expenditures
model_sim <- add_scenario(
  model = model_sim, name = "expansion", origin = "baseline",
  origin_start = 1,
  origin_end = 100,
  shock = sim_shock
)
```
where: 
model — name of the model to which new scenario will be added (required) \
name — name of a new scenario (optional) \
origin — name of baseline scenario (optional) \
origin_period — period from the baseline scenario which will be used as initial period in new scenario (optional) \
shock — name of created SFC_shock object (required)

### 4. Simulating shock scenario
To simulate new shock scenario please use ```simulate_scenario()``` function:

```{r}
# calculate shock scenario for increased government expenditures
model_sim <- simulate_scenario(
  model = model_sim, max_iter = 350, periods = 100, hidden_tol = 0.1,
  tol = 1e-08, method = "Gauss"
)
```

## Model — sensitivity
Sensitivity allows to create in separate SFC model multiple scenarios with changed initial value for selected variable. To create and simulate sensitivity please use ```create_sensitivity()``` and  ```simulate_scenario()``` functions as in example below:
 
```{r}
# create and calculate sensitivity
model_sen <- create_sensitivity(
  model_pass = model_sim, variable = "alpha1",
  lower = 0.1, upper = 0.8, step = 0.1
)

model_sen <- simulate_scenario(
  model = model_sen, max_iter = 350, periods = 100, hidden_tol = 0.1,
  tol = 1e-08, method = "Gauss"
)
```
where: \
model_pass — name of the SFC model or name of scenario template that will be used in sensitivity calculation (required) \
variable — name of the variable that will be check in sensitivity calculation (required) \
lower — minimum value of the variable that will be used in sensitivity calculation (optional) \
upper — maximum value of the variable that will be used in sensitivity calculation (optional) \
step — attribute describing by how many step the variable will be changed in sensitivity analysis (optional)
